### 2019-11-22 FEM_Stata/Estimation/estimate_summary-swv.tex

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{mathpazo,color}

\StataweaveOpts{echo=false, outfmt="frame=single, fontfamily=courier, fontsize=\small"}

\begin{document}

\begin{Statacode}{hide}
set more off
include ../../fem_env.do

*set up work directory
global estdir ../Estimates

*load estimates 
**Disease
estimates use $estdir/hearte.ster
estimates store Heart_disease
estimates use $estdir/stroke.ster
estimates store Stroke
estimates use $estdir/hibpe.ster
estimates store HBP
estimates use $estdir/diabe.ster
estimates store Diabetes
estimates use $estdir/lunge.ster
estimates store Lung
estimates use $estdir/memrye.ster
estimates store Memory_disorder

**Functional Status

estimates use $estdir/died.ster
estimates store Motality
estimates use $estdir/nhmliv.ster
estimates store Nursing_home
estimates use $estdir/iadlstat.ster
estimates store IADL
estimates use $estdir/adlstat.ster
estimates store ADL

**Predicted Medical Costs

estimates use $estdir/totmd_mcbs.ster
estimates store Total_MCBS
estimates use $estdir/totmd_meps.ster
estimates store Total_MEPS
estimates use $estdir/oopmd_meps.ster
estimates store OOP_MEPS
estimates use $estdir/mcare_pta.ster
estimates store Medicare_A
estimates use $estdir/mcare_ptb.ster
estimates store Medicare_B

**Program Participation

estimates use $estdir/ssclaim.ster 
estimates store SS_claim
estimates use $estdir/ssiclaim.ster
estimates store SSI_claim
estimates use $estdir/diclaim.ster 
estimates store DI_claim
estimates use $estdir/mcareb_takeup_newenroll.ster
estimates store MedicareB_new
estimates use $estdir/mcareb_takeup_curenroll.ster
estimates store MedicareB_current
estimates use $estdir/mcare_ptd.ster
estimates store Medicare_D

**Labor Outcomes

estimates use $estdir/work.ster
estimates store Working
estimates use $estdir/dbclaim.ster
estimates store DB_pension_claiming

\end{Statacode}

\section*{\centering Type of Estimators}

\begin{Statacode}

estimates dir

\end{Statacode}

\pagebreak
\section*{\centering Summary Characteristics}

\begin{Statacode}

estimates stat _all

\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Diseases}
\begin{Statacode}
estimates table Heart_disease Stroke HBP, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Diseases(continued)}
\begin{Statacode}
estimates table Diabetes Lung Memory_disorder, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Functional Status}
\begin{Statacode}
estimates table Motality Nursing_home IADL ADL, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Predicted Medical Costs}
\begin{Statacode} 
estimates table Total_MCBS Total_MEPS , star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Predicted Medical Costs(Continued)}
\begin{Statacode} 
estimates table OOP_MEPS Medicare_A Medicare_B, star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Program Participation}
\begin{Statacode}
estimates table SS_claim SSI_claim DI_claim, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Program Participation(Continued)}
\begin{Statacode}
estimates table MedicareB_new MedicareB_current Medicare_D, star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl1 adl2 adl3p widowed work) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}
estimates table Working DB_pension_claiming, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Coeffient Point Estimates}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}
estimates table Working DB_pension_claiming, star(.05 .01 .001) keep(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) sty(columns)
\end{Statacode}

\begin{Statacode}
qui set mem 300m
qui use $outdata/hrs17_transition.dta

*logdeltaage
qui gen logdeltaage = log(age_iwe - lage_iwe)

*llogbmi_l30 llogbmi_30p flogbmi_l30 flogbmi_30p
qui {
  local log_30 = log(30)
  mkspline llogbmi_l30 `log_30' llogbmi_30p = llogbmi
  mkspline flogbmi_l30 `log_30' flogbmi_30p = flogbmi
}

*flogq flogaime

qui gen flogq=0
qui gen flogaime=0

local age_var age_iwe

gen lage62e = floor(l`age_var') == 60 if l`age_var' < .
gen lage63e = floor(l`age_var') == 61 if l`age_var' < .

mkspline la6 58 la7 73 la7p = l`age_var'

*** GENERATE WEAVE DUMMIES
gen w3 = wave == 3
gen w4 = wave == 4
gen w5 = wave == 5
gen w6 = wave == 6
gen w7 = wave == 7

\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Diseases}
\begin{Statacode}
**marginal effects
*desease
qui estimates restore Heart_disease
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fstroke fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix heart=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\.\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Stroke
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fcancre fhibpe fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix stroke=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\.\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore HBP
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fdiabe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix hbp=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\.\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Diabetes
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe flunge fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix diab=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\.\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Lung
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix lung=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\.\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

qui estimates restore Memory_disorder
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix mem=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\.\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16])
matrix drop temp temp1

matrix disease=(heart,stroke,hbp,diab,lung,mem)
matrix colnames disease = "heart disease""stroke""hypertension""diabetes""lung disease""memory disorder"
matrix rownames disease = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "smoke" "widowed" "single" "bmi<30" "bim>30" "work"
matlist disease, lines(oneline) twidth(15)  
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Functional Health}
\begin{Statacode}
qui estimates restore Motality
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix mortality=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore Nursing_home
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix nursingh=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore IADL
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix iadl=(temp,temp1)
matrix iadl=iadl'
*matrix iadl=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

qui estimates restore ADL
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fsmoken fwidowed fsingle flogbmi_l30 flogbmi_30p fwork) noesample
matrix temp1=r(b)
matrix adl=(temp,temp1)
matrix adl=adl'
*matrix iadl=(temp\temp1[1,1]\temp1[1,2]\temp1[1,3]\temp1[1,4]\temp1[1,5]\temp1[1,6]\temp1[1,7]\temp1[1,8]\temp1[1,9]\temp1[1,10]\temp1[1,11]\temp1[1,12]\temp1[1,13]\temp1[1,14]\temp1[1,15]\temp1[1,16]\temp1[1,17]\.\.\temp1[1,18])
matrix drop temp temp1

matrix functional=(mortality, nursingh, iadl, adl)
matrix colnames functional = "mortality""nursing home""iadl""adl"
matrix rownames functional = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "smoke" "widowed" "single" "work"
matlist functional, lines(oneline) twidth(15)  
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Predicted Medical Costs}
\begin{Statacode} 
estimates table Total_MCBS Total_MEPS , star(.05 .01 .001) keep(black hispan hsless college male hearte stroke cancre hibpe diabe lunge adl3p widowed single) sty(columns)
\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Program Participation}
\begin{Statacode}

qui estimates restore SS_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix ss=(temp,temp1)
matrix ss=ss'
matrix drop temp temp1

qui estimates restore SSI_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix ssi=(temp,temp1)
matrix ssi=ssi'

qui estimates restore DI_claim
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix diclaim=(temp,temp1)
matrix diclaim=diclaim'
matrix drop temp temp1

matrix programp=(ss, ssi, diclaim)
matrix colnames programp = "ss claim""ssi claim""di claim"
matrix rownames programp = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "stroke" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "widowed" "single" "work"
matlist programp, lines(oneline) twidth(15)

\end{Statacode}

\pagebreak
\section*{\centering Marginal Effects}
\subsection*{\centering Labor Outcomes}
\begin{Statacode}

qui estimates restore Working
margins, noesample
matrix temp=r(b)
margins, dydx(black hispan hsless college male fhearte fstroke fcancre fhibpe fdiabe flunge fadl1 fadl2 fadl3p fwidowed fsingle fwork) noesample
matrix temp1=r(b)
matrix working=(temp,temp1)
matrix working=working'

matrix labor=(working)
matrix colnames labor = "working"
matrix rownames labor = "mean""black" "hispanic" "<high school" "college" "male" "heart disease" "cancer" "hypertension" "diabetes" "lung disease" "adl=1" "adl=2" "adl>=3" "smoke" "widowed" "single" "work"
matlist labor, lines(oneline) twidth(15)

\end{Statacode}
\end{document}