### 2019-11-22 analysis/depratios-swv.tex

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{url}

\title{Dependency Ratios}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}{!echo,hide}
library(ggplot2)
library(foreign)
\end{Rcode}

Population projections are from \url{http://www.census.gov/population/www/projections/files/nation/download/NP2008_D1.csv}

\begin{Rcode}
pop <- read.csv('NP2008_D1.csv')
pop <- subset(pop, HISP==0 & RACE==0 & SEX==0)
vec1864 <- paste('POP_', 18:64, sep='')
vec1850 <- paste('POP_', 18:50, sep='')
vec65p <- paste('POP_', 65:100, sep='')
pop1864 <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec1864)
pop1850 <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec1850)
pop65p <- by(pop, pop[,'YEAR'], function(x, a.names) sum(x[,a.names]), a.names=vec65p)
pops.df <- data.frame(pop1864=unlist(as.list(pop1864)), pop1850=unlist(as.list(pop1850)), pop65p=unlist(as.list(pop65p)))
row.names(pops.df) <- names(as.list(pop1864))
print(pops.df)
\end{Rcode}

\begin{Rcode}
summ <- read.dta('../output/depratios/depratios_summary.dta')
row.names(summ) <- summ[,'year']
summ[,'pop1864'] <- pops.df[row.names(summ),1]/1e6
summ[,'pop1850'] <- pops.df[row.names(summ),2]/1e6
summ[,'pop65p'] <- pops.df[row.names(summ),3]/1e6
summ <- subset(summ, year >= 2010 & !is.na(pop1864))
\end{Rcode}

\begin{Rcode}{savefig,label=standard}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM Simulation')) + 
geom_line(aes(y=pop65p/pop1864, colour='Census')) +
opts(legend.position='bottom') + labs(y='OADR', colour='65+ Population Projection')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{standard}
\caption{OADR: Age 65+ over Ages 18-64}
\end{figure}

\begin{Rcode}{savefig,label=qalystats}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=qaly_75Q, colour='75th Percentile')) + 
geom_line(aes(y=qaly_50Q, colour='50th Percentile')) + 
geom_line(aes(y=qaly, colour='Mean')) +
geom_line(aes(y=qaly_25Q, colour='25th Percentile')) + 
geom_line(aes(y=qaly_10Q, colour='10th Percentile')) + 
opts(legend.position='bottom') + labs(colour='QALY Statistic', y='QALY')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{qalystats}
\caption{QALY Over Time}
\end{figure}

\begin{Rcode}{savefig,label=qaly65}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_65/pop1864, colour='QALY <= 0.80')) + 
geom_line(aes(y=ttl_qaly_under7_65/pop1864, colour='QALY <= 0.70')) + 
geom_line(aes(y=ttl_qaly_under6_65/pop1864, colour='QALY <= 0.60')) + 
geom_line(aes(y=ttl_qaly_under5_65/pop1864, colour='QALY <= 0.50')) +
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qaly51}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_51/pop1850, colour='QALY <= 0.80')) +
geom_line(aes(y=ttl_qaly_under7_51/pop1850, colour='QALY <= 0.70')) + 
geom_line(aes(y=ttl_qaly_under6_51/pop1850, colour='QALY <= 0.60')) + 
geom_line(aes(y=ttl_qaly_under5_51/pop1850, colour='QALY <= 0.50')) +
 opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qaly51nw}
ggplot(summ, aes(x=year)) + 
geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=ttl_qaly_under8_51_nw/pop1850, colour='QALY <= 0.80 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under7_51_nw/pop1850, colour='QALY <= 0.70 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under6_51_nw/pop1850, colour='QALY <= 0.60 and Not Working')) + 
geom_line(aes(y=ttl_qaly_under5_51_nw/pop1850, colour='QALY <= 0.50 and Not Working')) +
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=disabled}
ggplot(summ, aes(x=year)) + geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=pop_disability/pop1850, colour='Age 51+ and Any Disability')) + 
geom_line(aes(y=ttl_adl3p/pop1850, colour='Age 51+ and ADL 3 or More')) + opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{Rcode}{savefig,label=qalyweighting}
ggplot(summ, aes(x=year)) + geom_line(aes(y=pop_medicare_age/pop1864, colour='FEM OADR')) + 
geom_line(aes(y=(pop_medicare_age - ttl_qaly65)/pop1864, colour='QALY-weighted FEM OADR')) + 
opts(legend.position='bottom') + labs(colour='Dependency Measure', y='Dependency Ratio')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{qaly65}
\caption{QALY Thresholds on Age 65+}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qaly51}
\caption{QALY Thresholds on Age 51+}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qaly51nw}
\caption{QALY Thresholds on Age 51+ and Not Working}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{disabled}
\caption{Different Definitions of Disability}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{qalyweighting}
\caption{QALY-weighted Population Ratios}
\end{figure}

\begin{Rcode}{savefig,label=prevelance2010a}
s <- subset(summ, year==2010)
xes <- seq(0.85, 0.35, -0.1)
buckets <- seq(90,40,-10)

cs <- as.numeric(s[,paste('cancre_q',buckets, sep='')])
ds <- as.numeric(s[,paste('diabe_q', buckets, sep='')])
hes <- as.numeric(s[,paste('hearte_q', buckets, sep='')])
his <- as.numeric(s[,paste('hibpe_q', buckets, sep='')])
ls <- as.numeric(s[,paste('lunge_q', buckets, sep='')])
ss <- as.numeric(s[,paste('stroke_q', buckets, sep='')])
is <- as.numeric(s[,paste('iadl1_q', buckets, sep='')])
a1s <- as.numeric(s[,paste('adl1_q', buckets, sep='')])
a3s <- as.numeric(s[,paste('adl3p_q', buckets, sep='')])
ns <- as.numeric(s[,paste('nhmliv_q', buckets, sep='')])
bs <- as.numeric(s[,paste('bmiGT30_q', buckets, sep='')])

ggplot(s) +
geom_line(aes(x=xes, y=cs, colour='Cancer'), na.rm=T) +
geom_line(aes(x=xes, y=ds, colour='Diabetes'), na.rm=T) +
geom_line(aes(x=xes, y=hes, colour='Heart Disease'), na.rm=T) +
geom_line(aes(x=xes, y=his, colour='High Blood Pressure'), na.rm=T) +
geom_line(aes(x=xes, y=ls, colour='Lung Disease'), na.rm=T) +
geom_line(aes(x=xes, y=ss, colour='Stroke'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{Rcode}{savefig,label=prevelance2010b}
ggplot(s) +
geom_line(aes(x=xes, y=is, colour='IADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a1s, colour='ADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a3s, colour='ADL 3+'), na.rm=T) +
geom_line(aes(x=xes, y=ns, colour='Nursing Home'), na.rm=T) +
geom_line(aes(x=xes, y=bs, colour='BMI > 30'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelance2010a}
\caption{Prevelance of Disease by QALY in 2010}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{prevelance2010b}
\caption{Prevelance of Health Conditions by QALY in 2010}
\end{figure}

\begin{Rcode}{savefig,label=prevelance2050a}
s <- subset(summ, year==2050)

cs <- as.numeric(s[,paste('cancre_q',buckets, sep='')])
ds <- as.numeric(s[,paste('diabe_q', buckets, sep='')])
hes <- as.numeric(s[,paste('hearte_q', buckets, sep='')])
his <- as.numeric(s[,paste('hibpe_q', buckets, sep='')])
ls <- as.numeric(s[,paste('lunge_q', buckets, sep='')])
ss <- as.numeric(s[,paste('stroke_q', buckets, sep='')])
is <- as.numeric(s[,paste('iadl1_q', buckets, sep='')])
a1s <- as.numeric(s[,paste('adl1_q', buckets, sep='')])
a3s <- as.numeric(s[,paste('adl3p_q', buckets, sep='')])
ns <- as.numeric(s[,paste('nhmliv_q', buckets, sep='')])
bs <- as.numeric(s[,paste('bmiGT30_q', buckets, sep='')])

ggplot(s) +
geom_line(aes(x=xes, y=cs, colour='Cancer'), na.rm=T) +
geom_line(aes(x=xes, y=ds, colour='Diabetes'), na.rm=T) +
geom_line(aes(x=xes, y=hes, colour='Heart Disease'), na.rm=T) +
geom_line(aes(x=xes, y=his, colour='High Blood Pressure'), na.rm=T) +
geom_line(aes(x=xes, y=ls, colour='Lung Disease'), na.rm=T) +
geom_line(aes(x=xes, y=ss, colour='Stroke'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{Rcode}{savefig,label=prevelance2050b}
ggplot(s) +
geom_line(aes(x=xes, y=is, colour='IADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a1s, colour='ADL 1'), na.rm=T) +
geom_line(aes(x=xes, y=a3s, colour='ADL 3+'), na.rm=T) +
geom_line(aes(x=xes, y=ns, colour='Nursing Home'), na.rm=T) +
geom_line(aes(x=xes, y=bs, colour='BMI > 30'), na.rm=T) +
opts(legend.position='bottom') +
xlab('QALY Bucket') +
ylab('Prevelance') +
scale_x_continuous(trans='reverse') +
scale_y_continuous(breaks=seq(0,1,0.2)) +
labs(colour='Health Condition')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelance2050a}
\caption{Prevelance of Disease by QALY in 2050}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{prevelance2050b}
\caption{Prevelance of Health Conditions by QALY in 2050}
\end{figure}

\begin{Rcode}{savefig,label=prevelanceevolutionA}
conds <- c('cancre','diabe','hearte','hibpe','lunge','stroke')
condlist <- c(paste(conds, '_qLT60', sep=''), paste(conds, '_qLT70', sep=''), paste(conds, '_qLT80', sep=''), paste(conds, '_qLT90', sep=''))

dat.l <- lapply(condlist, FUN=function(c, s) data.frame(year=as.numeric(s[,'year']),
                                                        val=s[,c],
							qaly.thresh=paste('QALY <= 0.', substr(c,nchar(c)-1, nchar(c)), sep=''),
							condition=strsplit(c,'_')[[1]][1]), s=summ)
dat <- do.call(rbind, dat.l)
ggplot(dat, aes(x=year, y=val)) + 
geom_line(aes(colour=condition)) +
scale_y_continuous(limits=c(0,1), breaks=seq(0,1,0.2)) +
facet_wrap( ~ qaly.thresh) +
labs(colour='Health Condition', x='Year', y='Prevelance') +
opts(legend.position='bottom')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelanceevolutionA}
\caption{Evolution of Prevelance by QALY Upper Threshold}
\end{figure}

\begin{Rcode}{savefig,label=prevelanceevolutionB}
conds <- c('iadl1','adl1','adl3p','nhmliv', 'bmiGT30')
condlist <- c(paste(conds, '_qLT60', sep=''), paste(conds, '_qLT70', sep=''), paste(conds, '_qLT80', sep=''), paste(conds, '_qLT90', sep=''))

dat.l <- lapply(condlist, FUN=function(c, s) data.frame(year=as.numeric(s[,'year']),
                                                        val=s[,c],
							qaly.thresh=paste('QALY <= 0.', substr(c,nchar(c)-1, nchar(c)), sep=''),
							condition=strsplit(c,'_')[[1]][1]), s=summ)
dat <- do.call(rbind, dat.l)
ggplot(dat, aes(x=year, y=val)) + 
geom_line(aes(colour=condition)) +
scale_y_continuous(limits=c(0,1), breaks=seq(0,1,0.2)) +
facet_wrap( ~ qaly.thresh) +
labs(colour='Health Condition', x='Year', y='Prevelance') +
opts(legend.position='bottom')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{prevelanceevolutionB}
\caption{Evolution of Prevelance by QALY Upper Threshold}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/depratios/depratios_summary.dta
keep year ttl_qaly_u* ttl_qaly_a*
drop *_nw
outsheet using depratios_survivor.csv, comma replace
\end{Statacode}

\end{document}