### 2019-11-22 analysis/medicare-swv.tex

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{datatool}

\title{Analysis of Medicare Reform Options}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}{!echo,hide}
library(foreign)
library(ggplot2)
\end{Rcode}

\begin{Statacode}{!echo,hide}
set linesize 100
global discount_rate = 0.03
\end{Statacode}

\section{Summary of Results}

Figure~\ref{fig:summary} shows a comparison of the costs of various Medicare reform scenarios. Each variant shows a particular parameter set for the examined scenarios. A complete listing of the excursions and their parameters is in Table~\ref{tab:scenarios}. The ACA schedule for Part B premium thresholds freezes them in nominal dollars until 2019 then allows them to rise with inflation, while the nominal dollars scenario never adjusts for inflation and the real dollars scenario always does.

\begin{Statacode}{!echo,hide,fig,savefig,label=aca_totmd}
use ../output/ACA/summary, clear
tsset year
tsfill
foreach v of varlist ttl_totmd_mcare {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist ttl_totmd_mcare {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected ttl_totmd_mcare year
outsheet year ttl_totmd_mcare using aca_totmd.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{aca_totmd}
\caption{NPV of Total Medical Expenditures for Medicare Eligible}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=cost_summary}
use ../output/ACA/summary, clear
gen summ_ACA = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_ACAGross = ttl_mcare
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
gen summ_nominal = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_nominalGross = ttl_mcare
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
gen summ_drop00 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop00Gross = ttl_mcare
label variable summ_drop00 "Part A Drop 0 ACA Thresholds"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
gen summ_drop03 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop03Gross = ttl_mcare
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
gen summ_drop0310 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop0310Gross = ttl_mcare
label variable summ_drop0310 "Part A Drop 3, 10 Percent"
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
gen summ_drop05 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_drop05Gross = ttl_mcare
label variable summ_drop05 "Part A Drop 5 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
gen summ_gdp = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
gen summ_gdpGross = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace summ_gdp = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium if year < 2012
replace summ_gdpGross = ttl_mcare if year < 2012
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
gen summ_cpi = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
gen summ_cpiGross = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace summ_cpi = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium if year < 2012
replace summ_cpiGross = ttl_mcare if year < 2012
label variable summ_cpi "Vouchers CPI"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
gen summ_age67 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age67Gross = ttl_mcare
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/elig.age70/summary, sort
gen summ_age70 = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age70Gross = ttl_mcare
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
gen summ_acanolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_acanolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
gen summ_partAnolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_partAnolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
gen summ_gdpnolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_gdpnolimitGross = ttl_mcare
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
gen summ_age67nolimit = ttl_mcare - ttl_mcare_ptb_premium - ttl_mcare_pta_premium
gen summ_age67nolimitGross = ttl_mcare
keep year summ_*

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Trillions of 2010 Dollars")
outsheet year summ_* using cost_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{cost_summary}
\caption{Net Present Value of Medicare Reform Options}
\label{fig:summary}
\end{figure}

\begin{table}[ht]
\centering
\caption{Net Present Value of Reform Options (Trillions 2010 Dollars)}
\recallout{cumm_cost}
\label{tab:cumm_cost}
\end{table}

\begin{Statacode}{!echo,hide,fig,savefig,label=enroll_a_summary}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll summ_ACA
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
rename ttl_mcare_pta_enroll summ_nominal
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
rename ttl_mcare_pta_enroll summ_drop03
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
rename ttl_mcare_pta_enroll summ_gdp
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
rename ttl_mcare_pta_enroll summ_age67
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_pta_enroll summ_cpi
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
rename ttl_mcare_pta_enroll summ_drop00
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
rename ttl_mcare_pta_enroll summ_drop0310
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
rename ttl_mcare_pta_enroll summ_drop05
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_mcare_pta_enroll summ_ACAnolimit
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_mcare_pta_enroll summ_partAnolimit
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_mcare_pta_enroll summ_gdpnolimit
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_mcare_pta_enroll summ_age67nolimit
keep year summ_*

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("People Enrolled in Part A (Millions)")
outsheet year summ_* using enroll_a_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{enroll_a_summary}
\caption{Annual Enrollment in Medicare Part A}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=enroll_b_summary}
use ../output/ACA/summary, clear
rename ttl_mcare_ptb_enroll summ_ACA
label variable summ_ACA "ACA"
keep year summ_*
merge year using ../output/partB.nominal.dollars/summary, sort
rename ttl_mcare_ptb_enroll summ_nominal
label variable summ_nominal "Part B Nominal"
keep year summ_*
merge year using ../output/partA.drop03/summary, sort
rename ttl_mcare_ptb_enroll summ_drop03
label variable summ_drop03 "Part A Drop 3 ACA Thresholds"
keep year summ_*
merge year using ../output/voucher.gdp/summary, sort
rename ttl_mcare_ptb_enroll summ_gdp
label variable summ_gdp "Vouchers GDP"
keep year summ_*
merge year using ../output/elig.age67/summary, sort
rename ttl_mcare_ptb_enroll summ_age67
label variable summ_age67 "Eligibility Age 67"
keep year summ_*
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_ptb_enroll summ_cpi
keep year summ_*
merge year using ../output/partA.drop00/summary, sort
rename ttl_mcare_ptb_enroll summ_drop00
keep year summ_*
merge year using ../output/partA.drop03.10percent/summary, sort
rename ttl_mcare_ptb_enroll summ_drop0310
keep year summ_*
merge year using ../output/partA.drop05/summary, sort
rename ttl_mcare_ptb_enroll summ_drop05
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_ACAnolimit
keep year summ_*
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_partAnolimit
keep year summ_*
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_gdpnolimit
keep year summ_*
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_mcare_ptb_enroll summ_age67nolimit
keep year summ_*

twoway connected summ_ACA summ_drop03 summ_gdp summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("People Enrolled in Part B (Millions)")
outsheet year summ_* using enroll_b_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{enroll_b_summary}
\caption{Annual Enrollment in Medicare Part B}
\end{figure}

\begin{Statacode}{!echo,hide,savefig,label=uncovered_upper}
use ../output/ACA/summary, clear
rename ttl_uncovered_upper_mcare summ_aca
label variable summ_aca "ACA"
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03/summary, sort
rename ttl_uncovered_upper_mcare summ_parta
label variable summ_parta "Part A Drop 3 ACA Thresholds"
keep year summ_* unc_6566
merge year using ../output/voucher.gdp/summary, sort
rename ttl_uncovered_upper_mcare summ_voucher
label variable summ_voucher "Voucher GDP"
keep year summ_* unc_6566
merge year using ../output/elig.age67/summary, sort
rename ttl_uncovered_upper_mcare summ_age67
replace summ_age67 = summ_age67 + unc_6566 if year > 2010
label variable summ_age67 "Eligibility Age 67"
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_ACAnolimit
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_partAnolimit
keep year summ_* unc_6566
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_gdpnolimit
keep year summ_* unc_6566
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_uncovered_upper_mcare summ_age67nolimit
replace summ_age67nolimit = summ_age67nolimit + unc_6566 if year > 2010
keep year summ_* unc_6566

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_aca-summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("NPV of OOP Expenditures (Trillions)")
outsheet year summ_* using uncovered_upper.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{uncovered_upper}
\caption{Upper Bound on Total Uncovered Expenditures for Traditional Medicare Eligible Population}
\end{figure}

\begin{Statacode}{!echo,hide,savefig,label=uncovered_lower}
use ../output/ACA/summary, clear
rename ttl_uncovered_lower_mcare summ_aca
label variable summ_aca "ACA"
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03/summary, sort
rename ttl_uncovered_lower_mcare summ_parta
label variable summ_parta "Part A Drop 3 ACA Thresholds"
keep year summ_* unc_6566
merge year using ../output/voucher.gdp/summary, sort
rename ttl_uncovered_lower_mcare summ_voucher
label variable summ_voucher "Voucher GDP"
keep year summ_* unc_6566
merge year using ../output/elig.age67/summary, sort
rename ttl_uncovered_lower_mcare summ_age67
replace summ_age67 = summ_age67 + 18.0/41 * unc_6566 if year > 2010
label variable summ_age67 "Eligibility Age 67"
keep year summ_*

merge year using ../output/ACAnolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_ACAnolimit
rename ttl_totmd_6566 unc_6566
keep year summ_* unc_6566
merge year using ../output/partA.drop03nolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_partAnolimit
keep year summ_* unc_6566
merge year using ../output/voucher.gdpnolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_gdpnolimit
keep year summ_* unc_6566
merge year using ../output/elig.age67nolimit/summary, sort
rename ttl_uncovered_lower_mcare summ_age67nolimit
replace summ_age67nolimit = summ_age67nolimit + unc_6566 if year > 2010
keep year summ_* unc_6566

tsset year
tsfill
foreach v of varlist summ_* {
  ipolate `v' year if year >= 2012, gen(i`v')
  replace `v' = i`v'
}
foreach v of varlist summ_* {
  replace `v' = `v' * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
}
* $ to fix syntax highlighting

twoway connected summ_aca-summ_age67 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("NPV of OOP Expenditures (Trillions)")
outsheet year summ_* using uncovered_lower.csv, comma replace
\end{Statacode}

\begin{figure}
\centering
\recallfig{uncovered_lower}
\caption{Lower Bound on Total Uncovered Expenditures for Traditional Medicare Eligible Population}
\end{figure}

\begin{table}[ht]
\centering
\caption{Scenario Definitions}\vspace{0.2in}
\begin{tabular}{|l|c|c|c|c|c|c|c|}\hline
 & \textbf{Medicare} & \textbf{Part A} & \textbf{Part A} & \textbf{Part A} & \textbf{Part B} & \textbf{Part B} & \textbf{Medical Cost} \\
\textbf{Scenario} & \textbf{Eligibility} & \textbf{Premium} & \textbf{Drop} & \textbf{Premium} & \textbf{Premium} & \textbf{Premium} & \textbf{Growth} \\
& \textbf{Age} & \textbf{Schedule} & \textbf{Fraction} & \textbf{Elasticity} & \textbf{Elasticity} & \textbf{Thresholds} & \textbf{Rate}\\\hline
ACA & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
Zero Elasticity & 65 & 0\% & 0\% & 0.0 & 0.0 & ACA & SSA Intermediate\\\hline
-0.2 Elasticity & 65 & 0\% & 0\% & 0.0 & -0.2 & ACA & SSA Intermediate\\\hline
-0.5 Elasticity & 65 & 0\% & 0\% & 0.0 & -0.5 & ACA & SSA Intermediate\\\hline
Part B Nominal Dollars & 65 & 0\% & 0\% & 0.0 & -0.1 & Nominal & SSA Intermediate\\\hline
Part B Real Dollars & 65 & 0\% & 0\% & 0.0 & -0.1 & 2010 & SSA Intermediate\\\hline
Part B 25 Percent & 65 & 0\% & 0\% & 0.0 & -0.1 & N/A & SSA Intermediate\\\hline
Part A Drop 3 & 65 & ACA & 3\% & -0.1 & -0.1 & ACA & SSA Intermediate\\\hline
Vouchers GDP & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & OASDI GDP \\\hline
Vouchers GDP+1 & 65 & 0\% & 0\% & 0.0 & -0.1 & ACA & OASDI GPD + 1\%\\\hline
Eligibility Age 67 & 67 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
Eligibility Age 70 & 70 & 0\% & 0\% & 0.0 & -0.1 & ACA & SSA Intermediate\\\hline
\end{tabular}
\label{tab:scenarios}
\end{table}

\section{Imposing a Premium on Part A}
Starting in 2010, premiums are imposed for Part A according to Table~\ref{tab:sliding}. The thresholds do not grow with inflation, so in real-dollar terms they decline over time. Any household under 138\% of the poverty line does not have any premium. Figure~\ref{fig:exemptions} shows the evolution of this population over time.

\begin{table}[ht]
\caption{Individual Sliding Scale Premiums as percentage of per-enrollee program cost}
\label{tab:sliding}
\centering
\begin{tabular}{|c|c|c|c|}\hline
\textbf{Band} & \textbf{Income Bracket} & \textbf{Part A Premium} & \textbf{Part B Premium}\\\hline
Band 1 & MAGI <= \$85,000 & 25\% & 25\%\\\hline
Band 2 & \$85,000 < MAGI <= \$107,000 & 35\% & 35\%\\\hline
Band 3 & \$107,000 < MAGI <= \$160,000  & 55\% & 55\%\\\hline
Band 4 & \$160,000 < MAGI <= \$214,000 & 65\% & 65\%\\\hline
Band 5 & \$214,000 < MAGI & 80\% & 80\%\\\hline
\end{tabular}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partB.real.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=income_pops_real}
gen b0 = income_band0/pop_medicare_age
gen b1 = income_band1/pop_medicare_age
gen b2 = income_band2/pop_medicare_age
gen b3 = income_band3/pop_medicare_age
gen b4 = income_band4/pop_medicare_age
gen b5 = income_band5/pop_medicare_age
gen pool = pop_medicare_age
global vlist b0 b1 b2 b3 b4 b5 pool
format $vlist %5.2f
outsheet year $vlist if year == 2010 | year == 2020 | year == 2036 using income_pops_real.csv, comma replace
\end{Statacode}
\begin{Statacode}{!echo,hide}
use ../output/partB.nominal.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=income_pops_nominal}
gen b0 = income_band0/pop_medicare_age
gen b1 = income_band1/pop_medicare_age
gen b2 = income_band2/pop_medicare_age
gen b3 = income_band3/pop_medicare_age
gen b4 = income_band4/pop_medicare_age
gen b5 = income_band5/pop_medicare_age
gen pool = pop_medicare_age
global vlist b0 b1 b2 b3 b4 b5 pool
format $vlist %5.2f
outsheet year $vlist if year == 2010 | year == 2020 | year == 2036 using income_pops_nominal.csv, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,Medicaid,Band 1,Band 2,Band 3,Band 4,Band 5,Total 65+}]{income_pops_real}{income_pops_real.csv}
\DTLloaddb[headers={Year,Medicaid,Band 1,Band 2,Band 3,Band 4,Band 5,Total 65+}]{income_pops_nominal}{income_pops_nominal.csv}

\begin{table}[ht]
\centering
\caption{Percentage in Real Dollar Threshold Bands by Year (Millions of People)}
\DTLdisplaydb{income_pops_real}
\label{tab:income_pops_real}
\end{table}

\begin{table}[ht]
\centering
\caption{Percentage in Nominal Dollar Threshold Bands by Year (Millions of People)}
\DTLdisplaydb{income_pops_nominal}
\label{tab:income_pops_nominal}
\end{table}

We looked at four different scenarios all based on Buchmueller, 2006. When premium jumps from nothing to the sliding-scale premium, there is a small chance that the person will drop (or not enroll in the first place). Once enrolled, all further premium adjustments are subject to a -0.1 elasticity to price and a takeup probit based on the model for Part B takeup. The average premium paid does not change significantly from scenario to scenario.

\begin{Statacode}{!echo,hide,fig,savefig,label=exemptions}
use ../output/ACA/summary, clear
gen bpop = ttl_mcare_ptb_exempt / pop_medicare
label variable bpop "Portion of Medicare Eligible Exempt from Part B Premium"
gen benroll = ttl_mcare_ptb_exempt / ttl_mcare_ptb_enroll
label variable benroll "Portion of Part B Enrolled Exempt from Premium under ACA"
keep year bpop benroll
sort year
merge year using ../output/partA.drop03/summary
gen aenroll = ttl_mcare_pta_exempt / ttl_mcare_pta_enroll
label variable aenroll "Portion of Part A Enrolled Exempt from Premium under Drop 3"
twoway connected bpop aenroll benroll year if year >= 2010, legend(cols(1))
outsheet year bpop aenroll benroll using exemptions.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{exemptions}
\caption{Fraction of Medicare Enrollees Exempt from Premium in ACA}
\label{fig:exemptions}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=drop_enrolled}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll base
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
rename ttl_mcare_pta_enroll drop03
label variable drop03 "Part A Drop 3 ACA Thresholds"
keep year base drop03
sort year
merge year using ../output/partA.drop03.real.dollars/summary
rename ttl_mcare_pta_enroll drop03Real
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
rename ttl_mcare_pta_enroll drop0325
label variable drop0325 "Part A Drop 3 25 Percent"
keep year base drop03 drop03Real drop0325
sort year
merge year using ../output/partA.drop03.10percent/summary
rename ttl_mcare_pta_enroll drop0310
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Number of Enrolled (Millions)")
outsheet year base drop03 drop03Real drop0310 drop0325 using drop_enrolled.csv if year >= 2010, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,ACA,Drop 03, Drop 03 Real,Drop 03 10 Percent,Drop 03 25 Percent}]{parta_enrolled}{drop_enrolled.csv}

\begin{table}[ht]
\centering
\caption{Enrollment in Part A}
\DTLdisplaydb{parta_enrolled}
\label{tab:parta_c_cost}
\end{table}

\begin{figure}[ht]
\centering
\recallfig{drop_enrolled}
\caption{Enrollment in Part A}
\label{fig:drop_enrolled}
\end{figure}

Part A costs (Figure~\ref{fig:drop_costs}) are relative to the enrollment, but the net costs are roughly the same, as shown in Figure~\ref{fig:parta_net_costs}.
\begin{Statacode}{!echo,hide,fig,savefig,label=drop_costs}
use ../output/ACA/summary, clear
gen base = ttl_mcare_pta
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
gen drop03 = ttl_mcare_pta
label variable drop03 "Part A Drop 3 ACA Thresholds"
keep year base drop03
sort year
merge year using ../output/partA.drop03.real.dollars/summary
rename ttl_mcare_pta drop03Real
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
rename ttl_mcare_pta drop0325
label variable drop0325 "Part A Drop 3 25 Percent"
keep year base drop03 drop03Real drop0325
sort year
merge year using ../output/partA.drop03.10percent/summary
rename ttl_mcare_pta drop0310
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base drop03 drop03Real drop0310 drop0325 using drop_costs.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{drop_costs}
\caption{Part A Total Costs with varying drop rates (excluding payroll tax revenue)}
\label{fig:drop_costs}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=parta_net_costs}
use ../output/ACA/summary, clear
gen base = ttl_mcare_pta
label variable base "ACA"
keep year base
sort year
merge year using ../output/partA.drop03/summary
gen drop03 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop03 "Part A Drop 3 ACA Thresholds"
sort year
keep year base drop03
merge year using ../output/partA.drop03.real.dollars/summary
gen drop03Real = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop03Real "Part A Drop 3 Real Dollars"
keep year base drop03 drop03Real
sort year
merge year using ../output/partA.drop03.25percent/summary
gen drop0325 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop0325 "Part A Drop 3 25 Percent"
sort year
keep year base drop03 drop03Real drop0325
merge year using ../output/partA.drop03.10percent/summary
gen drop0310 = ttl_mcare_pta - ttl_mcare_pta_premium
label variable drop0310 "Part A Drop 3 10 Percent"
twoway connected base drop03 drop03Real drop0310 drop0325 year if year >= 2010, legend(cols(1)) yscale(titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base drop03 drop03Real drop0310 drop0325 using parta_net_costs.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{parta_net_costs}
\caption{Part A Net Costs with varying drop rates (excluding payroll tax revenue)}
\label{fig:parta_net_costs}
\end{figure}

\begin{Statacode}{!echo,hide}
tsset year
tsfill
foreach v in base drop03 {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace drop03 = drop03 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
gen c_base = sum(base) if year >= 2010
gen c_drop03 = sum(drop03) if year >= 2010
gen diffDrop03 = c_drop03 - c_base
format c_base diffDrop03 %10.2f
outsheet year c_base diffDrop03 if year >= 2010 using parta_c_cost.csv, replace comma
\end{Statacode}
\DTLloaddb[headers={Year,ACA,Drop 03}]{parta_c_cost}{parta_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Part A Options Less Net Present Value of ACA Part A (Billions 2010 Dollars)}
\DTLdisplaydb{parta_c_cost}
\label{tab:parta_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partA.drop03/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,saveout,label=parta_premium}
gen basePremium = 0.25 * mcare_pta
gen ptaPremium = mcare_pta_premium_paid
format basePremium ptaPremium %10.2f
outsheet year basePremium ptaPremium if year >= 2010 using parta_premium.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,Base Premium,Average Premium}]{parta_premium}{parta_premium.csv}

\begin{table}[ht]
\centering
\caption{Part A Drop 3 Nominal and Average Part A Premiums}
\DTLdisplaydb{parta_premium}
\label{tab:parta_premium}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
Here is the code for calculating the deviation:
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/partA.drop03/summary
gen drop_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen drop_a_mean = mcare_pta - mcare_pta_premium_paid
gen drop_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen drop_d_mean = mcare_ptd
gen drop_a_enrl = ttl_mcare_pta_enroll * 1e6
gen drop_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen drop_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* drop_* pop_medicare {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen baseline = base_ttl / 1e9
gen dropCost = drop_ttl / 1e9
gen enrollment_deviation = (base_ttl_a * (drop_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (drop_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (drop_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_deviation = (base_a_enrl * (drop_a_mean - base_a_mean) + base_b_enrl * (drop_b_mean - base_b_mean) + base_d_enrl * (drop_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace dropCost = dropCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_deviation = enrollment_deviation * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_deviation = outlay_deviation * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $$ syntax highlighting fix
gen cBaseline = sum(baseline) if year >= 2010
gen cDropCost = sum(dropCost) if year >= 2010
gen enrollment = sum(enrollment_deviation) if year >= 2010
gen outlay = sum(outlay_deviation) if year >= 2010
format cBaseline enrollment outlay %10.2f
outsheet year cBaseline cDropCost enrollment outlay using parta_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,Real Dollars Cost,Drop 03,Enrollment Deviation,Enrollee Outlay Deviation}]{parta_deviation}{parta_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs for Drop 03 from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{parta_deviation}
\label{tab:parta_deviation}
\end{table}

\section{Impact of Sliding Scale on Part B}

Since Medicare Part B already has a premium, we don't need to make any assumptions about people dropping when a change is introduced, we can just use the elasticity directly. The current ACA plan is to keep the premium thresholds in nominal dollars. We also ran two scenarios with thresholds in real dollars (2010) and with no thresholds where everyone not exempt pays the same premium (25\%).

Figure~\ref{fig:partB_enrollment} shows the change in enrollment fraction and Figure~\ref{fig:partB_cost} shows the change in net costs. Since the 25\% premium is a continuation of previous policy, it is not surprising that enrollment continues relatively flat. The initial large discrepancy between real and nominal dollar thresholds is due to higher average premiums under the nominal dollar scheme as shown in Figure~\ref{fig:partB_premium}. The convergence is due to the lowest threshold falling below the exemption threshold.

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_enrollment}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb_enroll
keep year base
label variable base "ACA"
merge year using ../output/partB.25percent/summary
gen p25 = ttl_mcare_ptb_enroll
label variable p25 "Part B 25 Percent"
keep year base p25
sort year
merge year using ../output/partB.real.dollars/summary
gen real = ttl_mcare_ptb_enroll
label variable real "Part B Real Dollars"
keep year base p25 real
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = ttl_mcare_ptb_enroll
label variable nominal "Part B Nominal Dollars"
twoway connected base nominal real p25 year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Number of Persons Enrolled (Millions)")
outsheet year base nominal real using partB_enrollment.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_cost}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable base "ACA"
keep year base
merge year using ../output/partB.25percent/summary
gen p25 = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable p25 "Part B 25 Percent"
keep year base p25
sort year
merge year using ../output/partB.real.dollars/summary
gen real = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable real "Part B Real Dollars"
keep year base p25 real
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = ttl_mcare_ptb - ttl_mcare_ptb_premium
label variable nominal "Part B Nominal Dollars"
twoway connected base nominal real p25 year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Billions of 2010 Dollars")
outsheet year base nominal real p25 using partB_cost.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{partB_enrollment}
\caption{Medicare Part B Enrollment}
\label{fig:partB_enrollment}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{partB_cost}
\caption{Medicare Part B Net Cost (excluding payroll tax revenue)}
\label{fig:partB_cost}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=partb_c_cost}
tsset year
tsfill
foreach v in base p25 real nominal {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace p25 = p25 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace nominal = nominal * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace real = real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
gen c_base = sum(base)  if year >= 2010
gen c_p25 = sum(p25) if year >= 2010
gen c_real = sum(real) if year >= 2010
gen c_nominal = sum(nominal) if year >= 2010
gen diffP25 = c_p25 - c_base
gen diffReal = c_real - c_base
gen diffNominal = c_nominal - c_base
global vlist diffP25 diffReal diffNominal
format c_base $vlist %10.2f
outsheet year c_base $vlist if year >= 2010 using partb_c_cost.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,ACA,25\% Premium,Real Dollar Thresholds,Nominal Dollar Thresholds}]{partb_c_cost}{partb_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Part B Options Less Net Present Value of ACA Part B (Billions 2010 Dollars)}
\DTLdisplaydb{partb_c_cost}
\label{tab:partb_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/partB.25percent/summary, clear
twoway connected mcare_ptb_enroll5564 mcare_ptb_enroll6569 mcare_ptb_enroll7074 mcare_ptb_enroll7579 mcare_ptb_enroll80pp year if year >= 2010, yscale(titlegap(*2)) ytitle("Fraction of Eligible Enrolled") legend(cols(1))
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=premium_summary}
use ../output/ACA/summary, clear
gen baseB = mcare_ptb_premium_paid
label variable baseB "ACA"
keep year baseB
sort year
merge year using ../output/partB.real.dollars/summary
gen real = mcare_ptb_premium_paid
label variable real "Part B Real Dollars"
keep year baseB real
sort year
merge year using ../output/partB.25percent/summary
gen partB25 = mcare_ptb_premium_paid
label variable partB25 "Part B 25 Percent"
keep year baseB real partB25
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nominal = mcare_ptb_premium_paid
label variable nominal "Part B Nominal Dollars"
twoway connected baseB real nominal partB25 year if year >= 2010, legend(cols(1)) ytitle("Yearly Premium in 2010 Dollars")
outsheet year baseB real nominal partB25 using  premium_summary.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{premium_summary}
\caption{Average Part B Premium Paid}
\label{fig:partB_premium}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=partB_elasticity}
use ../output/ACA/summary, clear
gen base = ttl_mcare_ptb_enroll
label variable base "ACA"
keep base year
sort year
merge year using ../output/mcare.0elas/summary
gen zero = ttl_mcare_ptb_enroll
label variable zero "Zero Elasticity"
keep base zero year
sort year
merge year using ../output/mcare.neg2elas/summary
gen neg2 = ttl_mcare_ptb_enroll
label variable neg2 "-0.2 Elasticity"
keep base zero neg2 year
sort year
merge year using ../output/mcare.neg5elas/summary
gen neg5 = ttl_mcare_ptb_enroll
label variable neg5 "-0.5 Elasticity"
twoway connected zero neg2 base neg5 year if year >= 2010, legend(cols(1)) ytitle("Number of Enrolled (Millions)")
outsheet year zero neg2 base neg5 using partB_elasticity.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{partB_elasticity}
\caption{Effect of Elasticity on Part B Enrollment}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/partB.real.dollars/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen real_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen real_a_mean = mcare_pta - mcare_pta_premium_paid
gen real_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen real_d_mean = mcare_ptd
gen real_a_enrl = ttl_mcare_pta_enroll * 1e6
gen real_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen real_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year real_*
sort year
merge year using ../output/ACA/summary
gen aca_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen aca_a_mean = mcare_pta - mcare_pta_premium_paid
gen aca_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen aca_d_mean = mcare_ptd
gen aca_a_enrl = ttl_mcare_pta_enroll * 1e6
gen aca_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen aca_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen aca_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen aca_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen aca_ttl_d = ttl_mcare_ptd * 1e9
keep year real_* aca_*
sort year
merge year using ../output/partB.nominal.dollars/summary
gen nomin_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen nomin_a_mean = mcare_pta - mcare_pta_premium_paid
gen nomin_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen nomin_d_mean = mcare_ptd
gen nomin_a_enrl = ttl_mcare_pta_enroll * 1e6
gen nomin_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen nomin_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist real_* aca_* nomin_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen realCost = real_ttl / 1e9
gen acaCost = aca_ttl / 1e9
gen nominCost = nomin_ttl / 1e9
gen enrollment_real = (aca_ttl_a * (real_a_enrl - aca_a_enrl)/aca_a_enrl + aca_ttl_b * (real_b_enrl - aca_b_enrl)/aca_b_enrl + aca_ttl_d * (real_d_enrl - aca_d_enrl)/aca_d_enrl) / 1e9
gen outlay_real = (aca_a_enrl * (real_a_mean - aca_a_mean) + aca_b_enrl * (real_b_mean - aca_b_mean) + aca_d_enrl * (real_d_mean - aca_d_mean)) / 1e9
gen enrollment_nomin = (aca_ttl_a * (nomin_a_enrl - aca_a_enrl)/aca_a_enrl + aca_ttl_b * (nomin_b_enrl - aca_b_enrl)/aca_b_enrl + aca_ttl_d * (nomin_d_enrl - aca_d_enrl)/aca_d_enrl) / 1e9
gen outlay_nomin = (aca_a_enrl * (nomin_a_mean - aca_a_mean) + aca_b_enrl * (nomin_b_mean - aca_b_mean) + aca_d_enrl * (nomin_d_mean - aca_d_mean)) / 1e9
replace realCost = realCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace acaCost = acaCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace nominCost = nominCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_real = enrollment_real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_real = outlay_real * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_nomin = enrollment_nomin * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_nomin = outlay_nomin * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cReal = sum(realCost) if year >= 2010
gen cACA = sum(acaCost) if year >= 2010
gen cNomin = sum(nominCost) if year >= 2010
gen cummEnrollmentReal = sum(enrollment_real) if year >= 2010
gen cummOutlayReal = sum(outlay_real) if year >= 2010
gen cummEnrollmentNominal = sum(enrollment_nomin) if year >= 2010
gen cummOutlayNominal = sum(outlay_nomin) if year >= 2010
global vlist cACA  cReal cummEnrollmentReal cummOutlayReal cNomin cummEnrollmentNominal cummOutlayNominal 
format $vlist %10.2f
outsheet year $vlist using partb_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{Real\\Dollars\\Cost},\shortstack{Real\\Dollars\\Enrollment\\Deviation}, \shortstack{Real\\Dollars\\Enrollee\\Outlay\\Deviation}, \shortstack{Nominal\\Dollars\\Cost},\shortstack{Nominal\\Dollars\\Enrollment\\Deviation}, \shortstack{Nominal\\Dollars\\Enrollee\\Outlay\\Deviation}}]{partb_deviation}{partb_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{partb_deviation}
\label{tab:partb_deviation}
\end{table}

\section{Medicare Vouchers can Save Money}

The analysis here assumes that the government will issue vouchers whose value will grow at a rate different from the Medicare cost growth rate. The difference between the cost of vouchers and the projected cost of Medicare would be the savings (or loss) to the government. In the base scenario, there are no vouchers and Medicare costs grow as expected. The first voucher scenario grows at the rate of GDP (normalized by labor-force growth). The second scenario grows at GDP plus 1\%. Results are shown in Figure~\ref{fig:vouchers}.

The GDP plus 1\% scenario is more expensive in the out years because the Medicare cost growth is projected to fall below GDP+1\% by 2034 and to stay below until the end of the simulation period. In the near term, either voucher scenario will save money since Medicare costs are rising at about 1.5\% above GDP.

\begin{Statacode}{!echo,hide,fig,savefig,label=voucher_a_enroll}
use ../output/ACA/summary, clear
rename ttl_mcare_pta_enroll base
label variable base "ACA Part A Enrollment"
keep year base
merge year using ../output/voucher.gdp/summary
rename ttl_mcare_pta_enroll gdp
label variable gdp "GDP Part A Enrollment"
sort year
keep year base gdp
merge year using ../output/voucher.gdp.plus1/summary
rename ttl_mcare_pta_enroll plus1
label variable plus1 "GDP Plus 1 Part A Enrollment"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_pta_enroll cpi
label variable cpi "CPI Part A Enrollment"
twoway connected base gdp cpi year if year >= 2010, yscale(titlegap(*2)) ytitle("Number Enrolled (Millions) in Part A") legend(cols(1))
outsheet year base gdp plus1 cpi using voucher_a_enroll.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=voucher_b_enroll}
use ../output/ACA/summary, clear
rename ttl_mcare_ptb_enroll base
label variable base "ACA Part B Enrollment"
keep year base
merge year using ../output/voucher.gdp/summary
rename ttl_mcare_ptb_enroll gdp
label variable gdp "GDP Part B Enrollment"
sort year
keep year base gdp
merge year using ../output/voucher.gdp.plus1/summary
rename ttl_mcare_ptb_enroll plus1
label variable plus1 "GDP Plus 1 Part B Enrollment"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
rename ttl_mcare_ptb_enroll cpi
label variable cpi "CPI Part B Enrollment"
twoway connected base gdp cpi year if year >= 2010, yscale(titlegap(*2)) ytitle("Number Enrolled (Millions) in Part B") legend(cols(1))
outsheet year base gdp plus1 cpi using voucher_b_enroll.csv, comma replace
\end{Statacode}

\begin{Statacode}{!echo,hide,fig,savefig,label=vouchers}
use ../output/ACA/summary, clear
gen base_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
keep year base_medicare
label variable base_medicare "ACA"
merge year using ../output/voucher.gdp/summary
sort year
gen gdp_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace gdp_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable gdp_medicare "Vouchers GDP"
keep year base_medicare gdp_medicare
merge year using ../output/voucher.gdp.plus1/summary
gen plus1_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace plus1_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable plus1_medicare "Vouchers GDP+1"
keep year base gdp plus1
merge year using ../output/voucher.cpi/summary, sort
gen cpi_medicare = ttl_mcare_pta_subsidy + ttl_mcare_ptb_subsidy + ttl_mcare_ptd
replace cpi_medicare = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium if year < 2012
label variable cpi_medicare "Vouchers CPI"
twoway connected base_medicare gdp_medicare cpi_medicare year if year >= 2010, legend(cols(1)) yscale(axis(1) titlegap(*2)) ytitle("Total Medicare Cost (Billions of 2010 Dollars)")
outsheet year base_medicare gdp_medicare plus1_medicare cpi_medicare using vouchers.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{vouchers}
\caption{Comparison of Medicare Cost Growth with Vouchers - Long Term}
\label{fig:vouchers}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{voucher_a_enroll}
\caption{Number Enrolled in Part A Under Voucher Scenarios}
\label{fig:voucher_enroll}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{voucher_b_enroll}
\caption{Number Enrolled in Part B Under Voucher Scenarios}
\label{fig:voucher_enroll}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=voucher_c_cost}
tsset year
tsfill
foreach v in base_medicare gdp_medicare plus1_medicare {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base_medicare = base_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace gdp_medicare = gdp_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace plus1_medicare = plus1_medicare * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ for syntax highlights
gen c_base = sum(base_medicare) if year >= 2010
gen c_gdp = sum(gdp_medicare) if year >= 2010
gen c_plus1 = sum(plus1_medicare) if year >= 2010
gen diffGDP = c_gdp - c_base
gen diffPlus1 = c_plus1 - c_base
format c_base diffGDP diffPlus1 %10.2f
outsheet year c_base diffGDP diffPlus1 if year >= 2010 using voucher_c_cost.csv, comma replace
\end{Statacode}
\DTLloaddb[headers={Year, ACA, GDP, GDP Plus 1\%}]{voucher_c_cost}{voucher_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Voucher Programs Less Net Present Value of ACA (Billions 2010 Dollars)}
\DTLdisplaydb{voucher_c_cost}
\label{tab:voucher_c_cost}
\end{table}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/voucher.gdp/summary
gen gdp_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen gdp_a_mean = mcare_pta - mcare_pta_premium_paid
gen gdp_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen gdp_d_mean = mcare_ptd
gen gdp_a_enrl = ttl_mcare_pta_enroll * 1e6
gen gdp_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen gdp_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year base_* gdp_*
sort year
merge year using ../output/voucher.gdp.plus1/summary
gen plus_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen plus_a_mean = mcare_pta - mcare_pta_premium_paid
gen plus_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen plus_d_mean = mcare_ptd
gen plus_a_enrl = ttl_mcare_pta_enroll * 1e6
gen plus_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen plus_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* gdp_* plus_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
gen baseline = base_ttl / 1e9
gen gdpCost = gdp_ttl / 1e9
gen plusCost = plus_ttl / 1e9
gen enrollment_gdp = (base_ttl_a * (gdp_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (gdp_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (gdp_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_gdp = (base_a_enrl * (gdp_a_mean - base_a_mean) + base_b_enrl * (gdp_b_mean - base_b_mean) + base_d_enrl * (gdp_d_mean - base_d_mean)) / 1e9
gen enrollment_plus = (base_ttl_a * (plus_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (plus_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (plus_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_plus = (base_a_enrl * (plus_a_mean - base_a_mean) + base_b_enrl * (plus_b_mean - base_b_mean) + base_d_enrl * (plus_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace gdpCost = gdpCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace plusCost = plusCost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_gdp = enrollment_gdp * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_gdp = outlay_gdp * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_plus = enrollment_plus * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_plus = outlay_plus * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cBaseline = sum(baseline) if year >= 2010
gen cGDP = sum(gdpCost) if year >= 2010
gen cPlus = sum(plusCost) if year >= 2010
gen cummEnrollmentGDP = sum(enrollment_gdp) if year >= 2010
gen cummOutlayGDP = sum(outlay_gdp) if year >= 2010
gen cummEnrollmentPlusal = sum(enrollment_plus) if year >= 2010
gen cummOutlayPlusal = sum(outlay_plus) if year >= 2010
global vlist cBaseline cGDP cummEnrollmentGDP cummOutlayGDP cPlus cummEnrollmentPlusal cummOutlayPlusal 
format $vlist %10.2f
outsheet year $vlist using voucher_deviation.csv if year >= 2010, comma replace
\end{Statacode}
\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{GDP\\Cost},\shortstack{GDP\\Enrollment\\Deviation}, \shortstack{GDP\\Enrollee\\Outlay\\Deviation}, \shortstack{Plus 1\\Cost},\shortstack{Plus 1\\Enrollment\\Deviation}, \shortstack{Plus 1\\Enrollee\\Outlay\\Deviation}}]{voucher_deviation}{voucher_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Voucher Scenario Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{voucher_deviation}
\label{tab:voucher_deviation}
\end{table}

\section{Increasing Eligibility Age Reduces Total Cost and Enrollment}

Figure~\ref{fig:age_enroll} shows the number of people eligible for Medicare under three different age-based criteria.
Figure~\ref{fig:age_cost} shows that this decline in enrollment produces the expected decline in total cost, but Figure~\ref{fig:age_average} shows that the average cost per enrollee is higher, due to the elderly requiring more medical care per person.

\begin{Statacode}{!echo,hide,fig,savefig,label=age_enroll}
use ../output/ACA/summary, clear
rename pop_medicare base
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
rename pop_medicare age67
label variable age67 "Eligibility Age 67"
sort year
keep year base age67
merge year using ../output/elig.age70/summary
rename pop_medicare age70
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Millions Eligible for Medicare") legend(cols(1))
outsheet year base age67 age70 using age_enroll.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_enroll}
\caption{Medicare-eligigble population after raising eligibility age}
\label{fig:age_enroll}
\end{figure}

\begin{Statacode}{!echo,hide,fig,savefig,label=age_cost}
use ../output/ACA/summary, clear
gen base = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
gen age67 = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable age67 "Eligibility Age 67"
keep base age67 year
sort year
merge year using ../output/elig.age70/summary
gen age70 = ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Total Medicare Cost (Billions of 2010 Dollars)") legend(cols(1))
outsheet year base age67 age70 using age_cost.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_cost}
\caption{Total Medicare Costs Fall when Eligibility Rises}
\label{fig:age_cost}
\end{figure}

\begin{Statacode}{!echo,hide,saveout,label=age_c_cost}
tsset year
tsfill
foreach v in base age67 age70 {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
replace base = base * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace age67 = age67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace age70 = age70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen c_base = sum(base) if year >= 2010
gen c_age67 = sum(age67) if year >= 2010
gen c_age70 = sum(age70) if year >= 2010
gen diffAge67 = c_age67 - c_base
gen diffAge70 = c_age70 - c_base
global vlist c_base diffAge67 diffAge70
format $vlist %10.2f
outsheet year $vlist if year >= 2010 using age_c_cost.csv, replace comma
\end{Statacode}
\DTLloaddb[headers={Year, ACA, Eligibility Age 67, Eligibility Age 70}]{age_c_cost}{age_c_cost.csv}

\begin{table}[ht]
\centering
\caption{Net Present Value of Eligibility Age Options Less Net Present Value of ACA (Billions of 2010 Dollars)}
\DTLdisplaydb{age_c_cost}
\label{tab:age_c_cost}
\end{table}

\begin{Statacode}{!echo,hide,fig,savefig,label=age_average}
use ../output/ACA/summary, clear
gen base = ttl_mcare/pop_medicare
label variable base "ACA"
keep year base
merge year using ../output/elig.age67/summary
gen age67 = ttl_mcare/pop_medicare
label variable age67 "Eligibility Age 67"
keep base age67 year
sort year
merge year using ../output/elig.age70/summary
gen age70 = ttl_mcare/pop_medicare
label variable age70 "Eligibility Age 70"
twoway connected base age67 age70 year if year >= 2010, yscale(titlegap(*2)) ytitle("Medicare Cost per Enrollee (2010 Dollars)") legend(cols(1))
outsheet year base age67 age70 using age_average.csv, comma replace
\end{Statacode}

\begin{figure}[ht]
\centering
\recallfig{age_average}
\caption{Average Medicare Costs Rise when Eligibility Age Rises}
\label{fig:age_average}
\end{figure}

\begin{Statacode}{!echo,hide}
use ../output/ACA/summary, clear
\end{Statacode}
\begin{Statacode}{!echo,hide}
gen base_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen base_a_mean = mcare_pta - mcare_pta_premium_paid
gen base_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen base_d_mean = mcare_ptd
gen base_a_enrl = ttl_mcare_pta_enroll * 1e6
gen base_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen base_d_enrl = ttl_mcare_ptd_enroll * 1e6
gen base_ttl_a = ttl_mcare_pta * 1e9 - ttl_mcare_pta_premium * 1e9
gen base_ttl_b = ttl_mcare_ptb * 1e9 - ttl_mcare_ptb_premium * 1e9
gen base_ttl_d = ttl_mcare_ptd * 1e9
keep year base_*
sort year
merge year using ../output/elig.age67/summary
gen a67_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen a67_a_mean = mcare_pta - mcare_pta_premium_paid
gen a67_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen a67_d_mean = mcare_ptd
gen a67_a_enrl = ttl_mcare_pta_enroll * 1e6
gen a67_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen a67_d_enrl = ttl_mcare_ptd_enroll * 1e6
keep year base_* a67_*
sort year
merge year using ../output/elig.age70/summary
gen a70_ttl = (ttl_mcare - ttl_mcare_pta_premium - ttl_mcare_ptb_premium) * 1e9
gen a70_a_mean = mcare_pta - mcare_pta_premium_paid
gen a70_b_mean = mcare_ptb - mcare_ptb_premium_paid
gen a70_d_mean = mcare_ptd
gen a70_a_enrl = ttl_mcare_pta_enroll * 1e6
gen a70_b_enrl = ttl_mcare_ptb_enroll * 1e6
gen a70_d_enrl = ttl_mcare_ptd_enroll * 1e6
tsset year
tsfill
foreach v of varlist base_* a67_* a70_* {
  ipolate `v' year if year >= 2010, gen(i`v')
  replace `v' = i`v'
}
foreach v in "a_enrl" "b_enrl" "d_enrl" "a_mean" "b_mean" "d_mean" {
  replace a67_`v' = base_`v' if year < 2014
  replace a70_`v' = base_`v' if year < 2014
} 
gen baseline = base_ttl / 1e9
gen a67Cost = a67_ttl / 1e9
gen a70Cost = a70_ttl / 1e9
gen enrollment_a67 = (base_ttl_a * (a67_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (a67_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (a67_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_a67 = (base_a_enrl * (a67_a_mean - base_a_mean) + base_b_enrl * (a67_b_mean - base_b_mean) + base_d_enrl * (a67_d_mean - base_d_mean)) / 1e9
gen enrollment_a70 = (base_ttl_a * (a70_a_enrl - base_a_enrl)/base_a_enrl + base_ttl_b * (a70_b_enrl - base_b_enrl)/base_b_enrl + base_ttl_d * (a70_d_enrl - base_d_enrl)/base_d_enrl) / 1e9
gen outlay_a70 = (base_a_enrl * (a70_a_mean - base_a_mean) + base_b_enrl * (a70_b_mean - base_b_mean) + base_d_enrl * (a70_d_mean - base_d_mean)) / 1e9
replace baseline = baseline * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace a67Cost = a67Cost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace a70Cost = a70Cost * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_a67 = enrollment_a67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_a67 = outlay_a67 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace enrollment_a70 = enrollment_a70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
replace outlay_a70 = outlay_a70 * 1/(1 + $discount_rate)^(year - 2010) if year > 2010
* $ fix syntax highlighting
gen cBaseline = sum(baseline) if year >= 2010
gen cA67 = sum(a67Cost) if year >= 2010
gen cA70 = sum(a70Cost) if year >= 2010
gen cummEnrollmentA67 = sum(enrollment_a67) if year >= 2010
gen cummOutlayA67 = sum(outlay_a67) if year >= 2010
gen cummEnrollmentA70 = sum(enrollment_a70) if year >= 2010
gen cummOutlayA70 = sum(outlay_a70) if year >= 2010
global vlist cBaseline cA67 cummEnrollmentA67 cummOutlayA67 cA70 cummEnrollmentA70 cummOutlayA70
format $vlist %10.2f
outsheet year $vlist using age_deviation.csv if year >= 2010, comma replace
\end{Statacode}

\DTLloaddb[headers={Year,\shortstack{ACA\\Cost},\shortstack{Age 67\\Cost},\shortstack{Age 67\\Enrollment\\Deviation},\shortstack{Age 67\\Outlay\\Deviation},\shortstack{Age 70\\Cost},\shortstack{Age 70\\Enrollment\\Deviation},\shortstack{Age 70\\Outlay\\Deviation}}]{age_deviation}{age_deviation.csv}

\begin{table}[ht]
\centering
\caption{Deviation of Total Medicare Costs from ACA (Net Present Value in Billions of 2010 Dollars)}
\DTLdisplaydb{age_deviation}
\label{tab:age_deviation}
\end{table}

\begin{Statacode}{!echo,hide}
clear all
\end{Statacode}

\end{document}