### 2019-11-22 analysis/techdoc/latex/FAM/FAM_tables.tex

\section{Tables}

% FAM initial conditions are:
%Economic: 	Work Status, 	Earnings, 				Wealth
%Health: 		BMI Category, Smoking Category, Hypertension 
%Other: 		Education, 		Partnered, 				Partner Type, Health Insurance

\begin{table}[h]\centering
\begin{tabular}{l l l}
\textbf{Economic Outcomes} & \textbf{Health Outcomes} & \textbf{Other Outcomes}\\
Work Status & BMI Category & Education\\
Earnings & Smoking Category & Partnered\\
Wealth & Hypertension & Partner Type\\
& & Health Insurance
\end{tabular}~\\~\\
\caption{Estimated outcomes in replenishing cohorts module}
\label{tab:initial_conditions_model}
\end{table}

% FAM transition models are:
%Economic: Social Security Claiming, Disability Claiming, Supplemental Security Income Claiming, Non-Zero Capital Income, Capital Income, Non-zero Government Transfers, Government Transfers, Non-zero Wealth, Wealth, Laborforce Status, Full- or Part-time, Any Earnings if Unemployed, Any Earnings if Not in Labor Force, Earnings if Full-time, Earnings if Part-time, Earnings if Unemployed, Earnings if Not in Labor Force
%Health: Mortality, Heart Disease, Stroke, Cancer, Hypertension, Diabetes, Lung Disease, Start Smoking, Stop Smoking, ADL Status, IADL Status, Births/Paternity, Self-reported Health, BMI, Partner Death
%Marital Status (by sex): Exit Single, Exit Cohabitation, Exit Married, Single to Married, Cohabitation to Married, Married to Cohabitation
%Other: Insurance Type

\begin{sidewaystable}
\centering
\begin{tabular}{l l l l}
\textbf{Economic Outcomes} & \textbf{Health Outcomes} &  \textbf{Marital Status} & \textbf{Other Outcomes}\\
Social Security Claiming & Mortality & Exit Single & Insurance Type\\
Disability Claiming & Heart Disease & Exit Cohabitation &\\
Non-Zero Capital Income & Cancer & Exit Married & \\
Capital Income (if non-zero) & Hypertension & Single to Married &\\
Non-zero Government Transfers & Diabetes & Cohabitation to Married & \\
Government Transfers (if non-zero) & Lung Disease & Married to Cohabitation & \\
Non-zero Wealth & Start Smoking & & \\
Wealth (if non-zero) & Stop Smoking & &\\
Labor Force Status (out, unemployed, working) & ADL Status & &\\
Employed Full- or Part-time (if working) & IADL Status & &\\
Any Earnings (if Unemployed) & Births/Paternity & &\\
Any Earnings (if Not in Labor Force) & Self-reported Health & &\\
Earnings (if Full-time) & BMI & &\\
Earnings (if Part-time) & Partner Death & &\\
Earnings (if Unemployed and any) & & &\\
Earnings (if Not in Labor Force and any) & & &\\
\end{tabular}
\caption{Estimated outcomes in transitions module}
\label{tab:transition_model}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\begin{tabular}[h]{l rrrrrrrr}
\input{../tables/FAM/svy_disease_prevalence.tex}
\end{tabular}
\caption{Health condition prevalences in survey data}
\label{tab:svy_disease_prevalence}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.75}{
\begin{tabular}{l p{2.5in}p{2.5in}p{2.5in}p{2.5in}}
& \multicolumn{4}{c}{Survey}\\
Disease & \multicolumn{1}{c}{PSID/HRS} & \multicolumn{1}{c}{NHIS} & \multicolumn{1}{c}{MEPS} & \multicolumn{1}{c}{MCBS} \\
\hline
Cancer &
Has a doctor ever told you that you have cancer or a malignant tumor, excluding minor skin cancers?&
Have you ever been told be a doctor or other health professional that you had … cancer or a malignanacy of any kind? (WHEN RECODED, SKIN CANCERS WERE EXCLUDED)&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 11-21, 24-45&
Has a doctor ever told you that you had any (other) kind of cancer malignancy, or tumor other than skin cancer?\\
\hline
Heart Diseases &
Has a doctor ever told you that you had a heart attack, coronary heart disease, angina, congestive heart failure, or other heart problems?&
Four separate questions were asked about whether ever told by a docotor or oither health professional that had: CHD, Angina, MI, other heart problems.&
Have you ever been told by a doctor or health professional that you have … CHD; Angina; MI; other heart problems&
Six separate questions were asked about whether ever told by a doctor that had: Angina or MI; CHD; other heart problems (included four questions)\\
\hline
Stroke &
Has a doctor ever told you that you had a stroke?&
Have you EVER been told by a doctor or other health professional that you had a stroke?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have a stroke or TIA (transient ischemic attack)&
[Since (PREV$<$ SUPP. RD. INT. DATE),] has a doctor (ever) told (you/SP) that (you/he/she) had a stroke, a brain hemorrhage, or a cerebrovascular accident?\\
\hline
Diabetes&
Has a doctor ever told you that you have diabetes or high blood sugar?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
If Female, add: [Other than during pregnancy,] Have you hever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
Has a doctor (ever) told (you/SP) that (you/he/she) had diabetes, high blood sugar, or sugar in (your/his/her) urine? [DO NOT INCLUDE BOERDERLINE PREGNANCY, OR PRE-DIABETEIC DIABETES.]\\
\hline
Hypertension&
Has a doctor ver told you that you have high blood pressure or hypertension?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Has a doctor (ever) told (you/SP) that (you/he/she) (still) (had) (have/has) hypertension, sometimes called high blood pressure?\\
\hline
Lung Disease&
Has a doctor ever told you that you have chronic lung disease such a schronic bronchitis or emphysema? [IWER: DO NOT INCLUDE ASTHMA]&
Question 1: During the PAST 12 MONTHS, have you ever been told by a doctor or other health professional that you had chronic bronchitis? Question 2: Have you EVER been told by a docotor or other health professional that you had … emphysema?&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 127, 129-312&
Has a doctor (ever) told (you/SP) that (you/he/she) had emphysema, asthma, or COPD? [COPD=CHRONIC OBSTRUCTIVE PULMONARY DISEASE.]\\
\hline
Overweight & \multicolumn{4}{c}{\multirow{2}{*}{Self-reported body weight and height}}\\
\hhline{-~~~~}
Obese & \multicolumn{4}{c}{} \\
\hline
\end{tabular}
}
\caption{Survey questions used to determine health conditions}
\label{tab:svy_disease_questions}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Dynamically fill in numeric results using Sweave
\centering
\scalebox{0.9}{
\begin{tabular}{lllccr}
 &  &  & Type & At risk & Mean/fraction\\
\hline
\input{../tables/FAM/transitioned_outcomes_disease.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_smoking.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_logbmi.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_adlstat.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_iadlstat.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_work.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_lfpben.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_mstat.tex}
\hhline{~-----}
\input{../tables/FAM/transitioned_outcomes_birth.tex}
\hline
\input{../tables/FAM/transitioned_outcomes_financial.tex}
\hline
\end{tabular}
}
\caption[Outcomes in the transition model]{Outcomes in the transition model. Estimation sample is PSID 1999-2013 waves.}
\label{tab:transitioned_outcomes}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Write a script that dynamically generates this table from the .est files
\centering
\scalebox{0.53}{
\begin{tabular}{l*{20}{c}}
Value at time $T-1$ & \multicolumn{20}{c}{Outcome at time $T$}  \\
\hline
 & Heart  &  & & Lung  & & &  &  & Smoking  &  &  & &  &  & & Nursing & & &  Nonzero & \\
 & disease & hypertension & stroke & disease & diabetes & cancer & disability & mortality & status & BMI & Any HI & DI Claim & SS Claim & DB Claim & SSI Claim & Home & Work & Earnings & Wealth & Wealth\\
Heart disease &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Blood pressure & \checkmark &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Stroke &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Lung disease &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Diabetes & \checkmark & \checkmark & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Cancer &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Disability &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DI &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SS &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DB &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SSI &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  &  &  & \\
Work &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Earnings &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Nonzero wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Nursing home stay &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark &  &  & \checkmark & \checkmark
\end{tabular}
}
\caption[Restrictions on transition model]{Restrictions on transition model. \checkmark indicates that an outcome at time $T-1$ is allowed in the transition model for an outcome at time $T$.}
\label{tab:transition_restrictions}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{1}{
\input{../tables/FAM/controlvar_descs.tex}
}
\caption[Descriptive statistics for stock population]{Desciptive statistics for variables in 2009 PSID ages 25+ sample used as simulation stock population}
\label{tab:controlvar_descs}
\end{table}

%\begin{table}
%% \todo Dynamically fill in numeric results using Sweave and reduce the number of sig. digits after decimal
%\centering
%\scalebox{0.9}{
%\begin{tabular}{ccccrr}
% &  &  & Selection & 1992 & 2004 \\
%\hline
%\input{../tables/FAM/desc_init_conditions_binary.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_bmistat.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_smoking.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_funcstat.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_continuous.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_censcont.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_censdiscrete.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_earlyagedb.tex}
%\hhline{~-----}
%\input{../tables/FAM/desc_init_conditions_normalagedb.tex}
%\hline
%\input{../tables/FAM/desc_init_conditions_covars.tex}
%\hline
%\end{tabular}
%}
%\caption{Initial conditions used for estimation (1992) and simulation (2004)}
%\label{tab:desc_init_conditions}
%\end{table}

\begin{sidewaystable}
\centering
\scalebox{1}{
\input{../tables/FAM/latent_model_mean_est.tex}
}
\caption[Parameter estimates for latent model: conditional means and thresholds]{Parameter estimates for latent model: conditional means and thresholds. Sample is respondents age 25-30 in 2005-2011 PSID waves}
\label{tab:latent_model_mean_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{1}{
\input{../tables/FAM/latent_model_vcmat_est.tex}
}
\caption[Parameter estimates for latent model: parameterized covariance matrix]{Parameter estimates for latent model: parameterized covariance matrix. Sample is respondents age 25-30 in 2005-2011 PSID waves}
\label{tab:latent_model_vcmat_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_health.tex}
\caption{Health and risk factor trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_health}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_educ.tex}
\caption{Education trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_educ}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FAM/new_cohort_trends_social.tex}
\caption{Social trends for replenishing cohorts, prevalences relative to 2009}
\label{tab:new_cohort_trends_social}
\end{sidewaystable}

%\begin{table}
%\centering
%\input{../tables/FAM/NHEA_adjustment.tex}
%\caption{Per capita medical spending by payment source, age group, and year}
%\label{tab:NHEA_adjustment}
%\end{table}

%\begin{table}
%\centering
%\scalebox{0.8}{
%\input{../tables/FAM/time_series_by_year.tex}
%}
%\caption{Assumptions for each calendar year}
%\label{tab:time_series_by_year}
%\end{table}

%\begin{table}
%\centering
%\scalebox{0.8}{
%\input{../tables/FAM/time_series_by_yob.tex}
%}
%\caption[Assumptions for each birth year]{Assumptions for each birth year. In years after 1960, all values are held constant at their 1960 levels.}
%\label{tab:time_series_by_yob}
%\end{table}

\begin{table}
\centering
\input{../tables/FAM/crossval_unweighted.tex}
\caption{Crossvalidation of 1999 cohort: Mortality in 2001, 2007, and 2013}
\label{tab:crossval_unweighted}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_demog.tex}
\caption{Crossvalidation of 1999 cohort: Demographic outcomes in 2001, 2007, and 2013}
\label{tab:crossval_demog}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_binhlth.tex}
\caption{Crossvalidation of 1999 cohort: Binary health outcomes in 2001, 2007, and 2013}
\label{tab:crossval_binhlth}
\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_risk.tex}
\caption{Crossvalidation of 1999 cohort: Risk factor outcomes in 2001, 2007, and 2013}
\label{tab:crossval_risk}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FAM/crossval_binecon.tex}
\caption{Crossvalidation of 1999 cohort: Binary economic outcomes in 2001, 2007, and 2013}
\label{tab:crossval_binecon}
\end{table}

%\begin{table}
%% \todo clean up the significant digits in this table
%% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
%\centering
%\input{../tables/FAM/crossval_cntecon.tex}
%\caption{Crossvalidation of 1999 cohort: Simulated vs reported continuous economic outcomes in 2001, 2007, and 2013}
%\label{tab:crossval_cntecon}
%\end{table}

\begin{sidewaystable}
\centering
\input{../tables/FAM/external_pop.tex}
\caption{Population forecasts: Census compared to FAM}
\label{tab:external_pop}
\end{sidewaystable}