### 2019-11-22 analysis/techdoc/latex/FAM/acknowledgments.tex

\section{Acknowledgments}

The Future Elderly Model and Future Americans Model has been developed by a large team over the last decade.  Jay Bhattacharya, Eileen Crimmins, Christine Eibner, \'Etienne Gaudette, Geoff Joyce, Darius Lakdawalla,
Pierre-Carl Michaud, and Julie Zissimopoulos have all provided expert guidance.  Adam Gailey, Baoping Shang, and Igor Vaynman provided programming and analytic 
support during the first years of FEM development at RAND.  Jeff Sullivan then led the technical development for several years.  More recently, the University of 
Southern California research programming team has supported model development, including FAM development.  These programmers include Patricia St. Clair, Laura Gascue, Henu Zhao, and Yuhui Zheng. 
Barbara Blaylock, Malgorzata Switek, and Wendy Cheng have greatly aided model development while working as research assistants at USC.