### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation.tex

\section{Data sources used for estimation}
The Panel Survey of Income Dynamics is the main data source for the model. We estimate models
for assigning characteristics for the replacement cohorts in Replenishing Conditions Module.  These
are summarized in Table \ref{tab:initial_conditions_model}.  We estimate transition models for 
the entire PSID population in the Transition Model Module.  Transitioned outcomes are described in
Table \ref{tab:transition_model}.

\input{data_sources_estimation_psid.tex}
\input{data_sources_estimation_hrs.tex}
%\input{data_sources_estimation_ssearnings.tex}
%\input{data_sources_estimation_nhis.tex}
\input{data_sources_estimation_meps.tex}
\input{data_sources_estimation_mcbs.tex}