### 2019-11-22 analysis/techdoc/latex/FAM/data_sources_estimation_hrs.tex

\subsection{Health and Retirement Study}
The Health and Retirement Study (HRS), waves 1998-2012 are pooled with the PSID for estimation of mortality and 
widowhood models. The HRS has a similar structure to the PSID, with interviews occurring every two years. The HRS data
is harmonized to the PSID for all relevant variables. We use the dataset created by RAND (RAND HRS, version O) as our basis 
for the analysis. We use all cohorts in the analysis. When appropriately weighted, the HRS in 2010 is representative of U.S. households 
where at least one member is at least 51.  Compared to the PSID, the HRS includes more older Hispanics and interviews more respondents 
once they have entered nursing homes.