### 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model.tex

\section{Model for replenishing cohorts}
\label{sec:model_for_new_cohorts}
We first discuss the empirical strategy, then present the model and estimation 
results. The model for replinishing cohorts integrates information coming from trends 
among younger cohorts with the joint distribution of outcomes in the current 
population of age 25 respondents in the PSID.

%\input{new_cohorts_model_empirical_strategy.tex}
\input{new_cohorts_model_model_and_estimation.tex}
\input{new_cohorts_model_trends.tex}