### 2019-11-22 analysis/techdoc/latex/FAM/new_cohorts_model_trends.tex

\subsection{Trends for replenishing cohorts}
Using the jointly estimated models previously described, we then assign outcomes to the replenishing cohorts, imposing trends for some health, 
risk factor, and social outcomes.  We currently impose trends on BMI, education, number of children, marital status, hypertension, and smoking 
status for these 25-26 year olds.  These trends are estimated using the National Health Interview Survey (health and risk factors) or the American Community 
Survey (social outcomes). All trends are halted after 2029. The trends are shown in Table \ref{tab:new_cohort_trends_health}, Table \ref{tab:new_cohort_trends_educ} and 
Table \ref{tab:new_cohort_trends_social}.