### 2019-11-22 analysis/techdoc/latex/FAM/validation_external_corroboration.tex

\subsection{External Corroboration}
Finally, we compare FAM population forecasts to Census forecasts of the US population..  Here, we focus on the full PSID population (25 and older) and those 65 and 
older.  For this exercise, we begin the simulation in 2009 and simulate the full population through 2049.  Population projections are compared to the 2012 Census 
projections for years 2012 through 2049.  See resuts in Table \ref{tab:external_pop}.  By 2049, FAM forecasts for 25 and older remain within 2\% of Census forecasts.