### 2019-11-22 analysis/techdoc/latex/FEM/FEM_tables.tex

\section{Tables}

\begin{sidewaystable}
\centering
\begin{tabular}[h]{l rrrrrrrr}
\input{../tables/FEM/svy_disease_prevalence.tex}
\end{tabular}
\caption{Health condition prevalences in survey data}
\label{tab:svy_disease_prevalence}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.75}{
\begin{tabular}{l p{2.5in}p{2.5in}p{2.5in}p{2.5in}}
& \multicolumn{4}{c}{Survey}\\
Disease & \multicolumn{1}{c}{HRS} & \multicolumn{1}{c}{NHIS} & \multicolumn{1}{c}{MEPS} & \multicolumn{1}{c}{MCBS} \\
\hline
Cancer &
Has a doctor ever told you that you have cancer or a malignant tumor, excluding minor skin cancers?&
Have you ever been told be a doctor or other health professional that you had … cancer or a malignanacy of any kind? (WHEN RECODED, SKIN CANCERS WERE EXCLUDED)&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 11-21, 24-45&
Has a doctor ever told you that you had any (other) kind of cancer malignancy, or tumor other than skin cancer?\\
\hline
Heart Diseases &
Has a doctor ever told you that you had a heart attack, coronary heart disease, angina, congestive heart failure, or other heart problems?&
Four separate questions were asked about whether ever told by a docotor or oither health professional that had: CHD, Angina, MI, other heart problems.&
Have you ever been told by a doctor or health professional that you have … CHD; Angina; MI; other heart problems&
Six separate questions were asked about whether ever told by a doctor that had: Angina or MI; CHD; other heart problems (included four questions)\\
\hline
Stroke &
Has a doctor ever told you that you had a stroke?&
Have you EVER been told by a doctor or other health professional that you had a stroke?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have a stroke or TIA (transient ischemic attack)&
[Since (PREV$<$ SUPP. RD. INT. DATE),] has a doctor (ever) told (you/SP) that (you/he/she) had a stroke, a brain hemorrhage, or a cerebrovascular accident?\\
\hline
Diabetes&
Has a doctor ever told you that you have diabetes or high blood sugar?&
If Female, add: [Other than during pregnancy,] Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
If Female, add: [Other than during pregnancy,] Have you hever been told by a doctor or health professional that you have diabetes or sugar diabetes?&
Has a doctor (ever) told (you/SP) that (you/he/she) had diabetes, high blood sugar, or sugar in (your/his/her) urine? [DO NOT INCLUDE BOERDERLINE PREGNANCY, OR PRE-DIABETEIC DIABETES.]\\
\hline
Hypertension&
Has a doctor ver told you that you have high blood pressure or hypertension?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Have you EVER been told by a doctor or other health professional that you had Hypertension, also called high blood pressure?&
Has a doctor (ever) told (you/SP) that (you/he/she) (still) (had) (have/has) hypertension, sometimes called high blood pressure?\\
\hline
Lung Disease&
Has a doctor ever told you that you have chronic lung disease such a schronic bronchitis or emphysema? [IWER: DO NOT INCLUDE ASTHMA]&
Question 1: During the PAST 12 MONTHS, have you ever been told by a doctor or other health professional that you had chronic bronchitis? Question 2: Have you EVER been told by a docotor or other health professional that you had … emphysema?&
List all the conditions that have bothered (the person) from (START time) to (END time) CCS codes for the conditions list are 127, 129-312&
Has a doctor (ever) told (you/SP) that (you/he/she) had emphysema, asthma, or COPD? [COPD=CHRONIC OBSTRUCTIVE PULMONARY DISEASE.]\\
\hline
Overweight & \multicolumn{4}{c}{\multirow{2}{*}{Self-reported body weight and height}}\\
\hhline{-~~~~}
Obese & \multicolumn{4}{c}{} \\
\hline
\end{tabular}
}
\caption{Survey questions used to determine health conditions}
\label{tab:svy_disease_questions}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{0.85}{
\begin{tabular}[h]{p{0.95in}p{1.2in}p{1.6in}p{1.8in}}
\textbf{Conditions} & \textbf{Data source} & \textbf{Projection method} & \textbf{Other sources}\\
\hline
Diabetes\newline Heart disease\newline Hypertension\newline & 
National Health Interview Survey 1997-2006 & 
Use synthetic cohort approach to estimate age-specific incidence rate for each condition & 
There are other forecasts \citep{honeycutt2003dynamic,mainous2007impact} for the trends of diabetes in the U.S population; we compare their forecasts to ours and they are reasonably close \\
\hline
Overweight and obese & {
Prevalence of over-weight and obese for aged 46-56 from year 2001 to 2030, generated by Ruhm upon request} & {
Assume annual rate of change during year 2031-2050 linearly decreases from the 2030 rate to zero in 2050} & \citet{ruhm2007current} \\
\hline
Ever-smoked and smoking now & {
Forecast of prevalence of ever-smoked and smoking now for aged 45-54 from year 2005 to 2025, by \citet{levy2006smoking}} & {
For ever-smoked, assume that the prevalence at age 45-54 in year 2035 (2045) is the same as prevalence at age 35-44 (25-34) in year 2025. Assume that the annual change in prevalence at age 45-54 in year 2046-2050 the same as average in 2040-2045. For smoking-now, after year 2025, use the moving average of the past five years} & {
}\\
\hline
Any DB from current job & {
} & {
Assume annual relative declining rate for DB entitlement decrease by 2\% a year} & {
Historical trends of DB participation rates among all persons by different birth cohorts and by age, by \citet{poterba2007decline}}\\
\hline
Any DC from current job & {
} & {
Assume annual relative increasing rate for DC entitlement increase by 2\% a year until 2026 then stays the same after 2026} & {
Forecast of DC participation rates among all persons by different birth cohorts and by age, by \citet{poterba2008new}}\\
\hline
Population size 50-52\newline Male\newline Hispanic\newline Non-Hispanic black\newline & {
Census Bureau 2000-2010 Intercensal Population Estimates, 2012 National Population Estimates, and 2012 National Population Projections} & {
Projected 2060 - 2080 using linear trend based on 2040-2060} & {
} \\
\hline
\end{tabular}
}
\caption{Data sources and methods for projecting future cohort trends}
\label{tab:cohort_projection_data_methods}
\end{table}

\begin{sidewaystable}
\centering
\input{../tables/FEM/baseline_trends.tex}
\caption{Projected baseline trends for future cohorts}
\label{tab:baseline_trends}
\end{sidewaystable}

\begin{table}
\centering
\input{../tables/FEM/prevalence_1978_2004.tex}
\caption[Prevalence of obesity, hypertension, diabetes and current smokers among ages 46-56 in 1978 and 2004]{Prevalence of obesity, hypertension, diabetes and current smokers among ages 46-56 in 1978 and 2004. Prevalence in 1978 is based on NHANES II 1976-1980; Prevalence in 2004 is based on NHANES 2003-2004. 
BMI is calculated using self-reported weight and height.}
\label{tab:prevalence_1978_2004}
\end{table}

\begin{sidewaystable}
% \todo Dynamically fill in numeric results using Sweave
% For now, the numerical results can be copied from analysis/techdoc/table6.csv
\centering
\begin{tabular}{lllccr}
 &  &  & Type & At risk & Mean/fraction\\
\hline
\input{../tables/FEM/transitioned_outcomes_disease.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_smoking.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_logbmi.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_adlstat.tex}
\hhline{~-----}
\input{../tables/FEM/transitioned_outcomes_iadlstat.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_lfpben.tex}
\hline
\input{../tables/FEM/transitioned_outcomes_financial.tex}
\hline
\end{tabular}
\caption[Outcomes in the transition model]{Outcomes in the transition model. Estimation sample is HRS 1991-2008 waves.}
\label{tab:transitioned_outcomes}
\end{sidewaystable}

\begin{sidewaystable}
% \todo Write a script that dynamically generates this table from the .est files
\centering
\scalebox{0.53}{
\begin{tabular}{l*{20}{c}}
Value at time $T-1$ & \multicolumn{20}{c}{Outcome at time $T$}  \\
\hline
 & Heart  &  & & Lung  & & &  &  & Smoking  &  &  & &  &  & & Nursing & & &  Nonzero & \\
 & disease & hypertension & stroke & disease & diabetes & cancer & disability & mortality & status & BMI & Any HI & DI Claim & SS Claim & DB Claim & SSI Claim & Home & Work & Earnings & Wealth & Wealth\\
Heart disease &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Blood pressure & \checkmark &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Stroke &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Lung disease &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Diabetes & \checkmark & \checkmark & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Cancer &  &  & \checkmark &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Disability &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DI &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SS &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed DB &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Claimed SSI &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark &  &  &  &  & \\
Work &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Earnings &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark\\
Nonzero wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Wealth &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
Nursing home stay &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \checkmark & \checkmark &  &  & \checkmark & \checkmark
\end{tabular}
}
\caption[Restrictions on transition model]{Restrictions on transition model. \checkmark indicates that an outcome at time $T-1$ is allowed in the transition model for an outcome at time $T$.}
\label{tab:transition_restrictions}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\input{../tables/FEM/controlvar_descs.tex}
\caption[Desciptive statistics for exogeneous control variables]{Desciptive statistics for exogeneous control variables in 2004 HRS ages 51+ sample used as simulation stock population}
\label{tab:controlvar_descs}
\end{sidewaystable}

%\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
%\scalebox{0.75}{
%\input{../tables/FEM/validation_waves_4_7_9.tex}
%}
%\caption{Simulated and actual outcomes in 1998, 2004, and 2008}
%\label{tab:validation_waves_4_7_9}
%\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_unweighted.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported mortality and nursing home outcomes in 2000, 2006, and 2012}
\label{tab:crossval_unweighted}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_demog.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported demographic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_demog}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_binhlth.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported binary health outcomes in 2000, 2006, and 2012}
\label{tab:crossval_binhlth}
\end{table}

\clearpage

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_risk.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported risk factor outcomes in 2000, 2006, and 2012}
\label{tab:crossval_risk}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_binecon.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported binary economic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_binecon}
\end{table}

\begin{table}
% \todo clean up the significant digits in this table
% \todo create a variable sort order (file that maps variable names to integers) that can be merged on and then labels can be sorted in this table
\centering
\input{../tables/FEM/crossval_cntecon.tex}
\caption{Crossvalidation of 1998 cohort: Simulated vs reported continuous economic outcomes in 2000, 2006, and 2012}
\label{tab:crossval_cntecon}
\end{table}

\begin{table}
\centering
\begin{tabular}{cl|rr|rccl|rr|r}
 & \multicolumn{4}{c}{MEPS 2001 (ages 51+)} & \multicolumn{2}{c}{}  & \multicolumn{4}{c}{HRS 1998 (ages 51+)}\\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{ADL limitation} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{ADL limitation} & \multicolumn{1}{c}{} \\
%\hhline{~~--~~~~--~}
 &  & No & Yes & All & & &  & No & Yes & All\\
\hhline{~----~~----}
IADL & No & 91.5 & 0.4 & 92.0 & & IADL & No & 82.0 & 10.5 & 92.5\\
limitation & Yes & 4.4 & 3.7 & 8.0 & & limitation & Yes & 3.1 & 4.5 & 7.5\\
\hhline{~----~~----}
  & All & 95.9 & 4.1 & 100.0 & &  & All & 85.1 & 14.9 & 100.0\\
\end{tabular}
\caption[Prevalence of IADL and ADL limitations among ages 51+ in MEPS 2001 and HRS 1998]
{Prevalence of IADL and ADL limitations among ages 51+ in MEPS 2001 and HRS 1998.
The IADL limitations in MEPS are defined as receiving help or supervision using the 
telephone, paying bills, taking medications, preparing light meals, doing laundry, 
or going shopping; the ADL limitations in HRS are defined as receiving help or 
supervision with personal care such as bathing, dressing, or getting around the 
house. 
The IADL limitations in HRS are defined as having any difficulty in at least one of 
the following activities: using the phone, taking medications, and managing money. 
The ADL limitations in HRS are defined as having any difficulty in at least one of the 
following activities: bathing, dressing, eating, walking across the room, and 
getting out of bed. 
}
\label{tab:meps_hrs_adl_iadl_prevalence}
\end{table}

\begin{table}
\centering
\begin{tabular}{cl|rr|rccl|rr|r}
 & \multicolumn{4}{c}{MEPS 2001 (ages 51+)} & \multicolumn{2}{c}{}  & \multicolumn{4}{c}{HRS 1998 (ages 51+)}\\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{Physical function} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{Physical function} & \multicolumn{1}{c}{} \\
\multicolumn{2}{c}{} & \multicolumn{2}{c}{limitation} & \multicolumn{4}{c}{} & \multicolumn{2}{c}{limitation} & \multicolumn{1}{c}{} \\
%\hhline{~~--~~~~--~}
 &  & No & Yes & All & & &  & No & Yes & All\\
\hhline{~----~~----}
IADL & No & 61.6 & 30.4 & 92.0 & & IADL & No & 60.0 & 32.5 & 92.5\\
limitation & Yes & 0.3 & 7.8 & 8.0 & & limitation & Yes & 1.0 & 6.5 & 7.5\\
\hhline{~----~~----}
  & All & 61.9 & 38.2 & 100.0 & &  & All & 61.0 & 39.0 & 100.0\\
\end{tabular}
\caption[Prevalence of IADL limitation and physical function limitation among ages 51+ in MEPS 2001 and HRS 1998]
{Prevalence of IADL limitation and physical function limitation among ages 51+ in MEPS 2001 and HRS 1998. 
The definition of IADL limitation is the same as in Table \ref{tab:meps_hrs_adl_iadl_prevalence}.
Physical function limitation in MEPS indicates that at least one of the following is true: 1) receiving help or 
supervision with bathing, dressing or walking around the house; 2) being limited in work/housework; 3) having 
difficulty walking, climbing stairs, grasping objects, reaching overhead, lifting, bending or stooping, or 
standing for long periods of time; or 4) having difficulty in hearing or vision.
Physical function limitation in HRS indicates at least one of the following is true: 1) having any difficulty in 
bathing/dressing/eating/walking across the room/getting out of bed; 2) limited in work/housework; or 3) limited in 
any other activities.
}
\label{tab:meps_hrs_physlim_iadl_prevalence}
\end{table}

\begin{table}
\centering
\scalebox{0.86}{
\begin{tabular}{l|rrrrr}
  & Model I & Model II & Model III & Model IV & Model V\\
\hline
Constant & 0.877*** & 0.898*** & 0.874*** & 0.839*** & 0.869***\\
  & (0.002) & (0.003) & (0.005) & (0.002) & (0.003)\\
Physical function limitation & -0.115*** & -0.098*** & -0.094*** &   &  \\
  & (0.004) & (0.005) & (0.004) &   &  \\
IADL limitation & -0.041 & -0.019 & -0.008 &   &  \\
  & (0.037) & (0.042) & (0.036) &   &  \\
IADL limitation * Physical function limitation & -0.150*** & -0.156*** & -0.162*** &   &  \\
  & (0.037) & (0.044) & (0.037) &   &  \\
IADL limitation, no ADL limitation &   &   &   & -0.182*** &  \\
  &   &   &   & (0.009) &  \\
Any ADL limitation &   &   &   & -0.344*** &  \\
  &   &   &   & (0.010) &  \\
Ever diagnosed with cancer &   & -0.011 & -0.015** &   & -0.030***\\
  &   & (0.009) & (0.007) &   & (0.010)\\
Ever diagnosed with diabetes &   & -0.034*** & -0.032*** &   & -0.054***\\
  &   & (0.007) & (0.005) &   & (0.007)\\
Ever diagnosed with high blood pressure &   & -0.030*** & -0.028*** &   & -0.043***\\
  &   & (0.004) & (0.004) &   & (0.005)\\
Ever diagnosed with heart disease &   & -0.024*** & -0.029*** &   & -0.055***\\
  &   & (0.006) & (0.005) &   & (0.006)\\
Ever diagnosed with lung disease &   & -0.036*** & -0.032*** &   & -0.055***\\
  &   & (0.009) & (0.007) &   & (0.010)\\
Ever diagnosed with stroke &   & -0.045*** & -0.046*** &   & -0.115***\\
  &   & (0.012) & (0.008) &   & (0.013)\\
Age 65-74 &   &   & 0.010** &   &  \\
  &   &   & (0.004) &   &  \\
Age 75 and over &   &   & 0.015*** &   &  \\
  &   &   & (0.005) &   &  \\
Male &   &   & 0.028*** &   &  \\
  &   &   & (0.004) &   &  \\
Non-Hispanic black &   &   & 0.008 &   &  \\
  &   &   & (0.007) &   &  \\
Hispanic &   &   & -0.001 &   &  \\
  &   &   & (0.007) &   &  \\
Less than HS &   &   & -0.022*** &   &  \\
  &   &   & (0.005) &   &  \\
Some college &   &   & 0.016*** &   &  \\
  &   &   & (0.005) &   &  \\
College grad &   &   & 0.037*** &   &  \\
  &   &   & (0.005) &   &  \\
Census region: Northeast &   &   & 0.003 &   &  \\
  &   &   & (0.005) &   &  \\
Census region: Midwest &   &   & 0.004 &   &  \\
  &   &   & (0.005) &   &  \\
Census region: West &   &   & -0.012** &   &  \\
  &   &   & (0.005) &   &  \\
Marital status:widowed &   &   & 0.003 &   &  \\
  &   &   & (0.005) &   &  \\
Marital status: single &   &   & -0.013*** &   &  \\
  &   &   & (0.005) &   &  \\
\hline
$N$ & 7,358 & 7,317 & 7,317 & 7,361 & 7,322\\
\hline
Adjusted $R^2$ & .24 & .27 & .29 & .18 & .11\\
\hline
\multicolumn{6}{c}{* $p<0.10$, ** $p<0.05$, *** $p<0.01$}  
\end{tabular}
}
\caption[OLS regressions of EQ-5D utility index among ages 51+ in MEPS 2001]{OLS regressions of EQ-5D utility index among ages 51+ in MEPS 2001. $p$-values in parentheses. Data source: MEPS 2001 (ages 51+). EQ-5D scoring algorithm is based on \citet{shaw2005us}.}
\label{tab:meps_eq5d_model_est}
\end{table}

\begin{table}
\centering
\begin{tabular}{l|r}
\hline
Ever diagnosed with cancer & -0.020*\\
  & (0.001)\\
Ever diagnosed with diabetes & -0.042*\\
  & (0.001)\\
Ever diagnosed with heart disease & -0.044*\\
  & (0.001)\\
Ever diagnosed with high blood pressure & -0.034*\\
  & (0.001)\\
Ever diagnosed with lung disease & -0.054*\\
  & (0.001)\\
Ever diagnosed with stroke & -0.067*\\
  & (0.002)\\
IADL limitation only & -0.160*\\
  & (0.002)\\
One or two ADL limitations & -0.099*\\
  & (0.001)\\
Three or more ADL limitations  & -0.149*\\
  & (0.002)\\
Constant & 0.881*\\
  & (0.001)\\
\hline
$N$ & 19,676  \\
\hline
Adjusted $R^2$ & 0.67\\
\hline
\multicolumn{2}{c}{* $p<0.01$}
\end{tabular}
\caption[OLS regression of the predicted EQ-5D index score against chronic conditions and FEM-type functional status specification]
{OLS regression of the predicted EQ-5D index score against chronic conditions and FEM-type functional status 
specification. $p$-values in parentheses. Data source: Health and Retirement Study, 1998. Sample included the 
age 51 and over community respondents. EQ-5D score was predicted using Model II in Table \ref{tab:meps_eq5d_model_est}.   
}
\label{tab:fem_eq5d_model_est}
\end{table}

\begin{sidewaystable}
\centering
\begin{tabular}{|l|lrr|rrrrrr|}
 \multicolumn{4}{l}{}  & \multicolumn{6}{c}{Prevalence of chronic conditions}\\
\hhline{~~~~------}
\multicolumn{1}{l}{}                  &           & Predicted   &         &        &          & Heart   &              & Lung    &       \\
\multicolumn{1}{l}{Functional status} &  Estimate & EQ-5D score & Age     & Cancer & Diabetes & disease & Hypertension & disease & Stroke\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{Healthy}} & Mean & 0.841 & 63.9 & 0.114 & 0.142 & 0.189 & 0.481 & 0.072 & 0.045\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.042 & 10.0 & 0.318 & 0.349 & 0.391 & 0.500 & 0.258 & 0.207\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.000 & 0.0 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{IADL limitation only - not in nursing home}} & Mean & 0.655 & 70.0 & 0.151 & 0.223 & 0.334 & 0.581 & 0.152 & 0.162\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.054 & 12.9 & 0.358 & 0.416 & 0.472 & 0.493 & 0.359 & 0.368\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.005 & 0.005 & 0.006 & 0.006 & 0.005 & 0.005\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{1 or 2 ADL limitations - not in nursing home}} & Mean & 0.705 & 69.1 & 0.180 & 0.291 & 0.393 & 0.665 & 0.194 & 0.162\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.057 & 12.0 & 0.384 & 0.454 & 0.488 & 0.472 & 0.395 & 0.368\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.000 & 0.1 & 0.003 & 0.003 & 0.004 & 0.003 & 0.003 & 0.003\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{3 or more ADL limitations  - not in nursing home}} & Mean & 0.636 & 70.9 & 0.181 & 0.361 & 0.467 & 0.691 & 0.216 & 0.314\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.063 & 13.2 & 0.385 & 0.480 & 0.499 & 0.462 & 0.412 & 0.464\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.005 & 0.006 & 0.006 & 0.006 & 0.005 & 0.006\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{Nursing home residency}} & Mean & 0.568 & 82.0 & 0.144 & 0.256 & 0.428 & 0.651 & 0.142 & 0.372\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.077 & 10.0 & 0.351 & 0.437 & 0.495 & 0.477 & 0.349 & 0.484\\
\multicolumn{1}{|p{2in}|}{} & Std. err & 0.001 & 0.2 & 0.006 & 0.008 & 0.009 & 0.008 & 0.006 & 0.009\\
\hline
\multicolumn{1}{|p{2in}|}{\multirow{3}{2in}{All}} & Mean & 0.808 & 65.2 & 0.125 & 0.170 & 0.229 & 0.514 & 0.094 & 0.076\\
\multicolumn{1}{|p{2in}|}{} & Sd & 0.083 & 10.9 & 0.331 & 0.375 & 0.420 & 0.500 & 0.291 & 0.266\\
\multicolumn{1}{|p{2in}|}{}  & Std. err & 0.000 & 0.0 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001 & 0.001\\
\hline
\end{tabular}
\caption[Average predicted EQ-5D score, age, and prevalence of chronic conditions by functional status for the stock FEM simulation sample]
{Average predicted EQ-5D score, age, and prevalence of chronic conditions by functional status for the stock FEM simulation sample 
(ages 51 and over in 2004).
EQ-5D scores were predicted according to parameter estimates in Table \ref{tab:fem_eq5d_model_est}. The predicted 
score for nursing home residents is reduced by 10\% to account for the fact that the estimation sample in Table 
\ref{tab:fem_eq5d_model_est} only includes non-nursing home residents.
}
\label{tab:fem_stock_predicted_eq5d}
\end{sidewaystable}
\clearpage

\begin{table}
% \todo Dynamically fill in numeric results using Sweave and reduce the number of sig. digits after decimal
\centering
\scalebox{0.9}{
\begin{tabular}{ccccrr}
 &  &  & Selection & 1992 & 2004 \\
\hline
\input{../tables/FEM/desc_init_conditions_binary.tex}
\hline
\input{../tables/FEM/desc_init_conditions_bmistat.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_smoking.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_funcstat.tex}
\hline
\input{../tables/FEM/desc_init_conditions_continuous.tex}
\hline
\input{../tables/FEM/desc_init_conditions_censcont.tex}
\hline
\input{../tables/FEM/desc_init_conditions_censdiscrete.tex}
\hline
\input{../tables/FEM/desc_init_conditions_earlyagedb.tex}
\hhline{~-----}
\input{../tables/FEM/desc_init_conditions_normalagedb.tex}
\hline
\input{../tables/FEM/desc_init_conditions_covars.tex}
\hline
\end{tabular}
}
\caption{Initial conditions used for estimation (1992) and simulation (2004)}
\label{tab:desc_init_conditions}
\end{table}

\begin{sidewaystable}
\centering
\scalebox{0.53}{
\input{../tables/FEM/latent_model_mean_est.tex}
}
\caption[Parameter estimates for latent model: conditional means and thresholds]{Parameter estimates for latent model: conditional means and thresholds. Sample is respondents age 50-55 in 1992 HRS wave}
\label{tab:latent_model_mean_est}
\end{sidewaystable}

\begin{sidewaystable}
\centering
\scalebox{0.53}{
\input{../tables/FEM/latent_model_vcmat_est.tex}
}
\caption[Parameter estimates for latent model: parameterized covariance matrix]{Parameter estimates for latent model: parameterized covariance matrix. Sample is respondents age 50-55 in 1992 HRS wave}
\label{tab:latent_model_vcmat_est}
\end{sidewaystable}

\begin{table}
\centering
\input{../tables/FEM/NHEA_adjustment.tex}
\caption{Per capita medical spending by payment source, age group, and year}
\label{tab:NHEA_adjustment}
\end{table}

\begin{table}
% \todo Generate this table dynamically using Sweave
% \todo add cancer, lung disease and stroke
\centering
\scalebox{0.9}{
\begin{tabular}{lrrr}
Year & 2010 & 2030 & 2050\\
\hline
\input{../tables/FEM/statusquo_results_popsize.tex}
\hline
\input{../tables/FEM/statusquo_results_conditions.tex}
\hline
\input{../tables/FEM/statusquo_results_labor.tex}
\hline
\input{../tables/FEM/statusquo_results_govrev.tex}
\hline
\input{../tables/FEM/statusquo_results_govexp.tex}
\hline
\input{../tables/FEM/statusquo_results_totmd.tex}
\hline
\end{tabular}
}
\caption{Simulation results for status quo scenario}
\label{tab:status_quo_results}
\end{table}

\begin{sidewaystable}
% \todo Generate this table dynamically using Sweave
\centering
\begin{tabular}{lrrcrrcrr}
 & \multicolumn{2}{c}{}                   &  & \multicolumn{2}{c}{Relative Change}  &  & \multicolumn{2}{c}{Absolute Change} \\
 & \multicolumn{2}{c}{``Obese 1980'' Estimates}  &  & \multicolumn{2}{c}{from Status Quo}  &  & \multicolumn{2}{c}{from Status Quo} \\
\hhline{~--~--~--}
Year & 2030 & 2050 &  & 2030 & 2050 &  & 2030 & 2050\\
\hline
\input{../tables/FEM/obesity_results_popsize.tex}
\hline
\input{../tables/FEM/obesity_results_conditions.tex}
\hline
\input{../tables/FEM/obesity_results_labor.tex}
\hline
\input{../tables/FEM/obesity_results_govrev.tex}
\hline
\input{../tables/FEM/obesity_results_govexp.tex}
\hline
\input{../tables/FEM/obesity_results_totmd.tex}
\hline
\end{tabular}
\caption{Simulation results for obesity reduction scenario compared to status quo}
\label{tab:obesity_results}
\end{sidewaystable}

\begin{table}
\centering
\scalebox{0.8}{
\input{../tables/FEM/time_series_by_year.tex}
}
\caption{Assumptions for each calendar year}
\label{tab:time_series_by_year}
\end{table}

\begin{table}
\centering
\scalebox{0.65}{
\input{../tables/FEM/time_series_by_yob.tex}
}
\caption[Assumptions for each birth year]{Assumptions for each birth year. In years after 1960, all values are held constant at their 1960 levels.}
\label{tab:time_series_by_yob}
\end{table}