### 2019-11-22 analysis/techdoc/latex/FEM/acknowledgments.tex

\section{Acknowledgments}

The Future Elderly Model has been developed by a large team over the last decade.  Jay Bhattacharya, Eileen Crimmins, Christine Eibner, \'Etienne Gaudette, Geoff Joyce, Darius Lakdawalla,
Pierre-Carl Michaud, and Julie Zissimopoulos have all provided expert guidance.  Adam Gailey, Baoping Shang, and Igor Vaynman provided programming and analytic 
support during the first years of development at RAND.  In more recent years, the University of Southern California research programming team has supported 
model development.  These programmers include Patricia St. Clair, Laura Gascue, Henu Zhao, and Yuhui Zheng. Barbara Blaylock and Wendy Cheng have greatly aided 
model development while working as research assistants at USC.