### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation.tex

\section{Data sources used for estimation}
The Health and Retirement Study is the main data source for the model. We supplemented this data with 
merged Social Security covered earnings histories and data on health trends and health care costs 
coming from 3 major health surveys in the U.S. We describe these surveys below and the samples we 
selected for the analysis. We first list the variables used in the analysis. We then give details on 
the data sources.

\begin{center}
\begin{tabular}{l l}
\multicolumn{2}{c}{\textbf{Estimated Outcomes in Initial Conditions Model}}\\
 & \\
\textbf{Economic Outcomes}	& \textbf{Health Outcomes}\\
Employment & Hypertension\\
Earnings & Heart Disease\\
Wealth & Self-Reported Health\\
Defined Contribution Pension Wealth	& BMI Status\\
Pension Plan Type	& Smoking Status\\
AIME & Functional Status\\
Social Security Quarters of Coverage & \\
Health Insurance & 
\end{tabular}\end{center}~\\~\\

\begin{center}
\begin{tabular}{l l l}
\multicolumn{3}{c}{\textbf{Estimated Outcomes in/from Transition Model}}\\
 & & \\
\textbf{Economic Outcomes} & \textbf{Health Outcomes} & \textbf{Other Outcomes}\\
Employment & Death & Income Tax Revenue\\
Earnings & Heart & Social Security Revenue\\
Wealth & Stroke & Medicare Revenue\\
Demographics & Cancer & Medical Expenses\\
Health Insurance & Hypertension & Medicare Part A Expenses\\
Disability Insurance Claim & Diabetes & Medicare Part B Expenses\\
Defined Benefit Claim & Lung Disease & Medicare Part B Enrollment\\
SSI Claim & Nursing Home &  Medicare Part D Enrollment\\
Social Security Claim & BMI  & OASI Enrollment \\
 & Smoking Status & DI enrollment\\
 & ADL Limitations & SSI enrollment\\
 & IADL Limitations	& Medicaid Enrollment\\
 & & Medicaid Expenditures
\end{tabular}
\end{center}

\input{data_sources_estimation_hrs.tex}
\input{data_sources_estimation_ssearnings.tex}
\input{data_sources_estimation_nhis.tex}
\input{data_sources_estimation_meps.tex}
\input{data_sources_estimation_mcbs.tex}