### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_estimation_hrs.tex

\subsection{Health and Retirement Study}
The Health and Retirement Study (HRS) waves 2000-2008 are used to estimate the transition model. 
Interviews occur every two years. We use the dataset created by RAND (RAND HRS, version K) as our 
basis for the analysis. We use all cohorts in the analysis and consider sampling weights whenever 
appropriate. When appropriately weighted, the HRS in 2004 is representative of U.S. households where 
at least one member is at least 51. The HRS is also used as the host data for the simulation (pop 51+ in 2004) and for new cohorts (aged 51 and 52 in 2004).

The HRS adds new cohorts every six years. Until recently, the latest available cohort had been added 
in 2004, which is why that is the FEM's base year. The FEM is currently being updated to use the 
newly released 2010 data.