### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline.tex

\section{Data sources for trends and baseline scenario}
Two types of trends need to be projected in the model. First, we need to project trends in the 
incoming cohorts (the future new age 51/52 individuals). This includes trends in health and economic 
outcomes. Second, we need to project excess aggregate growth in real income and excess growth in 
health spending.

\input{data_sources_trends_and_baseline_entering_cohorts.tex}
\input{data_sources_trends_and_baseline_other_projections.tex}
\input{data_sources_trends_and_baseline_demographic_projections.tex}