### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline_demographic_projections.tex

\subsection{Demographic adjustments}
We make two adjustments to the weighting in the HRS to match population counts. Since we deleted some 
cases from the data and only considered the set of respondents with matched Social Security records, 
this takes account of selectivity based on these characteristics. First, we post-stratify the HRS sample 
by 5 year age groups, gender and race and rebalance weights using the Census Bureau 2000-2010 Intercensal 
Population Estimates. We do this for both the host data set and the new cohorts. We scale the weights for 
future new cohorts using 2012 National Population Projections based on race and gender. Second, we post-
stratify the HRS sample of deaths between the 2002 and 2004 interview waves by 5 year age groups, gender 
and race and rebalance weights based on the Human Mortality Database.

Once the simulation begins, trends in migration and mortality are applied.  We use net migration from
the SSA Trustees report intermediate cost scenario.  Seperate mortality rate adjustment factors are 
defined for the under and over 65 age groups based on the mortality projections from the 2013 SSA
Trustees report.  The SSA projections are interpolated through 2090, then extended using GLM with log 
link through 2150.