### 2019-11-22 analysis/techdoc/latex/FEM/data_sources_trends_and_baseline_other_projections.tex

\subsection{Data for other projections}
We make two assumptions relating to real growth in wages and medical costs. Firstly, as is done in the 2009 
Social Security Trustees report intermediate cost scenario, we assume a long term real increase in wages 
(earnings) of 1.1\% per year. Next, following the Centers for Medicare and Medicaid Services, we assume 
excess real growth in medical costs (that is additional cost growth to GDP growth), as 1.5\% in 2004, 
reducing linearly to 1\% in 2033, .4\% in 2053, and -.2\% in 2083. We also include the Affordable Care 
Act cost growth targets as an optional cap on medical cost growth. Baseline medical spending figures 
presented assume those targets are met. GDP growth in the near term (through 2019) is based on CBO 
projections, with the OASDI Trustees assumption of 2\% yearly afterwards.