### 2019-11-22 analysis/techdoc/latex/FEM/estimation_goodness-of-fit.tex

\subsection{Goodness-of-fit}
To judge the goodness-of-fit of the model, we estimated parameters on the 1998-2008 estimation sample and simulated outcomes of 1998 HRS 
respondents up to 2008. We then compared simulated and actual outcomes in 1998, 2004 and 2008. Table \ref{tab:validation_waves_4_7_9} presents the results. Some 
differences exist but in general the fit is satisfactory.

%\input{../tables/FEM/validation_waves_4_7_9.tex}
%\hline
%\caption{Simulated and actual outcomes in 1998, 2004, and 2008}
%\label{tab:validation_waves_4_7_9}
%\end{longtable}