### 2019-11-22 analysis/techdoc/latex/FEM/estimation_transition_model.tex

\subsection{Transition model}
\label{sec:estimation_transition_model} 
We consider a large set of outcomes for which we model transitions. Table \ref{tab:transitioned_outcomes} gives the set of outcomes considered for the transition model 
along with descriptive statistics and the population at risk when estimating the relationships.

Since we have a stock sample from the age 51+ population, each respondent goes through an individual-specific series of intervals. Hence, 
we have an unbalanced panel over the age range starting from 51 years old. Denote by $j_{i0}$ the first age at which respondent $i$ is 
observed and $j_{iT_i}$ the last age when he is observed. Hence we observe outcomes at ages $j_i = j_{i0},\ldots,j_{iT_i}$.

We first start with discrete outcomes which are absorbing states (e.g. disease 
diagnostic, mortality, benefit claiming). Record as $h_{i,j_i,m}=1$ if the 
individual outcome $m$ has occurred as of age $j_i$. We assume the 
individual-specific component of the hazard can be decomposed in a time 
invariant and variant part. The time invariant part is composed of the effect of 
observed characteristics $x_i$ that are constant over the entire life course and 
initial conditions $h_{i,j_0,-m}$ (outcomes other than the outcome 
$m$) that are determined before the first age in which each individual is observed
\footnote{Section \ref{sec:model_development_transition_model} explains why the $h_{i,j_0,-m}$ terms are included.}. 
The time-varying part is the effect of previously diagnosed outcomes $h_{i,j_i-1,-m}$, 
on the hazard for $m$.\footnote{With some abuse of notation, $j_i-1$ denotes 
the previous age at which the respondent was observed.}  We assume an index of 
the form  $z_{m,j_i} = x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + h_{i,j_0,-m}\psi_m$. Hence, the 
latent component of the hazard is modeled as 
\begin{equation}
h^*_{i,j_i,m}= x_i\beta_m + h_{i,j_i-1,-m} \gamma_m + h_{i,j_0,-m}\psi_m + a_{m,j_i} + \varepsilon_{i,j_i,m},
\label{eqn:transition_hzd_latent}
\end{equation}
\[
m = 1,\ldots,M_0 \mbox{, } j_i = j_{i0},\ldots,j_{i,T_i} \mbox{, } i=1,\ldots,N
\]
The term $\varepsilon_{i,j_i,m}$ is a time-varying shock specific to age $j_i$. 
We assume that this last shock is normally distributed and uncorrelated across 
diseases.  We approximate $a_{m,j_i}$ with an age spline. After several specification checks, 
knots at age 65 and 75 appear to provide the best fit. This simplification is made 
for computational reasons since the joint estimation with unrestricted age fixed 
effects for each condition would imply a large number of parameters.  The 
absorbing outcome, conditional on being at risk, is defined as 
\[
h_{i,j_i,m} = \max\{I(h^*_{i,j_i,m} > 0), h_{i,j_i-1,m}\}
%\label{eqn:transition_outcome}
\]
The occurrence of mortality censors observation of other 
outcomes in a current year. Mortality is recorded from exit interviews.

A number of restrictions are placed on the way feedback is allowed in the model.  
Table \ref{tab:transition_restrictions} documents restrictions placed on the 
transition model. We also include a set of other controls. A list of such controls 
is given in Table \ref{tab:controlvar_descs} along with descriptive statistics.

We have three other types of outcomes:
\begin{enumerate}
\item First, we have binary outcomes which are not an absorbing state, such as living in a nursing home. We specify latent 
indices as in (\ref{eqn:transition_hzd_latent}) for these outcomes as well but where 
the lag dependent outcome also appears as a right-hand side variable. This allows for state-dependence.

\item Second, we have ordered outcomes. These outcomes are also modeled as in (\ref{eqn:transition_hzd_latent}) recognizing the observation rule is 
a function of unknown thresholds $\varsigma_m$. Similarly to binary outcomes, we allow for state-dependence by including the lagged outcome 
on the right-hand side.

\item The third type of outcomes we consider are censored outcomes, earnings and financial wealth. Earnings are only observed when individuals 
work. For wealth, there are a non-negligible number of observations with zero and negative wealth. For these, we consider two part models 
where the latent variable is specified as in (\ref{eqn:transition_hzd_latent}) but model probabilities only when censoring does not occur. In 
total, we have $M$ outcomes.
\end{enumerate}

The parameters 
$\mathbf{\theta}_1 = \left(\left\{\beta_m, \gamma_m, \psi_m, \varsigma_m\right\}_{m=1}^M, \right)$, 
can be estimated by maximum likelihood. Given the normality distribution assumption on the 
time-varying unobservable, the joint probability of all time-intervals until failure, right-censoring 
or death conditional on the initial conditions $h_{i,j_0,-m}$ is the product of 
normal univariate probabilities. Since these sequences, conditional on initial 
conditions, are also independent across diseases, the joint 
probability over all disease-specific sequences is simply the product of 
those probabilities.

For a given respondent observed from initial age $j_{i0}$ to a last age $j_{T_i}$, the probability of the observed health history is 
(omitting the conditioning on covariates for notational simplicity)
\[
	l^{-0}_i(\mathbf{\theta}; h_{i,j_{i0}}) = \left[\prod_{m=1}^{M-1} \prod_{j=j_{i1}}^{j_{T_i}} P_{ij,m}(\mathbf{\theta})^{(1-h_{ij-1,m})(1-h_{ij,M})} \right] \times \left[\prod_{j=j_{i1}}^{j_{T_i}} P_{ij,M}(\mathbf{\theta}) \right]
\]
We use the ${-0}$ superscript to make explicit the conditioning on $\mathbf{h}_{i,j_{i0}} = (h_{i,j_{i0},0},\ldots,h_{i,j_{i0},M})'$. We have limited information on outcomes prior to this age. 
The likelihood is a product of $M$ terms with the $m$th term containing only 
$(\beta_m, \gamma_m, \psi_m, \varsigma_m)$.  This allows the estimation
to be done seperately for each outcome.

\subsubsection{Inverse Hyperbolic Sine Transformation}
One problem fitting the wealth and earnings distribution is that they have a long right tail and wealth has some negative values. We use a 
generalization of the inverse hyperbolic sine transform (IHT) presented in \citet{mackinnon1990transforming}. First denote the variable of 
interest $y$. The hyperbolic sine transform is 
\begin{equation}
y = \sinh(x) = \frac{\exp(x) - \exp(-x)}{2}
\label{eqn:sinh_y}
\end{equation}
The inverse of the hyperbolic sine transform is
\[
x = \sinh^{-1}(y) = h(y) = \log(y + (1+y^2)^{1/2})
\]
Consider the inverse transformation. We can generalize such transformation, first allowing for a 
shape parameter $\theta$,
\begin{equation}
r(y) = h(\theta y)/\theta
\label{eqn:generalized_ihs_shape}
\end{equation}
Such that we can specify the regression model as
\begin{equation}
r(y) = x\beta + \varepsilon, \varepsilon \sim \mathrm{N}(0, \sigma^2)
\label{eqn:ihs_regression_model}
\end{equation}
A further generalization is to introduce a location parameter $\omega$ such that the new 
transformation becomes
\begin{equation}
g(y) = \frac{h(\theta(y+\omega)) - h(\theta\omega)}{\theta h'(\theta \omega)}
\label{eqn:geralized_ihs_loc_scale}
\end{equation}
where $h'(a) = (1+a^2)^{-1/2}$.

We specify (\ref{eqn:ihs_regression_model}) in terms of the transformation $g$. The shape parameters 
can be estimated from the concentrated likelihood for $\theta, \omega$. We can then 
retrieve $\beta, \sigma$ by standard OLS.

Upon estimation, we can simulate 
\[
\tilde{g} = x \hat{\beta} + \sigma \tilde{\eta}
\]
where $\eta$ is a standard normal draw. Given this draw, we can retransform using 
(\ref{eqn:geralized_ihs_loc_scale}) and (\ref{eqn:sinh_y})
\begin{align*}
&h(\theta(y+\omega)) = \theta h'(\theta\omega)\tilde{g} + h(\theta\omega)\\
&\tilde{y} = \frac{\sinh\left[\theta h'(\theta\omega)\tilde{g} + h(\theta\omega)\right]-\theta\omega}{\theta}
\end{align*}

% \todo link to transition_estimates.xls on box -- make link unique to the version of the appendix in make process
%The estimates table\footnote{\input{estimates_table_link.tex}} gives parameter estimates for the transition models.

%\newpage
%\input{../tables/FEM/health_transition_est.tex}
%\caption{Coefficient estimates and $t$-statistics for mortality and chronic condition transitions}
%\label{tab:health_transition_est}
%\end{longtable}
%
%\newpage
%\input{../tables/FEM/econ_transition_est.tex}
%\caption{Coefficient estimates and $t$-statistics for economic outcome transitions}
%\label{tab:econ_transition_est}
%\end{longtable}
%
%\newpage
%\input{../tables/FEM/BMI_transition_est.tex}
%\caption{Coefficient estimates and $t$-statistics for BMI outcome transition}
%\label{tab:BMI_transition_est}
%\end{longtable}
%
%\newpage
%\input{../tables/FEM/ordered_transition_est.tex}
%\caption{Coefficient estimates and $t$-statistics for ordered outcome transitions}
%\label{tab:ordered_transition_est}
%\end{longtable}
%
%\newpage
%\input{../tables/FEM/IHS_transition_est.tex}
%\caption{Coefficient estimates and $t$-statistics for transitioning outcomes with shape-preserving Inverse Hyperbolic Sine transformation}
%\label{tab:IHS_transition_est}
%\end{longtable}