### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures.tex

\section{Government revenues and expenditures}
This gives a limited overview of how revenues and expenditures of the government are computed. These 
functions are based on 2004 rules, but we include predicted changes in program rules such changes 
based on year of birth (e.g. Normal retirement age).

We cover the following revenues and expenditures:\\

\begin{center}
\begin{tabular}{l l}
\textbf{Revenues} & \textbf{Expenditures}\\
Federal Income Tax & Social Security Retirement benefits\\
State and City Income Taxes & Social Security Disability benefits\\
Social Security Payroll Tax & Supplementary Security Income (SSI)\\
Medicare Payroll Tax & Medical Care Costs\\
Property Tax & Medicaid\\
 & Medicare (parts A, B, and D)
\end{tabular}
\end{center}

\input{govt_revenue_and_expenditures_SSben.tex}
\input{govt_revenue_and_expenditures_DIben.tex}
\input{govt_revenue_and_expenditures_SSIben.tex}
\input{govt_revenue_and_expenditures_medcost_estimation.tex}
\input{govt_revenue_and_expenditures_taxes.tex}