### 2019-11-22 analysis/techdoc/latex/FEM/govt_revenue_and_expenditures_SSIben.tex

\subsection{Supplemental Security Income benefits}
Self-reported receipt of supplemental security income (SSI) in the HRS provides estimates of the proportion of people receiving SSI under 
what administrative data would suggest. To correct for this bias, we link the HRS with administrative data from the social security 
administration identifying those receiving SSI. In the linked administrative data, 3.96\% of the population receives supplementary security 
income, while only 2.79\% of the sample reports social security income. We therefore estimate a probit of receiving SSI as a function of 
self-reporting social security income, as well as demographic, health, and wealth. 
%Model estimates can be found in the estimates table \footnote{\input{estimates_table_link.tex}}.

The benefit amount is taken from the average monthly benefits found in the 2004 Social Security Annual Statistical Supplement. We assign 
monthly benefit of \$450 for person aged 51 to 64, and \$350 for persons aged 65 and older.