### 2019-11-22 analysis/techdoc/latex/FEM/implementation.tex

\section{Implementation}
The FEM is implemented in multiple parts. Estimation of the transition and cross 
sectional models is performed in Stata.
The incoming cohort model is estimated in Stata using the CMP package \citep{statacmp2011}.
The simulation is implemented in C++ to increase speed.

To match the two year structure of the Health and Retirement Study (HRS) data used to estimate the transition models, the FEM simulation 
proceeds in two year increments. The end of each two year step is designed to occur on July 1st to allow for easier matching to population 
forecasts from Social Security. A simulation of the FEM proceeds by first loading a population representative of the age 51+ US population 
in 2004, generated from HRS. In two year increments, the FEM applies the transition models for mortality, health, working, wealth, 
earnings, and benefit claiming with Monte Carlo decisions to calculate the new states of the population. The population is also adjusted by 
immigration forecasts from the US Census Department, stratified by race and age. If incoming cohorts are being used, the new 51/52 year 
olds are added to the population. The number of new 51/52 year olds added is consistent with estimates from the Census, stratified by race. 
Once the new states have been determined and new 51/52 year olds added, the cross sectional models for medical costs, and calculations for 
government expenditures and revenues are performed. Summary variables are then computed. Computation of medical costs includes the persons 
that died to account for end of life costs. Other computations, such as Social Security benefits and government tax revenues, are 
restricted to persons alive at the end of each two year interval. To eliminate uncertainty due to the Monte Carlo decision rules, the 
simulation is performed multiple times (typically 100), and the mean of each summary variable is calculated across repetitions.

FEM simulation takes as inputs assumptions regarding growth in the national wage index, normal retirement age, real medical cost growth, 
interest rates, cost of living adjustments, the consumer price index, significant gainful activity, and deferred retirement credit. The 
default assumptions are taken from the 2010 Social Security Intermediate scenario, adjusted for no price increases after 2010. Therefore 
simulation results are in real 2009 dollars.  Table \ref{tab:time_series_by_year} shows the assumptions for each calendar year and Table 
\ref{tab:time_series_by_yob} shows assumptions for each birth year.

Different simulation scenarios are implemented by changing any of the following components: incoming cohort model, transition models, 
interventions that adjust the probabilities of specific transition, and changes to assumptions on future economic conditions.

\input{./implementation_interventions.tex}