### 2019-11-22 analysis/techdoc/latex/FEM/validation.tex

\section{Validation}
We perform three validation exercises:

\begin{enumerate}
\item Cross-validation
\item External validation
\item External corroboration
\end{enumerate}

Cross-validation is a test of the simulation’s internal validity that compares simulated outcomes to actual outcomes, external validation compares model forecasts 
with actual outcomes from other data sources, and external corroboration compares model forecasts to others’ forecasts.

\input{validation_crossvalidation.tex}
\input{validation_external_validation.tex}
\input{validation_external_corroboration.tex}