### 2019-11-22 analysis/volunteer-swv.tex

\documentclass{article}
\usepackage{savetrees}
\usepackage{hyperref}
\usepackage[section]{placeins}
\usepackage{datatool}

\title{Analysis of Volunteering in the HRS}

\begin{document}
\maketitle
\listoftables
\listoffigures

\begin{Rcode}
library(foreign)
library(ggplot2)
library(survey)
hrs <- read.dta('../input_data/hrs_analytic_recoded.dta')
hrs <- subset(hrs, !is.na(wtresp) & wtresp > 0 & !died & wave %in% 3:7, select=!'r8' %in% names(hrs) | !'r9' %in% names(hrs))
hrs[,'orgnonzero'] <- hrs[,'orghours'] > 0
hrs[,'friendnonzero'] <- hrs[,'friendhours'] > 0
wave3.age <- weighted.mean(hrs[hrs[,'wave'] == 3, 'age'], hrs[hrs[,'wave'] == 3, 'wtresp'])
rhs <- 'age + black + hispan + male + hsless + college + work + hearte + stroke + cancre + hibpe + diabe + adl12 + adl3 + iadl1 + widowed + diclaim + ssclaim + dbclaim + wave'
dorgnz <- svydesign( id=~ 1, data=hrs, weights= ~ wtresp)
dfriendnz <- svydesign( id=~ 1, data=hrs, weights= ~ wtresp)
org.ever.glm <- svyglm(paste('orgnonzero', rhs, sep=' ~ ') , family=quasibinomial(link='logit'), design=dorgnz)
print(summary(org.ever.glm))
friend.ever.glm <- svyglm(paste('friendnonzero', rhs, sep=' ~ '), family=quasibinomial(link='logit'), design=dfriendnz)
print(summary(friend.ever.glm))
orgnz <- subset(hrs, orgnonzero)
frdnz <- subset(hrs, friendnonzero)
dorg <- svydesign( id=~1, data=orgnz, weights=~wtresp)
dfrd <- svydesign( id=~1, data=frdnz, weights=~wtresp)
org.lm <- svyglm(paste('orghours', rhs, sep='~'), design=dorg)
print(summary(org.lm))
frd.lm <- svyglm(paste('friendhours', rhs, sep='~'), design=dfrd)
print(summary(frd.lm))
hrs[,'orgnonzero.p'] <- predict.glm(org.ever.glm, newdata=hrs, type='response', na.action=na.pass)
hrs[,'friendnonzero.p'] <- predict.glm(friend.ever.glm, newdata=hrs, type='response', na.action=na.pass)
hrs[,'orghours.p'] <- predict.glm(org.lm, newdata=hrs, type='response', na.action=na.pass) * hrs[,'orgnonzero.p']
hrs[,'friendhours.p'] <- predict.glm(frd.lm, newdata=hrs, type='response', na.action=na.pass) * hrs[,'friendnonzero.p']
hrs2 <- hrs
hrs2[,'age'] <- wave3.age
hrs[,'orgnonzero.a'] <- predict.glm(org.ever.glm, newdata=hrs2, type='response', na.action=na.pass)
hrs[,'friendnonzero.a'] <- predict.glm(friend.ever.glm, newdata=hrs2, type='response', na.action=na.pass)
hrs[,'orghours.a'] <- predict.glm(org.lm, newdata=hrs2, type='response', na.action=na.pass) * hrs[,'orgnonzero.a']
hrs[,'friendhours.a'] <- predict.glm(frd.lm, newdata=hrs2, type='response', na.action=na.pass) * hrs[,'friendnonzero.a']
rm(hrs2)
gc()
\end{Rcode}

\begin{Rcode}{savefig,label=orgs}
hrs.s <- subset(hrs, !is.na(orghours) & orghours <= 200)
ggplot(hrs.s, aes(x=orghours, weight=wtresp/1e6)) + geom_histogram() + facet_wrap( ~ wave) + labs(y='Millions of People')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{orgs}
\caption{Weighted Histograms of Organizational Volunteer Hours}
\end{figure}

\begin{Rcode}{savefig,label=friends}
hrs.s <- subset(hrs, !is.na(friendhours) & friendhours <= 200)
ggplot(hrs.s, aes(x=friendhours, weight=wtresp/1e6)) + geom_histogram() + facet_wrap( ~ wave) + labs(y='Millions of People')
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{friends}
\caption{Weighted Histograms of Friends and Family Volunteer Hours}
\end{figure}

\begin{Rcode}{}
hrs[,'orgweight'] <- ifelse(is.na(hrs[,'orghours']), NA, hrs[,'wtresp'])
hrs[,'friendweight'] <- ifelse(is.na(hrs[,'friendhours']), NA, hrs[,'wtresp'])

hrs.waves <- split(hrs, hrs[,'wave'])

orgnonzero <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero'], x[,'orgweight'], na.rm=T))
orgnonzero.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero.p'], x[,'wtresp'], na.rm=T))
orgnonzero.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orgnonzero.a'], x[,'wtresp'], na.rm=T))
orgmean <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours'], x[,'orgweight'], na.rm=T))
orgmean.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours.p'], x[,'wtresp'], na.rm=T))
orgmean.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'orghours.a'], x[,'wtresp'], na.rm=T))
orgwtfactor <- sapply(hrs.waves, FUN=function(x) sum(x[,'wtresp'], na.rm=T)/sum(x[,'orgweight'], na.rm=T))
orgtotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours'] * x[,'orgweight'], na.rm=T)/1e9) * orgwtfactor
orgtotal.p <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours.p'] * x[,'wtresp'], na.rm=T)/1e9)
orgtotal.a <- sapply(hrs.waves, FUN=function(x) sum(x[,'orghours.a'] * x[,'wtresp'], na.rm=T)/1e9)

friendnonzero <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero'], x[,'friendweight'], na.rm=T))
friendnonzero.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero.p'], x[,'wtresp'], na.rm=T))
friendnonzero.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendnonzero.a'], x[,'wtresp'], na.rm=T))
friendmean <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours'], x[,'friendweight'], na.rm=T))
friendmean.p <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours.p'], x[,'wtresp'], na.rm=T))
friendmean.a <- sapply(hrs.waves, FUN=function(x) weighted.mean(x[,'friendhours.a'], x[,'wtresp'], na.rm=T))
friendwtfactor <- sapply(hrs.waves, FUN=function(x) sum(x[,'wtresp'], na.rm=T)/sum(x[,'friendweight'], na.rm=T))
friendtotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours'] * x[,'friendweight'], na.rm=T)/1e9) * friendwtfactor
friendtotal.p <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours.p'] * x[,'wtresp'], na.rm=T)/1e9)
friendtotal.a <- sapply(hrs.waves, FUN=function(x) sum(x[,'friendhours.a'] * x[,'wtresp'], na.rm=T)/1e9)

iearntotal <- sapply(hrs.waves, FUN=function(x) sum(x[,'iearn'] * x[,'wtresp'], na.rm=T)/1e9)

dat <- data.frame(year=c(1996, 1998, 2000, 2002, 2004), orgtotal, orgtotal.p, orgtotal.a, friendtotal, friendtotal.p, friendtotal.a, iearntotal, orgnonzero, friendnonzero, orgnonzero.p, friendnonzero.p, orgnonzero.a, friendnonzero.a)
\end{Rcode}

\begin{Rcode}{savefig,label=value}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal*20.85, colour='Volunteering for Organization')) + 
geom_line(aes(y=friendtotal*11.22, colour='Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{Rcode}{savefig,label=valuea}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal.a*20.85, colour='Age-Adjusted Volunteering for Organization')) + 
geom_line(aes(y=friendtotal.a*11.22, colour='Age-Adjusted Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{Rcode}{savefig,label=valuep}

ggplot(dat, aes(x=year)) + 
geom_line(aes(y=orgtotal.p*20.85, colour='Predicted Volunteering for Organization')) + 
geom_line(aes(y=friendtotal.p*11.22, colour='Predicted Volunteering for Friends')) + 
labs(x='Year', y='Billions of 2009 Dollars', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,100))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{value}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{valuea}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{valuep}
\caption{Total Value of Volunteering in the 51+ Population}
\end{figure}

\begin{Rcode}{savefig,label=fraction}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero, colour='Fraction Volunteering for Organizations')) +
geom_line(aes(y=friendnonzero, colour='Fraction Volunteering for Friends, etc')) + 
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{Rcode}{savefig,label=fractiona}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero.a, colour='Age-Adjusted Volunteering for Organizations')) +
geom_line(aes(y=friendnonzero.a, colour='Age-Adjusted Volunteering for Friends')) +
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{Rcode}{savefig,label=fractionp}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgnonzero.p, colour='Predicted Fraction Volunteering for Organizations')) + 
geom_line(aes(y=friendnonzero.p, colour='Predicted Fraction Volunteering for Friends')) +
labs(x='Year', y='Portion of Population', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,1))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{fraction}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{fractiona}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{fractionp}
\caption{Fraction of 51+ Population who Volunteered}
\end{figure}

\begin{Rcode}{savefig,label=means}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean, colour='Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean, colour='Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{Rcode}{savefig,label=meansa}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean.a, colour='Age-Adjusted Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean.a, colour='Age-Adjusted Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{Rcode}{savefig,label=meansp}
ggplot(dat, aes(x=year)) +
geom_line(aes(y=orgmean.p, colour='Predicted Hours Volunteered for Organizations')) +
geom_line(aes(y=friendmean.p, colour='Predicted Hours Volunteered for Friends, etc')) + 
labs(x='Year', y='Number of Hours', colour='Type of Work') + opts(legend.position='bottom') +
scale_y_continuous(limits=c(0,70))
\end{Rcode}

\begin{figure}[ht]
\centering
\recallfig{means}
\caption{Mean number of hours}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{meansa}
\caption{Mean number of hours}
\end{figure}

\begin{figure}[ht]
\centering
\recallfig{meansp}
\caption{Mean number of hours}
\end{figure}

\end{document}