### 2020-07-21 FEM_R/pubHealthInterventionVis_Notebook.Rmd

---
title: "Public Health Intervention Visualisation"
output: html_notebook
---

This notebook will keep all visualisations for the English Future Elderly Model public health intervention scenarios.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(haven)
require(dplyr)
require(ggplot2)
```

# COMMIT

The commit smoking interventions are inspired by this paper. <!-- TODO: Add link to COMMIT paper here -->

The intervention resulted in a 3% increase in the smoking quit rate for light/moderate smokers (anyone smoking less than 25 cigs/day).

```{r}
 # First read in the baseline cohort data file and the commit scenario outputs
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
commit3 <- read_dta('output/COMMIT_cSmoken3/COMMIT_cSmoken3_summary.dta')

# As these are all cohort datasets we can calculate their age by (year - 2012 + 50) 
cohort$age <- (cohort$year - 2012) + 50
commit3$age <- (cohort$year - 2012) + 50

# Now calculate survival rates for all 3 (normalise pop_count to inital value)
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
commit3$survival <- commit3$m_endpop_all / commit3$m_endpop_all[1]

# Plot survival curve
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.5) +
        geom_line(data=commit3, col='red', size=0.5)

# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=commit3, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
```

# Smoke Stop Cohort Interventions

```{r}
# First read in the baseline cohort data file and the commit scenario outputs
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
x169 <- read_dta('output/Smoke_Stop_Cohort169/Smoke_Stop_Cohort169_summary.dta')

# As these are all cohort datasets we can calculate their age by (year - 2012 + 50)
cohort$age <- (cohort$year - 2012) + 50
x169$age <- (cohort$year - 2012) + 50

# Now calculate survival rates for all 3 (normalise pop_count to inital value)
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
x169$survival <- x169$m_endpop_all / x169$m_endpop_all[1]

# Plot survival curve
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.5) +
        geom_line(data=x169, col='red', size=0.5)

# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=x169, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')
```

# Smoke Stop Population Interventions

```{r}
# First read in the baseline cohort data file and the commit scenario outputs
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
p169 <- read_dta('output/Smoke_Stop_Pop169/Smoke_Stop_Pop169_summary.dta')

# No need to calculate year as this is population scenario (replenishing)
# Also means can't calculate survival rates

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=p169, size=0.3, col='red') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')
```

# Custom Smoke Stop Interventions (Cohort)

This is the custom intervention written in C++ - SmokeStopIntervention.cpp

```{r}

# Read in data
cohort <- read_dta('output/ELSA_cohort/ELSA_cohort_summary.dta')
SSPC <- read_dta('output/SmokeStopInterventionC/SmokeStopInterventionC_summary.dta')

# Generate age var
cohort$age <- (cohort$year - 2012) + 50
SSPC$age <- (cohort$year - 2012) + 50

# Survival curve
cohort$survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
SSPC$survival <- SSPC$m_endpop_all / SSPC$m_endpop_all[1]

# Now plot all survival curves together
ggplot(cohort, aes(x=age, y=survival)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Cohort Survival Over Time', y='Survival', x='Age')
# Smoking variables and chronic diseases
ggplot(cohort, aes(x=age, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Smoking by Age', y='Prevalence of Smoking', x='Age')
ggplot(cohort, aes(x=age, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease by Age', y='Prevalence of Lung Disease', x='Age')
ggplot(cohort, aes(x=age, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPC, size=0.3, col='red') +
        labs(title='Prevalence of Cancer by Age', y='Prevalence of Cancer', x='Age')

```

# Custom Smoke Stop Interventions (Population)

```{r}

# Load data
pop <- read_dta('output/ELSA_Baseline/ELSA_Baseline_summary.dta')
SSPP <- read_dta('output/SmokeStopInterventionP/SmokeStopInterventionP_summary.dta')

# Plot Average age at death
ggplot(pop, aes(x=year, y=d_age_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Average Age at Death over Time', y='Average age at death', x='Year')

# Smoking variables and chronic diseases
ggplot(pop, aes(x=year, y=n_smoke_stop_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Smoke Stop', y='# Stopped Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_smoken_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Smoking over Time', y='Prevalence of Smoking', x='Year')
ggplot(pop, aes(x=year, y=p_lunge_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Lung Disease over Time', y='Prevalence of Lung Disease', x='Year')
ggplot(pop, aes(x=year, y=p_cancre_all)) +
        geom_line(col='black', size=0.3) +
        geom_line(data=SSPP, size=0.3, col='red') +
        labs(title='Prevalence of Cancer over Time', y='Prevalence of Cancer', x='Year')

```