### 2020-08-12 FEM_R/ELSA_US_comparison.Rmd
Lots of changes to R scripts analysing outputs. Of note, started an R notebook for imputing missing data in ELSA before running through the model

---
title: "Comparing Understanding Society with ELSA"
output: html_notebook
---

This document exists to compare the distributions of key variables (mainly risk behaviours: smoking, drinking, exercise (and obesity)) between Understanding Society (US) and the English Longitudinal Survey of Ageing (ELSA).

# Setup

```{r "setup", include=FALSE}

# Set up the environment
US_dir <- "/home/luke/Documents/E_FEM_clean/Understanding_Society/UKDA-6614-stata/stata/stata11_se/ukhls_w2/"
ELSA_dir <- "/home/luke/Documents/E_FEM/UKDA-5050-stata/stata/stata11_se/"

require(haven)
require(dplyr)
require(ggplot2)

```

# Smoking Distributions by Age by Sex

```{r}

```