### 2020-08-12 FEM_R/ELSA_imputation.Rmd

---
title: "Multiple Imputation of Missing Data in ELSA"
output: html_notebook
---

This document will run through the steps taken to impute large amounts of missing data in ELSA.

Missing data in ELSA is a significant problem to running the FEM, as a number of key predictive variables have a high degree of missingness and/or are not missing at random. For example, BMI data is missing for every odd wave, and the even waves are missing a considerable proportion of that data.

First, the extent of missing information will be assessed, then a few methods for imputing these missing values will be trialled.

```{r "setup", include=FALSE}

```