### 2020-11-16 FEM_R/IJM_paper_vis_NEAT.Rmd
Copied visualisation script into FEM_R from the IJM paper directory. Added chunk to visualise cross-validation 2 output

---
title: "IJM Paper Visualisations NEAT"
output: html_notebook
---

This notebook will hold the finished visualisation functions written in IJM_paper1_vis_WIP.Rmd.

This is just so we can come to a single notebook and rerun ALL visualisations that will go into the IJM FEM paper.

Start with setup.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(plyr)
require(data.table)

```

# Validation Stuff

## Cross-validation 2

In this scenario, transition models were estimated from waves 1-4, and the stock population was based on wave 5, with a view to simulating from wave 5-8 (2010-2016). We want to check whether there is a big difference as the model takes over from ELSA data from 2008-2010, in terms of chronic disease prevalence etc.

```{r}

vis_CV2_prev <- function(long, CV2, path) {

  # Keep only waves 1 - 4
  longC <- subset(long, wave < 5)
  longC <- as.data.table(longC)

  # Replace all NA with 0
  longC[is.na(longC)] <- 0

  # Have to drop r*clust vars as they contain strings
  longC <- select(longC, -contains('clust'))

  # Now get the weighted mean by wave of every column based on cwtresp
  l2 <- longC[ , lapply(.SD,weighted.mean,w=cwtresp), by = list(wave)]

  # ELSA_long is now processed to the point where we can start extracting and combining variables
  # Better start by adding year to l2 and reordering the columns
  l2$year <- seq(from=2002, to = 2008, by = 2)
  setcolorder(l2, c('year', 'wave'))

  # Now lets extract chronic disease variables from both dataframes and do a row combine
  # Start with cancre, diabe, hearte
  cd_l2 <- select(l2, c(year, cancre, diabe, hearte, lunge, smoken, bmi, work, anyadl, anyiadl)) #, adl1, adl2, adl3p))
  cd_CV2 <- select(CV2, c(year, p_cancre_all, p_diabe_all, 
                            p_hearte_all, p_lunge_all, p_smoken_all, 
                            a_bmi_all, p_work_all,  p_anyadl_all, p_anyiadl_all))
                            #p_adl1_all, p_adl2_all, p_adl3p_all))

  CV2_rename <- cd_CV2 %>% dplyr::rename('cancre' = p_cancre_all, 
                                        'diabe' = p_diabe_all, 
                                        'hearte' = p_hearte_all,
                                        'lunge' = p_lunge_all,
                                        'smoken' = p_smoken_all,
                                        'bmi' = a_bmi_all,
                                        'work' = p_work_all,
                                        'anyadl' = p_anyadl_all,
                                        'anyiadl' = p_anyiadl_all)
                                        #'adl1' = p_adl1_all,
                                        #'adl2' = p_adl2_all,
                                        #'adl3p' = p_adl3p_all)

  combine <- rbind(cd_l2, CV2_rename)

  melted <- melt(combine, id.var = 'year', value.name = 'Prevalence')

  ggplot(data = melted, aes(x = year, y = Prevalence, color = variable)) +
          geom_line() +
          facet_wrap(. ~ variable, scales = "free")
          #theme(panel.spacing.x = unit(1.5, 'lines')) +
          #labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')
}

# Load in whole stock pop, plus the CV2 summary output file
long <- read_dta('../input_data/ELSA_long.dta')
CV2 <- read_dta('ELSA_CrossValidation2/ELSA_CrossValidation2_summary.dta')

vis_CV2_prev(long, CV2, path = 'path')

```

# ReduceBMI (NEW SCENARIO I - FULL POPULATION)

## Baseline Initial Prevalence

```{r}
### visInitBasePrev
# Inputs:   base
#           path
#
# Output:   bar - 
#
# This function... 
visInitBasePrev <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                filter(year == 2012) %>%
                melt(id.var = 'year', value.name = 'Prevalence') %>%
                subset(select = -year)

  bar <- ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
          geom_bar(stat = 'identity') +
          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'init_baseline_prevalence.png', 
         plot = bar,
         path = path,
         width = 7,
         height = 5)

  bar
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visInitBasePrev(baseline)

```

## Baseline Projections

```{r}
### visProjectedBaselineCD
# Inputs:   base - summary output file from the ELSA_Baseline scenario
#           path - path to the directory where the resulting plot will be saved
#
# Output:   tiled - A 3 column tiled plot with the projected prevalence of Cancer, Diabetes, and Heart Disease on the same x axis, with year on the y
#
# This function visualises the prevalence of Cancer, Diabetes, and Heart Disease in the Baseline projection of the FEM
visProjectedBaselineCD <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                melt(id.var='year', value.name = 'Prevalence')

  # Create the tiled plot of disease prevalence
  tiled <- ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
            geom_line() +
            facet_grid(. ~ variable) +
            theme(panel.spacing.x = unit(1.5, 'lines')) +
            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'baseline_prevalence.png', 
         plot = tiled,
         path = path,
         width = 10,
         height = 5)

  # Show plot
  tiled
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBaselineCD(baseline)

```

## ChronDisease Prevalence Comparison

```{r}

visCDComparison <- function(base, int, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # For file naming and stuff
  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  }

  var <- paste0('p_', cd, '_all')

  selBase <- baseline %>%
              select(c('year', all_of(var))) %>%
              rename('Baseline' = var)

  selInt <- int %>%
              select(c('year', all_of(var))) %>%
              rename('Intervention' = var)

  melted <- merge(selBase, selInt, by = 'year') %>%
              melt(id.var = 'year', value.name = 'Scenario')

  p <- ggplot(data = melted, aes(x = year, y = Scenario, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f <- paste0(dis, '_prev_intervention.png')

  ggsave(filename = f, 
           plot = p,
           path = path,
           width = 6,
           height = 5)

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visCDComparison(baseline, int, 'cancre')
visCDComparison(baseline, int, 'diabe')
visCDComparison(baseline, int, 'hearte')

```

## Disability Prevalence Comparison

```{r}

# First do % change from 2012 - 2050

visDisabilityChange <- function(base, int, targetYear, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selInt <- int %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Intervention' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  # Merge base and intervention df's
  selected <- inner_join(selBase, selInt, by = 'year')

  # Calculate the difference between targetYear prev and initial prev
  newline <- sapply(selected, function(x) x[2] - x[1])

  # Add the difference as a new row, then transpose and sort out the column names
  selected <- rbind(selected, newline)
  new <- data.frame(t(selected))
  colnames(new) <- c('y2012', 'y2050', 'Difference')

  # Calculate % change, add onto the df and move rownames into their own row
  new$change <- (new$Difference / new$y2012) * 100
  new <- cbind(Scenario = rownames(new), new)
  rownames(new) <- 1:nrow(new)

  # Delete first row, artifacts from the previous preparation
  new <- slice(new, -1)

  p <- ggplot(data = new, aes(x = Scenario, y = change)) +
      geom_bar(stat = 'identity') +
      labs(title = '% Change in Disability Prevalence 2012 - 2050', x = 'Scenario',y = '% Change')

  ggsave(filename = 'disability_prev_change.png', 
           plot = p,
           path = path,
           width = 6,
           height = 5)
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visDisabilityChange(baseline, int, 2050)

```

```{r}

visDisabilityChangeOverTime <- function(base, int, targetYear, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all)

  selInt <- int %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Intervention' = p_anyadl_all)

  # Merge base and intervention df's
  selected <- inner_join(selBase, selInt, by = 'year') %>%
                melt(id.var = 'year', value.name = 'Prevalence')

  p <- ggplot(data = selected, aes(x = year, y = Prevalence, color = variable)) +
      geom_line() +
      labs(title = 'Change in Prevalence of Disability 2012 - 2060', x = 'Year', y = 'Prevalence')

  ggsave(filename = 'disability_prev_change_overTime.png', 
           plot = p,
           path = path,
           width = 6,
           height = 5)

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
int <- read_dta('ELSA_pop_BMI_reduc_obese/ELSA_pop_BMI_reduc_obese_summary.dta')

visDisabilityChangeOverTime(baseline, int, 2050)

```

# Smoking Cessation (SCENARIO II - COHORT)

## Baseline CD prevalence and survival curve

```{r}

visCohortBase <- function(cohort, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_2/') {
  # Generate age var
  cohort$Age <- (cohort$year - 2012) + 50
  # Survival curve
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]

  cdMelt <- cohort %>%
              select(c('Age', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
              rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
              melt(id.var='Age', value.name='Prevalence')

  p1 <- ggplot(data=cohort, aes(x=Age, y=Survival)) +
          geom_line() +
          labs(title = 'Survival Curve')

  ggsave(filename = 'cohort_survival_curve.png', 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  p2 <- ggplot(data=cdMelt, aes(x=Age, y=Prevalence)) +
          geom_line() +
          facet_grid(. ~ variable) +
          theme(panel.spacing.x = unit(1.5, 'lines')) +
          labs(title = 'Chronic Disease Prevalence in Cohort')

  ggsave(filename = 'cohort_base_prev.png', 
         plot = p2,
         path = path,
         width = 10,
         height = 5)

  p1
  p2

}

cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')

visCohortBase(cohort)

```

## Intervention CD prev and survival curve comparison

```{r}

cohort <- read_dta('ELSA_cohort/ELSA_cohort_summary.dta')
noSmoke <- read_dta('ELSA_CnoSmoke/ELSA_CnoSmoke_summary.dta')

visSmokIntervention <- function(cohort, intervention, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_2/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)

  survivals <- survivals %>%
                 rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, 'No Smoking' = intervention.Survival) %>%
                 melt(id.var = 'Age', value.name = 'Survival')

  p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
        geom_line() +
        labs(title = 'Baseline vs Non-smoking', x = 'Age', y = 'Survival')

  ggsave(filename = 'cohort_and_intervention_survival_curve.png', 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  # Now do the chronic disease comparison graphs
  # Collect data for each disease separately
  cancre <- data.frame(cohort$Age, cohort$p_cancre_all, noSmoke$p_cancre_all)
  diabe <- data.frame(cohort$Age, cohort$p_diabe_all, noSmoke$p_diabe_all)
  hearte <- data.frame(cohort$Age, cohort$p_hearte_all, noSmoke$p_hearte_all)
  lunge <- data.frame(cohort$Age, cohort$p_lunge_all, noSmoke$p_lunge_all)

  cancre <- cancre %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_cancre_all, 'No Smoking' = noSmoke.p_cancre_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  diabe <- diabe %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_diabe_all, 'No Smoking' = noSmoke.p_diabe_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  hearte <- hearte %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_hearte_all, 'No Smoking' = noSmoke.p_hearte_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  lunge <- lunge %>%
              rename('Age' = cohort.Age, 'Baseline' = cohort.p_lunge_all, 'No Smoking' = noSmoke.p_lunge_all) %>%
              melt(id.var = 'Age', value.name = 'Prevalence')

  p2 <- ggplot(data = cancre, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Cancer', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'cancre_cohort_intervention.png', 
         plot = p2,
         path = path,
         width = 7,
         height = 5)

  p3 <- ggplot(data = diabe, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Diabetes', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'diabe_cohort_intervention.png', 
         plot = p3,
         path = path,
         width = 7,
         height = 5)

  p4 <- ggplot(data = hearte, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Heart Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'hearte_cohort_intervention.png', 
         plot = p4,
         path = path,
         width = 7,
         height = 5)

  p5 <- ggplot(data = lunge, aes(x = Age, y = Prevalence, color = variable)) +
          geom_line() +
          labs(title = 'Lung Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'lunge_cohort_intervention.png', 
         plot = p5,
         path = path,
         width = 7,
         height = 5)

  p1
  p2
  p3
  p4
  p5

}

visSmokIntervention(cohort, noSmoke)

```

# OLD - DEFUNKT CHUNKS

# Chronic Disease Prevalence (OLD SCENARIO I - FULL POPULATION)

## Baseline Initial Prevalence

```{r}
### visInitBasePrev
# Inputs:   base
#           path
#
# Output:   bar - 
#
# This function... 
visInitBasePrev <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                filter(year == 2012) %>%
                melt(id.var = 'year', value.name = 'Prevalence') %>%
                subset(select = -year)

  bar <- ggplot(data=cdMelted, aes(x=variable, y=Prevalence)) +
          geom_bar(stat = 'identity') +
          labs(title='Initial Prevalence', x='Prevalence', y='Chronic Disease')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'init_baseline_prevalence.png', 
         plot = bar,
         path = path,
         width = 7,
         height = 5)

  bar
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visInitBasePrev(baseline)

```

## Baseline Projections

```{r}
### visProjectedBaselineCD
# Inputs:   base - summary output file from the ELSA_Baseline scenario
#           path - path to the directory where the resulting plot will be saved
#
# Output:   tiled - A 3 column tiled plot with the projected prevalence of Cancer, Diabetes, and Heart Disease on the same x axis, with year on the y
#
# This function visualises the prevalence of Cancer, Diabetes, and Heart Disease in the Baseline projection of the FEM
visProjectedBaselineCD <- function(base, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  # Extract vars, rename, and melt to put it in better format for graphing with ggplot2
  cdMelted <- base %>%
                select(c('year', 'p_cancre_all', 'p_diabe_all', 'p_hearte_all')) %>%
                rename('Cancer' = p_cancre_all, 'Diabetes' = p_diabe_all, 'Heart Disease' = p_hearte_all) %>%
                melt(id.var='year', value.name = 'Prevalence')

  # Create the tiled plot of disease prevalence
  tiled <- ggplot(data=cdMelted, aes(x=year, y=Prevalence)) +
            geom_line() +
            facet_grid(. ~ variable) +
            theme(panel.spacing.x = unit(1.5, 'lines')) +
            labs(title = 'Prevalence of Chronic Disease', x = 'Year', y = 'Prevalence')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = 'baseline_prevalence.png', 
         plot = tiled,
         path = path,
         width = 10,
         height = 5)

  # Show plot
  tiled
}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBaselineCD(baseline)

```

```{r}
### visProjectedBySubpop
# Inputs:   base - 
#           path - 
#           cd - chronic disease to visualise
#
# Output:   None
#
# This function visualises the baseline prevalence of a chronic disease broken down by subpopulations. Currently, the subpopulations included are: 
# - obese/not obese 
# - disabled (anyadl==1)/not disabled 
# - age splines (55-64,65-74,75-84,85p)
visProjectedBySubpop <- function(base, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  }

  # Generate varnames
  all <- paste0('p_', cd, '_all')
  # obesity
  obese <- paste0('p_', cd, '_obese')
  notObese <- paste0('p_', cd, '_notObese')
  # disability
  adl <- paste0('p_', cd, '_anyadl')
  noadl <- paste0('p_', cd, '_noadl')
  # age
  a5564 <- paste0('p_', cd, '_5564')
  a6574 <- paste0('p_', cd, '_6574')
  a7584 <- paste0('p_', cd, '_7584')
  a85p <- paste0('p_', cd, '_85p')

  obeseMelt <- base %>%
              select(c('year', all_of(all), all_of(obese), all_of(notObese))) %>%
              rename('All' = all, 'Obese' = obese, 'Healthy Weight' = notObese) %>%
              melt(id.var='year', value.name = 'Prevalence')

  adlMelt <- base %>%
              select(c('year', all_of(all), adl, noadl)) %>%
              rename('All' = all, 'Disabled' = adl, 'Able' = noadl) %>%
              melt(id.var='year', value.name = 'Prevalence')

  ageMelt <- base %>%
              select(c('year', a5564, a6574, a7584, a85p)) %>%
              rename('55 to 64' = a5564, '65 to 74' = a6574, '75 to 64' = a7584, '85+' = a85p) %>%
              melt(id.var = 'year', value.name = 'Prevalence')

  p1 <- ggplot(data=obeseMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f1 <- paste0(cd, '_prev_by_obesity.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f1, 
         plot = p1,
         path = path,
         width = 7,
         height = 5)

  p2 <- ggplot(data=adlMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f2 <- paste0(cd, '_prev_by_disability.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f2, 
         plot = p2,
         path = path,
         width = 6,
         height = 5)

  p3 <- ggplot(data=ageMelt, aes(x=year, y=Prevalence, color = variable)) +
          geom_line() +
          labs(title = dis, x = 'Year', y = 'Prevalence')

  f3 <- paste0(cd, '_prev_by_age.png')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = f3, 
         plot = p3,
         path = path,
         width = 6,
         height = 5)
  p1
  p2
  p3

}

baseline <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

visProjectedBySubpop(baseline, cd = 'cancre')
visProjectedBySubpop(baseline, cd = 'diabe')
visProjectedBySubpop(baseline, cd = 'hearte')

```

## Disability Comparison 2012 - 2040

This one is slightly more tricky to think about so need a plan.

I want to compare the prevalence of anyadl in 2012 and 2050 (or 2040, or 2060). So first I need to extract the prevalence of anyadl from all scenarios I want to compare.

```{r}

visDisabilityComparison <- function(base, ten, fifty, hundred, targetYear, cd,  path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Scenario_1/') {

  selBase <- base %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('Baseline' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC10 <- ten %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('10%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC50 <- fifty %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('50%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selC100 <- hundred %>%
                select(c('year', 'p_anyadl_all')) %>%
                rename('100%' = p_anyadl_all) %>%
                filter(year == '2012' | year == targetYear)

  selected <- inner_join(selBase, selC10)
  selected <- inner_join(selected, selC50)
  selected <- inner_join(selected, selC100)
  selected <- select(selected, -year)

  newline <- sapply(selected, function(x) x[2] - x[1])

  selected <- rbind(selected, newline)
  new <- data.frame(t(selected))
  colnames(new) <- c('y2012', 'y2050', 'Difference')

  new$change <- (new$Difference / new$y2012) * 100
  #new <- rename(new, '%Increase' = change)

  new <- cbind(Scenario = rownames(new), new)
  rownames(new) <- 1:nrow(new)
  new$Scenario <- factor(new$Scenario, levels = c('Baseline', '10%', '50%', '100%'))

  x_name <- paste0(cd, ' Reduction Scenario')

  p <-ggplot(data = new, aes(x = Scenario, y = change)) +
        geom_bar(stat = 'identity') +
        labs(title = 'Increase in Prevalence of Disability (2012 - 2050)', x = x_name, y = '% Increase')

  fname <- paste0(cd, '_reduc_disability_prev.png')

  ggsave(filename = fname, 
         plot = p,
         path = path,
         width = 6,
         height = 5)

  p

}

base <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

cancre10 <- read_dta('ELSA_pcancre10/ELSA_pcancre10_summary.dta')
cancre50 <- read_dta('ELSA_pcancre50/ELSA_pcancre50_summary.dta')
cancre100 <- read_dta('ELSA_pcancre100/ELSA_pcancre100_summary.dta')

diabe10 <- read_dta('ELSA_pdiabe10/ELSA_pdiabe10_summary.dta')
diabe50 <- read_dta('ELSA_pdiabe50/ELSA_pdiabe50_summary.dta')
diabe100 <- read_dta('ELSA_pdiabe100/ELSA_pdiabe100_summary.dta')

hearte10 <- read_dta('ELSA_phearte10/ELSA_phearte10_summary.dta')
hearte50 <- read_dta('ELSA_phearte50/ELSA_phearte50_summary.dta')
hearte100 <- read_dta('ELSA_phearte100/ELSA_phearte100_summary.dta')

visDisabilityComparison(base, cancre10, cancre50, cancre100, 2050, 'Cancer')
visDisabilityComparison(base, diabe10, diabe50, diabe100, 2050, 'Diabetes')
visDisabilityComparison(base, hearte10, hearte50, hearte100, 2050, 'Heart Disease')

```