### 2021-01-10 FEM_R/mortality_adjustment.Rmd
Added an R notebook file that handles the production of the mortality adjustment table, based on life tables supplied by the ONS. Had to add a column definition into ELSA_vars.txt, and included the settings file in ELSA_baseline.csv for testing. All works now!

---
title: "Generate Mortality Adjustment for FEM"
output: html_notebook
author: Luke Archer
email: l.archer@leeds.ac.uk
---

This notebook will go through the steps taken to generate the mortality adjustment table for the FEM.

The mortality adjustment is a table that specifies how the mortality rate should be adjusted over time in the FEM. As a rule, the mortality rate is decreasing over time. The original US version divided the mortality adjustment into a rate for those aged below 65, and those aged above. As the data we are using contains information by gender and single year of age, we will instead attempt to make a version of this with a separate rate by gender and single year of age.

```{r setup, include = FALSE}

# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(readODS)

```

First read in the .xls document containing the life tables. The life tables are divided into period and cohort tables, which are further subdivided by gender, resulting in 4 separate life tables.

```{r}
# Read in male and female, cohort and period data separately 
m_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 3,
                     skip = 9,
                     row_names = TRUE)
f_period <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 4,
                     skip = 9,
                     row_names = TRUE)
m_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 5,
                     skip = 9,
                     row_names = FALSE)
f_cohort <- read_ods('../Mortality_Projections/engppp18ex.ods',
                     sheet = 6,
                     skip = 9,
                     row_names = FALSE)

# First row is empty so remove
m_period <- m_period[-1,]
f_period <- f_period[-1,]
m_cohort <- m_cohort[-1,]
f_cohort <- f_cohort[-1,]
```

Steps:

- Drop ages below 50
- Assign age to rownames and drop Age column. End up with a matrix indexed by age and year.
- Normalise matrix to 2012 values (column 32)
- Calculate inverse (value < 1 == decreased chance of mortality)

```{r}
# Rename age column and filter out ages below 50
male <- m_cohort %>% 
            rename(age = 'Attained age (years)') %>%
            filter(age >= 50)

female <- f_cohort %>% 
            rename(age = 'Attained age (years)') %>%
            filter(age >= 50)

# Replace rownames with values from Age row
rownames(male) <- male$age
rownames(female) <- female$age
male <- subset(male, select = -age)
female <- subset(female, select = -age)

# Scale all values to the 2012 column
male_norm <- male / male[,32]
female_norm <- female / female[,32]

head(female_norm)

# Take the inverse (we want lower values to represent reduced mortality)
m_inv <- 1 / male_norm
f_inv <- 1 / female_norm

head(f_inv)
```

We now have a value of 1 for all ages in the year 2012, which decreases into the future (reduced chance of mortality), and previous years show an increase (where mortality used to be higher).

The next step is to rearrange the matrices into a single table with 4 columns: gender, age, year, and adjustment (the latter column must be unnamed to be recognised by the model). We can do this with tidyr::gather.

```{r}
# Copy dfs so we don't have to rerun other chunks
m_copy <- m_inv
f_copy <- f_inv

# Add a column for gender and reinsert the rownames back as a column for reshaping
m_copy <- cbind(male = 1, rownames(m_copy), m_copy)
rownames(m_copy) <- 1:nrow(m_copy)
f_copy <- cbind(male = 0, rownames(f_copy), f_copy)
rownames(f_copy) <- 1:nrow(f_copy)
# Rename age column
m_copy <- rename(m_copy, age = "rownames(m_copy)")
f_copy <- rename(f_copy, age = "rownames(f_copy)")

# First attempt at reshaping with gather
m_long <- gather(m_copy, year, adjustment, 3:ncol(m_copy), factor_key = TRUE)
f_long <- gather(f_copy, year, adjustment, 3:ncol(f_copy), factor_key = TRUE)

head(f_long) # Looks good!
```

Last step is just to row bind and stick everything together, then can manually delete the 'adjustment' column heading before adding to a folder called 'mortality_adj_ELSA' in FEM_CPP_settings.

```{r}
# Combine male and female data
adjustment <- rbind(m_long, f_long)

head(adjustment)

# Write to folder outside trunk so we can't accidentally replace it later
write_tsv(adjustment, path = '/home/luke/Documents/E_FEM_clean/Mortality_Projections/mortality_adj.txt', 
                      append = FALSE, 
                      col_names = TRUE)
```