### 2021-01-27 FEM_R/validation.Rmd
Started the validation work in a separate notebook to model analysis

---
title: "External Validation"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
#require(reactable)
require(haven)
#require(data.table)
require(reshape2)
require(ggplot2)
#require(plyr)

```

# AgeUK Almanac External Validation

'The Age UK almanac of disease profiles in later life', published in October 2015, contains information on the prevalence of major diseases, conditions, and syndromes affecting older people in England, broken down by age and sex.

```{r}
# Just read in the data in this chunk, then don't have to have everything in the environment all the time
#AgeUK
mAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
fAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')

# FEM 2014 detailed output
FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
## extract males and females
mFEM_2014 <- FEM_2014 %>% filter(male == 1)
fFEM_2014 <- FEM_2014 %>% filter(male == 0)

# FEM 2014 summary
FEM_sum <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

```

```{r}
# Rename columns
colnames(mAGE_UK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
colnames(fAGE_UK) <- colnames(mAGE_UK)

# Collect conditions into a list for extracting from FEM
# Will need to be written in FEM shorthand
keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
f6064 <- fFEM_2014 %>% filter(age %in% (60:64))

m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
colnames(m6064_prev) <- 'FEM_male_6064'
colnames(f6064_prev) <- 'FEM_female_6064'

prev_df <- data.frame(m6064_prev, f6064_prev)

```

```{r}
male_au <- mAGE_UK
female_au <- fAGE_UK

#mFEM <- mFEM_2014
#fFEM <- fFEM_2014

FEM <- FEM_sum

ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')

m_ages <- paste0('m', ages)

m_ages
```

## First Try

```{r}
# Need a function here to handle the repetition
get_FEM_prev <- function(FEM, age_range, keep_vars) {
  # Keep only important vars
  FEM <- FEM[keep_vars]
  # Separate by gender
  mFEM <- FEM %>% filter(male == 1)
  fFEM <- FEM %>% filter(male == 0)
  # Filter by age from age_range
  mFEM <- mFEM %>% filter(age %in% (age_range))
  fFEM <- fFEM %>% filter(age %in% (age_range))
  # Get proportion of total pop. (prevalence)
  mFEM <- data.frame((colSums(mFEM) / nrow(mFEM)) * 100)
  fFEM <- data.frame((colSums(fFEM) / nrow(fFEM)) * 100)

  # produce a string for column naming
  age <- paste0(age_range[1], '-', age_range[5])

  colnames(mFEM) <- paste0('FEM_male_', age)
  colnames(fFEM) <- paste0('FEM_female_', age)

  #male <- cbind(male, mFEM)

  # Collect into 1 df
  prev_df <- data.frame(mFEM, fFEM)
  return(prev_df)

  #return(out)
}

package <- function() {
  # Set up objects for holding and returning the prevalences, separate dfs for male and female
  out <- list()
  male <- data.frame()
  female <- data.frame()

  age_list <- c('Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+')
  age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)
}

# Now read in male and female AgeUK data
male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
# Rename columns
colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
colnames(female_ageUK) <- colnames(male_ageUK)

# Collect conditions into a list for extracting from FEM
# Will need to be written in FEM shorthand
keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

# Now get FEM output for 2014
FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')

test <- get_FEM_prev(FEM_2014, 60:64, keep_vars)

#FEM_2014 <- FEM_2014[keep_vars]
## extract males and females
mFEM_2014 <- FEM_2014 %>% filter(male == 1)
fFEM_2014 <- FEM_2014 %>% filter(male == 0)

m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
f6064 <- fFEM_2014 %>% filter(age %in% (60:64))

m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
colnames(m6064_prev) <- 'FEM_male_6064'
colnames(f6064_prev) <- 'FEM_female_6064'

prev_df <- data.frame(m6064_prev, f6064_prev)

# AgeUK conditions to extract
#condlist <- c('Dementia', 'Hypertension', 'Diabetes', 'Stroke')

# correct colnames in ageUK stuff
#male_ageUK <- male_ageUK %>% filter(Condition == 'Dementia' | Condition == 'Hypertension' | Condition == 'Diabetes' | Condition == 'Stroke')
#male_ageUK <- male_ageUK %>% filter(Condition %in% condlist)

# Now calculate prevalence by age group for each chronic disease
# prevalence = total / affected
# Diabetes
#male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
#female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
#colnames(male_prev) <- 'FEM_male'
#colnames(female_prev) <- 'FEM_female'

```

```{r}

tmp <- function(m, f, age) {

  m_age <- m %>% filter(age %in% (age))
  f_age <- f %>% filter(age %in% (age))

  m_prev <- data.frame((colSums(m_age) / nrow(m_age)) * 100)
  f_prev <- data.frame((colSums(f_age) / nrow(f_age)) * 100)

  list <- list(m_prev, f_prev)
  return(list)
}

# Now read in male and female AgeUK data
male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
# Rename columns
colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
colnames(female_ageUK) <- colnames(male_ageUK)

# Collect conditions into a list for extracting from FEM
# Will need to be written in FEM shorthand
keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

# Now get FEM output for 2014
FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
FEM <- FEM_2014[keep_vars]

# extract males and females
mFEM <- FEM %>% filter(male == 1)
fFEM <- FEM %>% filter(male == 0)

age_list <- c(60:64, 65:69, 70:74, 75:79, 80:84, 85:89, 90:94, 95:99)

m6064 <- mFEM %>% filter(age %in% (60:64))
f6064 <- fFEM %>% filter(age %in% (60:64))

m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
colnames(m6064_prev) <- 'FEM_male_6064'
colnames(f6064_prev) <- 'FEM_female_6064'

male <- list()
female <- list()

mf_list <- tmp(mFEM, fFEM, 60:64)
male <- list(male, mf_list[1])
female <- list(female, mf_list[2])

mf_list <- tmp(mFEM, fFEM, 65:69)
male <- list(male, mf_list[1])
female <- list(female, mf_list[2])

```