### 2021-02-02 FEM_R/validation.Rmd
WIP - mostly done but trying to figure out the best way to show all the plots together

---
title: "External Validation"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
#require(reactable)
require(haven)
#require(data.table)
require(reshape2)
require(ggplot2)
#require(plyr)

```

# AgeUK Almanac External Validation

'The Age UK almanac of disease profiles in later life', published in October 2015, contains information on the prevalence of major diseases, conditions, and syndromes affecting older people in England, broken down by age and sex.

```{r}
# Just read in the data in this chunk, then don't have to have everything in the environment all the time
#AgeUK
mAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
fAGE_UK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')

# FEM 2014 detailed output
#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
## extract males and females
#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
#fFEM_2014 <- FEM_2014 %>% filter(male == 0)

# FEM 2014 summary
FEM_sum <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

```

```{r}
# Take copy of FEM summary
FEM <- FEM_sum

# Keep only year 2014 from FEM summary
FEM <- filter(FEM, year == 2014)

# Specify the age groups to extract from FEM
ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
# Add m and f for male and female variables
m_ages <- paste0('m_', ages)
f_ages <- paste0('f_', ages)

# Collect the age suffixes into a vector
keep_vars <- c(m_ages, f_ages)
# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

# Don't like using loops but this one was just too simple to bother with figuring out the apply version
final <- c()
ticker <- 1
# Generating the column (variable) names we want to keep in a loop
for(item in diseases) {
  for(anotherItem in keep_vars) {
    final[ticker] <- paste0('p_', item, '_', anotherItem)
    ticker <- ticker + 1
  }
}

# Keep the right vars using varnames generated in loop
FEM <- FEM %>%
          select(c('year', all_of(final)))

# Rename some vars to remove 'p_' before merging
FEM <- FEM %>%
          rename_at(vars(starts_with('p_')),
                    funs(str_replace(., 'p_', ''))) %>%
          select(!year)

pivoted.FEM <- pivot_longer(data = FEM,
                            cols = hearte_m_6064:demene_f_100p, 
                            names_to = c('Condition', 'Gender', 'Age_Group'), 
                            names_sep = '_')

# Convert value from fraction to percentage
pivoted.FEM$value <- pivoted.FEM$value * 100

```

Now process the AgeUK data and combine male and female

```{r}
# Copy over age uk data
male_au <- mAGE_UK
female_au <- fAGE_UK

# Rename cols to enable reshaping
male_au <- male_au %>%
                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
                select(!Total)

female_au <- female_au %>%
                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
                select(!Total)

# Now need to reshape these datasets to long format
pivoted.male <- pivot_longer(data = male_au,
                             cols = m_6064:m_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

pivoted.female <- pivot_longer(data = female_au,
                             cols = f_6064:f_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

# Now row bind these frames together, combining male and female data
ageUK <-rbind(pivoted.male, pivoted.female)

# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
# extract all heart related conditions for combining
ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
# Aggregate the values by gender and age group
ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
# Rename columns and add condition column back
colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
ageUK_heart$Condition <- 'Heart Problems'

# Now drop the heart problems from original ageUK and replace with aggregate
ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
ageUK <- rbind(ageUK, ageUK_heart)

# Rename a few just for completeness
ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)

# Finally, add column to specify that this is ageUK data
ageUK$source <- "ageUK"

# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
ageUK <- filter(ageUK, Condition != 'Asthma', Condition != 'Depression', Condition != 'Severe Mental Health Conditions')

```

```{r}

# COME ON!!!!!!
# Steps:
#   - Rename FEM rows to match AgeUK
#   - Add column to specify between FEM and AgeUK
#   - Merge datasets, maintain long format
#   - ggplot2 some nice grouped bar charts
#   - Also think about some nice reactable tables maybe? With some custom styling

# Make copy
newFEM <- pivoted.FEM

# Rename all the conditions to fit ageUK
# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
#       Therefore only validation from this will be that FEM is higher than ageUK
newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)

# Add column to specify this is FEM data
newFEM$source <- "FEM"

```

```{r}

# Now row bind the datasets
combined <- rbind(newFEM, ageUK)
# Specify factor levels for age group
combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
# Sort
combined <- arrange(combined, source, Condition, Gender, Age_Group)

```

```{r}

# Lets do some plotting!

# First try to plot the whole dataset in a facet grid
# This isn't perfect, and it's pretty complicated to plot altogether
ggplot(data = combined, aes(x = Age_Group, y = value, width = 0.7)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
  facet_wrap(~Condition)

```

```{r}

# Now instead, lets plot an individual condition - Hypertension

# First extract all hypertension info
hyper <- filter(combined, Condition == 'Hypertension')

# Big dirty plot
ggplot(data = hyper, aes(x = Age_Group, y = value, width = 0.7)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
  labs(title = 'Hypertension', x = 'Age Group', y = 'Proportion')

```

```{r}

# Next, lets try an all female and all male example

```