### 2021-02-24 FEM_R/results.Rmd
Many changes to get core scenarios working. Added ELSA_core_scen.csv file for running the core scenarios

---
title: "R Notebook"
output: html_notebook
---

This notebook will hold some initial testing and visualisation for the results section of the IJM paper.

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
#require(plyr)
#require(data.table)

```

```{r}

# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

```

Data has been read in, lets create some functions to visualise.

```{r}

compareSurvivalCurve <- function(cohort, intervention, int_name,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, intervention$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Baseline' = cohort.Survival, !!quo_name(int_name) := intervention.Survival) %>%
                 melt(id.var = 'Age', value.name = 'Survival')

  p1 <-ggplot(data = survivals, aes(x=Age, y=Survival, color = variable)) +
        geom_line() +
        labs(title = paste('Baseline vs', int_name), x = 'Age', y = 'Survival')

  ggsave(filename = paste0(int_name, '_survival_curve.png'),
         plot = p1,
         path = path)

  return(p1)

}

```

```{r}

compareDeathAge <- function(base, int, int_name, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  aDeath <- data.frame(base$year, base$d_age_all, int$d_age_all)

  print(colnames(aDeath))

  aDeath <- aDeath %>%
                dplyr::rename('Year' = base.year, 'Baseline' = base.d_age_all, !!quo_name(int_name) := int.d_age_all) %>%
                reshape2::melt(id.var = 'Year', value.name = 'Age')

  p1 <- ggplot(data = aDeath, aes(x = Year, y = Age, color = variable)) +
        geom_line() +
        labs(title = paste('Baseline vs', int_name), x = 'Year', y = 'Age at Death')

  ggsave(filename = paste0(int_name, '_death_age.png'),
         plot = p1,
         path = path)

  return(p1)
}
```

# Hearte

## Cohort

```{r}

coh <- cohort
h <- heart_c

compareSurvivalCurve(coh, h, 'No Heart Disease')

```

```{r}

coh <- cohort
h <- heart_c

cohortCDVis <- function(cohort, intervention, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  intervention$Age <- (intervention$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  intervention$Survival <- intervention$m_endpop_all / intervention$m_endpop_all[1]

  cohMelt <- cohort %>%
              select(c('Age', all_of(cds))) %>%
              melt(id.var='Age', value.name = 'Prevalence_cohort')

  intMelt <- intervention %>%
              select(c('Age', all_of(cds))) %>%
              melt(id.var='Age', value.name = 'Prevalence_intervention')

  merged <- merge(cohMelt, intMelt)

  pivot.merged <- pivot_longer(data = merged,
                             cols = Prevalence_cohort:Prevalence_intervention,
                             names_to = c('Prevalence', 'source'),
                             names_sep = '_')
  pivot.merged <- select(pivot.merged, -Prevalence)

  p1 <- ggplot(data = pivot.merged, aes(x = Age, y = value, color = source)) +
          geom_line() +
          facet_wrap(. ~ variable, scales = "free")

  ggsave(filename = 'noHearte_cd_lattice.png',
         plot = p1,
         path = path)

  return(p1)

}

cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')

cohortCDVis(coh, h, cds)

```

## Population

```{r}

b <- base
h_p <- heart_p

compareDeathAge(b, h_p, 'No Heart Disease')

```

# Smoking

## Average Age at Death

```{r}

b2 <- base
s <- smok

compareDeathAge(b2, smok, 'No Smoking')

```

## Chronic Diseases

```{r}

baseCDVis <- function(base, int, cds, cds.rename, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  selBase <- base %>%
              select(c('year', all_of(cds)))
  colnames(selBase) <- c('Year', cds.rename)

  selInt <- int %>%
              select(c('year', all_of(cds)))
  colnames(selInt) <- c('Year', cds.rename)

  pivot.selBase <- pivot_longer(data = selBase,
                              cols = 'Lung Disease':Diabetes,
                              names_to = 'Condition',
                              values_to = 'Prevalence')
  pivot.selBase$scen <- 'Baseline'

  pivot.selInt <- pivot_longer(data = selInt,
                              cols = 'Lung Disease':Diabetes,
                              names_to = 'Condition',
                              values_to = 'Prevalence')
  pivot.selInt$scen <- 'Intervention'

  merged <- rbind(pivot.selBase, pivot.selInt)

  p <- ggplot(data = merged, aes(x = Year, y = Prevalence, color = scen)) +
          geom_line() +
          facet_wrap(. ~ Condition, scales = "free")
          labs(x = 'Year', y = 'Prevalence')

  ggsave(filename = 'noSmoke_cd_lattice.png',
         plot = p,
         path = path)

  return(p)
}

cds <- c('p_lunge_all', 'p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all')
cds.rename <- c('Lung Disease', 'Heart', 'Cancer', 'Stroke', 'Diabetes')
baseCDVis(b2, s, cds, cds.rename)

```

```{r}

visCDComparison <- function(base, int, int_name, cd, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # For file naming and stuff
  if(cd == 'cancre') {
    dis <- 'Cancer'
  } else if(cd == 'diabe') {
    dis <- 'Diabetes'
  } else if(cd == 'hearte') {
    dis <- 'Heart Disease'
  } else if(cd == 'lunge') {
    dis <- 'Lung Disease'
  } else if(cd == 'stroke') {
    dis <- 'Stroke'
  }

  var <- paste0('p_', cd, '_all')

  selBase <- base %>%
              select(c('year', all_of(var))) %>%
              dplyr::rename('Baseline' = var)

  selInt <- int %>%
              select(c('year', all_of(var))) %>%
              dplyr::rename('Intervention' = var)

  melted <- merge(selBase, selInt, by = 'year') %>%
              reshape2::melt(id.var = 'year', value.name = 'Scenario')

  p <- ggplot(data = melted, aes(x = year, y = Scenario, color = variable)) +
          geom_line() +
          labs(title = paste(int_name, dis, sep = ' - '), x = 'Year', y = 'Prevalence')

  ggsave(filename = paste0(int_name, '_', dis, '_prev.png'),
         plot = p,
         path = path)

  return(p)

}

visCDComparison(b2, s, 'No Smoke', 'cancre')
visCDComparison(b2, s, 'No Smoke', 'diabe')
visCDComparison(b2, s, 'No Smoke', 'hearte')
visCDComparison(b2, s, 'No Smoke', 'lunge')
visCDComparison(b2, s, 'No Smoke', 'stroke')

```

# BMI

```{r}

c <- cohort
bmi1 <- obese1
bmi2 <- obese2
bmi3 <- nomorbid

compareSurvivalCurveBMI <- function(cohort, int1, int2, int3,
                                 path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  int1$Age <- (int1$year - 2012) + 50
  int2$Age <- (int2$year - 2012) + 50
  int3$Age <- (int3$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  int1$Survival <- int1$m_endpop_all / int1$m_endpop_all[1]
  int2$Survival <- int2$m_endpop_all / int2$m_endpop_all[1]
  int3$Survival <- int3$m_endpop_all / int3$m_endpop_all[1]

  survivals <- data.frame(cohort$Age, cohort$Survival, int1$Survival, int2$Survival, int3$Survival)

  # Had a small problem when trying to rename using the variable name given as function argument
  # Found solution here: https://stackoverflow.com/questions/49650394/how-to-rename-a-variable-using-a-dynamic-name-and-dplyr
  survivals <- survivals %>%
                 dplyr::rename('Age' = cohort.Age, 'Cohort' = cohort.Survival, 'No Obese 1' = int1.Survival, 'No Obese 2' = int2.Survival, 'No Morbid' = int3.Survival)

  pivot.survival <- pivot_longer(data = survivals,
                               cols = Cohort:'No Morbid',
                               names_to = 'Scenario',
                               values_to = 'Survival')

  p1 <-ggplot(data = pivot.survival, aes(x=Age, y=Survival, color = Scenario)) +
        geom_line() +
        labs(title = 'BMI Intervention Survival', x = 'Age', y = 'Survival')

  ggsave(filename = 'BMI_survival_curves.png',
         plot = p1,
         path = path)

  return(p1)

}

compareSurvivalCurveBMI(cohort, bmi1, bmi2, bmi3)

```

```{r}

c <- cohort
bmi1 <- obese1
bmi2 <- obese2
bmi3 <- nomorbid

# # Calculate age and survival curves
# c$Age <- (c$year - 2012) + 50
# bmi1$Age <- (bmi1$year - 2012) + 50
# bmi2$Age <- (bmi2$year - 2012) + 50
# bmi3$Age <- (bmi3$year - 2012) + 50
# c$Survival <- c$m_endpop_all / c$m_endpop_all[1]
# bmi1$Survival <- bmi1$m_endpop_all / bmi1$m_endpop_all[1]
# bmi2$Survival <- bmi2$m_endpop_all / bmi2$m_endpop_all[1]
# bmi3$Survival <- bmi3$m_endpop_all / bmi3$m_endpop_all[1]
# 
# cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')
# 
# coh <- c %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_cohort = p_hearte_all, cancer_cohort = p_cancre_all, 
#                           stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
#                           bmi_cohort = a_bmi_all)
# 
# b1 <- bmi1 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi1 = p_hearte_all, cancer_bmi1 = p_cancre_all, 
#                           stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
#                           bmi_bmi1 = a_bmi_all)
# 
# b2 <- bmi2 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi2 = p_hearte_all, cancer_bmi2 = p_cancre_all, 
#                           stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
#                           bmi_bmi2 = a_bmi_all)
# 
# b3 <- bmi3 %>%
#             select(Age, all_of(cds)) %>%
#             dplyr::rename(heart_bmi3 = p_hearte_all, cancer_bmi3 = p_cancre_all, 
#                           stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
#                           bmi_bmi3 = a_bmi_all)
# 
# merged <- Reduce(merge, list(coh, b1, b2, b3))
# 
# pivot.merged <- pivot_longer(data = merged,
#                              cols = heart_cohort:bmi_bmi3,
#                              names_to = c('Feature', 'Scenario'),
#                              names_sep = '_',
#                              values_to = 'Value')
# 
# ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()
# 
# ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
#   geom_line()

visCDComparisonBMI <- function(cohort, int1, int2, int3, cds, path='/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/results/') {

  # Calculate age and survival curves
  cohort$Age <- (cohort$year - 2012) + 50
  int1$Age <- (int1$year - 2012) + 50
  int2$Age <- (int2$year - 2012) + 50
  int3$Age <- (int3$year - 2012) + 50
  cohort$Survival <- cohort$m_endpop_all / cohort$m_endpop_all[1]
  int1$Survival <- int1$m_endpop_all / int1$m_endpop_all[1]
  int2$Survival <- int2$m_endpop_all / int2$m_endpop_all[1]
  int3$Survival <- int3$m_endpop_all / int3$m_endpop_all[1]

  coh <- cohort %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_cohort = p_hearte_all, cancer_cohort = p_cancre_all, 
                            stroke_cohort = p_stroke_all, diabetes_cohort = p_diabe_all, 
                            bmi_cohort = a_bmi_all)

  b1 <- int1 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi1 = p_hearte_all, cancer_bmi1 = p_cancre_all, 
                            stroke_bmi1 = p_stroke_all, diabetes_bmi1 = p_diabe_all,
                            bmi_bmi1 = a_bmi_all)

  b2 <- int2 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi2 = p_hearte_all, cancer_bmi2 = p_cancre_all, 
                            stroke_bmi2 = p_stroke_all, diabetes_bmi2 = p_diabe_all,
                            bmi_bmi2 = a_bmi_all)

  b3 <- int3 %>%
              select(Age, all_of(cds)) %>%
              dplyr::rename(heart_bmi3 = p_hearte_all, cancer_bmi3 = p_cancre_all, 
                            stroke_bmi3 = p_stroke_all, diabetes_bmi3 = p_diabe_all,
                            bmi_bmi3 = a_bmi_all)

  merged <- Reduce(merge, list(coh, b1, b2, b3))

  pivot.merged <- pivot_longer(data = merged,
                               cols = heart_cohort:bmi_bmi3,
                               names_to = c('Feature', 'Scenario'),
                               names_sep = '_',
                               values_to = 'Value')

  p1 <-ggplot(data = filter(pivot.merged, Feature == 'bmi'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Average BMI', x = 'Age', y = 'BMI')

  ggsave(filename = 'BMI_abmi.png',
         plot = p1,
         path = path)

  p2 <- ggplot(data = filter(pivot.merged, Feature == 'heart'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Heart Disease', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_phearte.png',
         plot = p2,
         path = path)

  p3 <- ggplot(data = filter(pivot.merged, Feature == 'stroke'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Stroke', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_pstroke.png',
         plot = p3,
         path = path)

  p4 <- ggplot(data = filter(pivot.merged, Feature == 'diabetes'), aes(x = Age, y = Value, color = Scenario)) +
        geom_line() +
        labs(title = 'Diabetes', x = 'Age', y = 'Prevalence')

  ggsave(filename = 'BMI_pdiabe.png',
         plot = p4,
         path = path)

  return(list(p1, p2, p3, p4))
}

cds <- c('p_hearte_all', 'p_cancre_all', 'p_stroke_all', 'p_diabe_all', 'a_bmi_all')

visList <- visCDComparisonBMI(c, bmi1, bmi2, bmi3, cds)

for(graph in visList) {
  print(graph)
}

```

```{r}

visCDComparison(c, bmi3, 'No Morbid', 'cancre')
visCDComparison(c, bmi3, 'No Morbid', 'diabe')
visCDComparison(c, bmi3, 'No Morbid', 'stroke')
visCDComparison(c, bmi3, 'No Morbid', 'hearte')
visCDComparison(c, bmi3, 'No Morbid', 'lunge')

```