### 2021-04-30 FEM_R/calculate_dflys_from_summary.Rmd
Started working on dflys, added some derived vars to ELSA_vars.txt and included them in measures_subpop_ELSA to work with in summary output files

---
title: "R Notebook"
output: html_notebook
---

Going to have a go at producing estimates of disease free life years and disability free life years.

To do this we need to use the 'disabled' and 'anydisease' variables that are defined in ELSA_vars.txt.

First setup and reading in data files.

# Setup
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
#require(reshape2)
require(ggplot2)
#require(ggpubr)
#require(plyr)
#require(data.table)

```

# Read in Data
```{r}
## Read in all the data
# Baselines
base <- read_dta('ELSA_core_base/ELSA_core_base_summary.dta')
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_summary.dta')

# Hearte (cohort and pop)
#heart_c <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_summary.dta')
heart_p <- read_dta('ELSA_core_remove_hearte_p/ELSA_core_remove_hearte_p_summary.dta')

# Smoking (cohort)
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_summary.dta')

# BMI
nomorbid <- read_dta('ELSA_core_nomorbid/ELSA_core_nomorbid_summary.dta')
obese2 <- read_dta('ELSA_core_noobese2/ELSA_core_noobese2_summary.dta')
obese1 <- read_dta('ELSA_core_noobese1/ELSA_core_noobese1_summary.dta')

# Drinking
nodrink <- read_dta('ELSA_core_remove_drink_p/ELSA_core_remove_drink_p_summary.dta')
noBadDrink <- read_dta('ELSA_core_remove_problem_drink_p/ELSA_core_remove_problem_drink_p_summary.dta')
```

# Disability Free Life Years

Going to attempt disease free first.

What we are looking for is a change in the total number of people with no diseases. i.e. change in n_anydisease_all

Bryan said to 'sum it up for the whole duration of the simulation'. I don't know what this means, but if I have a good go at it now I can always ask him next time we meet.

```{r}
c <- cohort
s <- smok

cSel <- c %>%
          select('year', n_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_disabled_all, 'total' = m_endpop_all)

sSel <- s %>%
          select('year', n_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_disabled_all, 'total' = m_endpop_all)

#combined <- merge(cSel, sSel, by='year')

# Add source and rbind
cSel$source <- 'cohort'
sSel$source <- 'intervention'

combined <- rbind(cSel, sSel)

ggplot(combined, aes(x = year, y = disabled, color = source)) +
  geom_line()
```

The graph above shows why we cannot rely on graphical interpretations of the 'n_' var alone.

Let's try to get the number of disabled years per person.

```{r}
cSel2 <- cSel
sSel2 <- sSel

# Population is in millions initially
cSel2$total <- cSel2$total * 1000000
sSel2$total <- sSel2$total * 1000000

cSel2$dis_by_person <- cSel2$disabled / cSel2$total
sSel2$dis_by_person <- sSel2$disabled / sSel2$total

combined2 <- rbind(cSel2, sSel2)

ggplot(data = combined2, aes(x = year, y = dis_by_person, color = source)) +
  geom_line()

```

## Try again, this time with the correct variable (not disabled)

```{r}
c <- cohort
s <- smok

cSel <- c %>%
          select('year', n_not_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_not_disabled_all, 'total' = m_endpop_all)

sSel <- s %>%
          select('year', n_not_disabled_all, m_endpop_all) %>%
          dplyr::rename('disabled' = n_not_disabled_all, 'total' = m_endpop_all)

#combined <- merge(cSel, sSel, by='year')

# Add source and rbind
cSel$source <- 'cohort'
sSel$source <- 'intervention'

combined <- rbind(cSel, sSel)

ggplot(combined, aes(x = year, y = disabled, color = source)) +
  geom_line()
```

```{r}
cSel2 <- cSel
sSel2 <- sSel

# Population is in millions initially
cSel2$total <- cSel2$total * 1000000
sSel2$total <- sSel2$total * 1000000

cSel2$dis_by_person <- cSel2$disabled / cSel2$total
sSel2$dis_by_person <- sSel2$disabled / sSel2$total

combined2 <- rbind(cSel2, sSel2)

ggplot(data = combined2, aes(x = year, y = dis_by_person, color = source)) +
  geom_line()

```