### 2021-05-28 FEM_R/dfly_detailed_output.Rmd

---
title: "R Notebook"
output: html_notebook
---

This script will attempt a couple of things using detailed output data files. We will attempt:
- Disability free life years (and disease free)
- Treatment on the treated analysis

First we're going to have a go at replicating the file tot_test.do, saved in E_FEM_clean/Bryan_scripts

```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/tmp_output/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)
require(ggplot2)
require(ggpubr)
#require(plyr)
#require(data.table)

theme('minimal')
```

Setup is complete (including the working directory), so let's read in some of the data files.

We are using detailed output, so need the appended data files that should be in the output folders.

```{r}
cohort <- read_dta('ELSA_core_cohort/ELSA_core_cohort_append.dta')
heart <- read_dta('ELSA_core_remove_hearte_c/ELSA_core_remove_hearte_c_append.dta')
smok <- read_dta('ELSA_core_remove_smoken/ELSA_core_remove_smoken_append.dta')
```

Now start with the cohort file, and try to calculate a host of things like life years, dflys, and other things like that.

Generate stats for lung disease and heart disease.

```{r}

# generate flag for initial wave with disease and year died (if known)
new <- cohort %>%
          group_by(hhidpn, mcrep, year) %>%

```

```{r}

aData <- read_dta('/home/luke/Documents/E_FEM_clean/E_FEM/input_data/detailed_output/tot/base_treated_merged_smoken_tot.dta')

```

```{r}

educ <- aData %>%
          group_by(educ) %>%
          summarise(improv.ly = mean(improved_lifeyears), improv.dfly = mean(improved_dflys))

educ

```

```{r}
educ2 <- aData %>%
          group_by(educ) %>%
          summarise(base.ly = mean(total_lifeyears_basecase), int.ly = mean(total_lifeyears),
                    base.dfly = mean(total_dflys_basecase), int.dfly = mean(total_dflys))

educ2
```