### 2021-11-22 FEM_R/impute_alcbase.Rmd
Imputed alcbase information for drinkers who report not having drank anything in the week before the survey

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(mice)

```

## Multiple imputation

We will do multiple imputation of

```{r}
wv <- read_dta('input_data/H_ELSA_g2_wv_specific.dta')
```

Now set up lists of the variables we're going to keep and impute. This will have to be the wave specific variables 
```{r}

id.var <- 'idauniq'
demog.vars <- c('ragender', 'raeducl')
age.vars <- c('r4agey', 'r5agey', 'r6agey', 'r7agey', 'r8agey', 'r9agey')
scako.vars <- c('r4scako', 'r5scako', 'r6scako', 'r7scako', 'r8scako', 'r9scako')
alc.vars <- c('r4alcbase', 'r5alcbase', 'r6alcbase', 'r7alcbase', 'r8alcbase', 
              'r9alcbase')

# first extract just these vars from original data
test <- wv %>% select(all_of(id.var), all_of(demog.vars), all_of(age.vars),
                      all_of(scako.vars), all_of(alc.vars))

# Calculate the annual and monthly consumption
for (v in alc.vars) {
  newVarName.a <- paste0(v, '_a')
  test[[newVarName.a]] <- test[[v]] * 52

  newVarName.m <- paste0(v, '_m')
  test[[newVarName.m]] <- test[[v]] * 4
}

#imp <- mice(data=test, m=5, maxit=10, printFlag = TRUE)

```

Right, data is prepared for mice, but we need to make it work in a very specific way.
Information must be imputed within scako groups. So I think a multi-layered for loop is the only way to do this. I am envisaging:
- Loop over alc.vars
- Loop over `wv'scako values
- Generate a df of

```{r}

miss.list <- list()
# generate a separate df for each wave
imp.list <- list()
for (x in 4:9) {
  scak <- paste0('r', x, 'scako')
  alcb <- paste0('r', x, 'alcbase')

  newDF <- test %>% select(all_of(id.var), all_of(demog.vars),
                                  paste0('r', x, 'agey'),
                                  all_of(scak), 
                                  all_of(alcb))

  # Extract records where scako != 1:7 (i.e. missing or not drank in last 12 months)
  scako.missings <- newDF[!newDF[[scak]] %in% 1:7,]

  # Add to a list so we can rejoin later
  miss.list[[x]] <- scako.missings

  # Now select only the non-missing drinkers
  impDF <- newDF[newDF[[scak]] %in% 1:7,]

  #impDF[[alcb]][impDF[[alcb]] == 0.0,] <- NA
  impDF[[alcb]][impDF[[alcb]] == 0] <- NA

  imp <- mice(data = impDF, m=5, maxit=10, printFlag = FALSE)
  imp.complete <- rbind(complete(imp), miss.list[[x]])

  # Make sure anyone who said they haven't drank in last 12 months has alcbase == 0
  imp.complete[[alcb]][imp.complete[[scak]] == 8] <- 0
  # and finally drop anything that isn't wave specific for the merge in the next chunk
  imp.complete <- imp.complete %>% select(-ragender, -raeducl)

  # save to the list
  imp.list[[x]] <- imp.complete
}

```

```{r}

# Last step is to merge all the dataframes together on the id var and save 
# so it can be merged with the original data in stata
# Also remove all other vars but the alcbase vars as we only want the imputed versions of them
final.imputed <- imp.list %>% discard(is.null) %>%
                reduce(inner_join, by = "idauniq")

# drop age and scako too? Yes for first attempt
final.imputed <- final.imputed %>% select(-contains('scako'), -contains('agey'))

write_dta(final.imputed, path = 'input_data/alcbase_imputed.dta')
```