### 2022-05-30 FEM_R/BMI_kdensity_plots.Rmd
Lots of WIP and changes for the lockdown_alcohol intervention

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(reshape2)

```

```{r}
ELSA <- read_dta('input_data/ELSA_long.dta')
raw.ELSA <- read_dta('input_data/H_ELSA_g2_wv_specific.dta')
```

```{r}
bmi <- ELSA %>% select(wave, male, bmi)

bmi.m <- bmi %>% filter(male == 1)
bmi.f <- bmi %>% filter(male == 0)

bmi.m.2 <- bmi.m %>% filter(wave == 2) %>%
                        select(bmi) %>%
                        drop_na()
bmi.f.2 <- bmi.f %>% filter(wave == 2) %>%
                        select(bmi) %>%
                        drop_na()
```

```{r}
m.dens <- density(bmi.m.2$bmi)
f.dens <- density(bmi.f.2$bmi)
```

```{r}
plot(m.dens, col='blue', lwd=2)
lines(f.dens, col='red', lwd=2)
```