### 2022-05-30 FEM_R/OLD_STUFF/summary_statistics_forPaper.Rmd

---
title: "Starting Populations Summary Stats"
output: html_notebook
---

# Setup

## WorkDir & Libraries
```{r "setup", include=FALSE}
# Need to set up the environment
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/input_data/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(haven)
require(ggplot2)
require(Hmisc)

rm(workingDir)
```

## Data
```{r}

stock <- read_dta('ELSA_stock.dta')
repl <- read_dta('ELSA_repl.dta')
trans <- read_dta('ELSA_transition.dta')

```

# Summary Statistics
## Stock
```{r}
# What do we need to know?
# Population average of:
# - Year
# - Age
# - Gender
# - BMI
#   - % Distribution in BMI categories (<25, 25-30, >30)
# - Smoking Status
#   - Ever Smoked
#   - Current Smoker
# - Education Level
#   - Level 1
#   - Level 2
#   - Level 3
# - Initial Disease Prevalence
#   - Cancer
#   - Diabetes
#   - Heart Disease
#   - MORE!

# Try this:
# https://thatdatatho.com/easily-create-descriptive-summary-statistic-tables-r-studio/

stock.sum <- stock %>%
            summarise(count = n(),
                      year = median(entry, na.rm = TRUE),
                      wave = median(wave, na.rm = TRUE),
                      age.mean = mean(age, na.rm = TRUE),
                      age.sd = sd(age, na.rm = TRUE),
                      male = mean(male, na.rm = TRUE),
                      bmi = mean(bmi, na.rm = TRUE))

stock.sum

#bmi.cut <- cut(stock$bmi, breaks = c(20, 25, 30))

rm(stock.sum)
```

# Weighted Means
```{r}
weighted.mean(stock$male, stock$weight)
print('-------')
weighted.mean(stock$cancre, stock$weight)
weighted.mean(stock$diabe, stock$weight)
weighted.mean(stock$hearte, stock$weight)
weighted.mean(stock$stroke, stock$weight)
weighted.mean(stock$lunge, stock$weight)
weighted.mean(stock$hibpe, stock$weight)
weighted.mean(stock$alzhe, stock$weight)
weighted.mean(stock$demene, stock$weight)
weighted.mean(stock$hchole, stock$weight)

```

```{r}
s2 <- stock

s2$bmi_cat <- cut(s2$bmi, breaks = c(25, 30))

#weighted.mean()
```

```{r}
require(survey)

stock.design <- svydesign(id = ~idauniq,
                          weights = ~weight,
                          data = stock)

svymean(~male, stock.design, na.rm=TRUE)
svymean(~cancre, stock.design, na.rm=TRUE)
svymean(~age, stock.design, na.rm=TRUE)

print('Standard deviation')
dev <- 0.1444 * sqrt(length(stock.design$variables))
dev

```

```{r}
require(survey)
weighted.survey.means <- function(init.pop, transition=FALSE) {

  if(!transition) {
    design <- svydesign(id = ~idauniq,
                      weights = ~weight,
                      data = init.pop)
  } else if(transition) {
    # Error from missing cwtresp var - drop all that are missing to solve
    init.pop <- init.pop[complete.cases(init.pop$cwtresp),]
    design <- svydesign(id = ~idauniq,
                      weights = ~cwtresp,
                      data = init.pop)
  }
  print(svymean(~age, design, na.rm=TRUE))
  print(svyvar(~age, design, na.rm=TRUE))
  print(svymean(~male, design, na.rm=TRUE))
  print(svymean(~bmi, design, na.rm=TRUE))
  print(svyvar(~bmi, design, na.rm=TRUE))
  print(svymean(~smoken, design, na.rm=TRUE))
  print(svymean(~smokev, design, na.rm=TRUE))
  print(svymean(~hsless, design, na.rm=TRUE))
  print(svymean(~college, design, na.rm=TRUE))
  print('---------------------')
  print(svymean(~cancre, design, na.rm=TRUE))
  print(svymean(~diabe, design, na.rm=TRUE))
  print(svymean(~hearte, design, na.rm=TRUE))
  print(svymean(~stroke, design, na.rm=TRUE))
  print(svymean(~lunge, design, na.rm=TRUE))
  print(svymean(~hibpe, design, na.rm=TRUE))
  print(svymean(~alzhe, design, na.rm=TRUE))
  print(svymean(~demene, design, na.rm=TRUE))
  print(svymean(~hchole, design, na.rm=TRUE))
}

weighted.survey.means(stock)
```

```{r}
# Replenishing
weighted.survey.means(repl)
```

```{r}
# Transition
weighted.survey.means(trans, transition=TRUE)
```

```{r}
# Now calculate incidence rates of chronic diseases (& some other stuff)
# incidence = total no. new cases / number of people at risk

# I think we'll need to group by year (or wave), and get counts of chronic disease ==1

getIncidenceRate <- function(t, var) {

  l2var <- paste0('l2', var)

  select.vars <- c(var, l2var)

  t <- select(t, year, all_of(select.vars))
  t$incidence.counts <- (t[,var] == 1) & (t[,l2var] == 0)

  t <- t %>%
          summarise(n = n(), newIn = sum(incidence.counts, na.rm=TRUE)) %>%
          mutate(incidence.rate = (newIn / n) * 100)

  return(t)
}

getIncidenceRate(trans, 'cancre')
getIncidenceRate(trans, 'diabe')
getIncidenceRate(trans, 'hearte')
getIncidenceRate(trans, 'stroke')
getIncidenceRate(trans, 'lunge')
getIncidenceRate(trans, 'hibpe')
getIncidenceRate(trans, 'alzhe')
getIncidenceRate(trans, 'demene')
getIncidenceRate(trans, 'hchole')
```