### 2022-07-12 FEM_R/dev_output_vis_and_test_alcohol.Rmd

---
title: "R Notebook"
output: html_notebook
---

# SETUP
## Libraries
```{r "setup", include=FALSE}
require(tidyverse)
require(haven)
require(ggplot2)
require(Hmisc)
require(reshape2)

workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

## Load in data
```{r}
base_path <- 'output/ALCOHOL/ELSA_full/'
lockdown_path <- 'output/ALCOHOL/ELSA_lockdown/'

ldown <- read_dta(paste0(lockdown_path, 'ELSA_lockdown_summary.dta'))

base <- read_dta(paste0(base_path, 'ELSA_full_summary.dta'))

mup <- read_dta('output/ALCOHOL/ELSA_alcInt_full/ELSA_alcInt_full_summary.dta')
#mup2 <- read_dta('output/ALCOHOL/ELSA_alcInt2_full/ELSA_alcInt2_full_summary.dta')
```

Now write a generic function to visualise a single continuous variable over time. Lets try with average unit consumption.

# Vis single continous

```{r}
vis_continuous_output <- function(data, var, name, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  # select column we want and rename for the visualisation
  selected <- data %>% select(year, all_of(var)) %>%
                        filter(year < end.yr) %>%
                        rename(!!quo_name(name) := var)
  # plot
  ggplot(selected, aes_string(x = 'year', y = name)) +
    geom_line()
}

vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption')
vis_continuous_output(ldown, 'a_alcbase_alc_ldown_treated', 'alcohol_consumption')
```

```{r}
vis_continuous_output(base, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption', horizon = 20)
vis_continuous_output(mup, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
```

```{r}
vis_continuous_output(ldown, 'a_alcbase_all', 'alcohol_consumption', horizon = 50)
vis_continuous_output(ldown, 'a_alcbase_alc_ldown_treated', 'alcohol_consumption', horizon = 50)
```

Now lets do a visualisation of an ordinal variable, in this case alcstat or alcstat4.

```{r}

vis_ordinal_output <- function(data, key.var.list, suffix, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  # add prefix and suffix
  key.var.list <- lapply(1:length(key.var.list), function(k) paste0('p_', key.var.list[[k]], suffix))
  # unlist and get the vars for pivoting
  first.var <- unlist(key.var.list[1])
  last.var <- unlist(key.var.list[length(key.var.list)])

  # select vars and pivot
  selected <- data %>% select(year, contains(all_of(unlist(key.var.list)))) %>%
                        filter(year < end.yr) %>%
                        pivot_longer(cols = first.var:last.var,
                                     names_to = 'Variable',
                                     values_to = 'Prevalence')
  # visualise
  ggplot(selected, aes(x = year, y = Prevalence, color = Variable)) +
    geom_line() +
    labs(title = 'Single Scenario Ordinal', x = 'Year')
}

```

```{r}
vis_ordinal_output(base, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all', horizon = 50)
#vis_ordinal_output(ldown, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all', horizon = 50)
vis_ordinal_output(mup, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all')
#vis_ordinal_output(mup2, c('abstainer', 'moderate', 'increasingRisk', 'highRisk'), '_all')
```

```{r}

vis_ordinal_out_compare <- function(base, scen2, prefix, var, suffix, horizon = 50) {

  # calculate horizon
  start.yr <- 2011
  end.yr <- start.yr + horizon

  key.var <- paste0(prefix, var, suffix)

  base.select <- base %>% select(year, all_of(key.var)) %>%
                            rename(Baseline = key.var) %>%
                            filter(year < end.yr)
  scen2.select <- scen2 %>% select(year, all_of(key.var)) %>%
                            rename(Intervention = key.var) %>%
                            filter(year < end.yr)

  merged <- merge(base.select, scen2.select, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Baseline:Intervention,
                          names_to = 'Scenario',
                          values_to = 'Proportion')

  ggplot(pivoted, aes(x = year, y = Proportion, color = Scenario)) +
    geom_line() +
    labs(title = var, x = 'Year')

}

vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'abstainer', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'moderate', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'increasingRisk', '_all', horizon = 20)
vis_ordinal_out_compare(base = base, scen2 = ldown, 'p_', 'highRisk', '_all', horizon = 20)

```

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'abstainer', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'moderate', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'increasingRisk', '_all', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'highRisk', '_all', horizon = 50)
```

```{r}
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'abstainer', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'moderate', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'increasingRisk', '_all', horizon = 50)
# vis_ordinal_out_compare(base = base, scen2 = mup2, 'p_', 'highRisk', '_all', horizon = 50)
```

NOW LOOK AT SOME HEALTH OUTCOMES IN THE MUP CASE.

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'hearte', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'diabe', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'alzhe', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'demene', '_highRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'd_', 'age', '_highRisk', horizon = 50)
```

```{r}
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'hearte', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'diabe', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'alzhe', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'p_', 'demene', '_increasingRisk', horizon = 50)
vis_ordinal_out_compare(base = base, scen2 = mup, 'd_', 'age', '_increasingRisk', horizon = 50)
```

# Internal Validation of alcbase and alcstat

```{r}

long <- read_dta('input_data/ELSA_long.dta')
cv1 <- read_dta('output/ALCOHOL/ELSA_CV1/ELSA_CV1_summary.dta')

```

```{r}

# 
# 
# var <- 'abstainer'
# #var <- enquo(var)
# 
# # , abstainer, moderate, increasingRisk, highRisk
# 
# long.alcstat.sum <- long %>% select(year, .data[[var]]) %>%
#                       drop_na() %>%
#                       group_by(year, .data[[var]]) %>%
#                       summarise(n = n())
# 
# long.alcstat.sum2 <- long.alcstat.sum %>%
#                       mutate(prop = n/sum(n) * 100) %>%
#                       filter(.data[[var]] == 1) %>%
#                       select(-.data[[var]], -n) %>%
#                       rename(Raw = prop)
# 
# cv1.alcstat.sum <- cv1 %>% select(year, p_abstainer_all) %>%
#                             rename(Simulated = p_abstainer_all) %>%
#                             mutate(Simulated = Simulated * 100)
# 
# merged <- merge(long.alcstat.sum2, cv1.alcstat.sum, by = 'year')
# 
# pivoted <- pivot_longer(data = merged,
#                         cols = Raw:Simulated,
#                         names_to = 'Data',
#                         values_to = 'Percentage')
# 
# ggplot(pivoted, aes(x = year, y = Percentage, color = Data)) +
#   geom_line()

compare_cv_binary <- function(long, cv1, var) {

  #var <- enquo(var)

  long.sum <- long %>% select(year, .data[[var]]) %>%
                      drop_na() %>%
                      group_by(year, .data[[var]]) %>%
                      summarise(n = n()) %>%
                      mutate(prop = n/sum(n) * 100) %>%
                      filter(.data[[var]] == 1) %>%
                      select(-.data[[var]], -n) %>%
                      rename(Raw = prop)

  p.var <- paste0('p_', var, '_all')

  cv1.sum <- cv1 %>% select(year, p.var) %>%
                              rename(Simulated = p.var) %>%
                              mutate(Simulated = Simulated * 100)
  merged <- merge(long.sum, cv1.sum, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Raw:Simulated,
                          names_to = 'Data',
                          values_to = 'Percentage')
  ggplot(pivoted, aes(x = year, y = Percentage, color = Data)) +
    geom_line() +
    labs(title = var, x = 'Year')

  #ggsave(filename = paste(var, '_crossval.png'),
  #       path = 'Deliverables/Conferences/NordicDemographic/cv')

}

compare_cv_binary(long, cv1, var = 'abstainer')
compare_cv_binary(long, cv1, var = 'moderate')
compare_cv_binary(long, cv1, var = 'increasingRisk')
compare_cv_binary(long, cv1, var = 'highRisk')

```

```{r}
long.alcbase.sum <- long %>% select(year, alcbase) %>%
                      drop_na() %>%
                      group_by(year) %>%
                      summarise(avg_alcbase = mean(alcbase))

cv1.alcbase.sum <- cv1 %>% select(year, a_alcbase_all)

compare_cv_continuous <- function(long, cv1, var) {

  long.sum <- long %>% select(year, .data[[var]]) %>%
                      drop_na() %>%
                      group_by(year) %>%
                      summarise(avg_alcbase = mean(alcbase)) %>%
                      rename(Raw = avg_alcbase)

  a.var <- paste0('a_', var, '_all')

  cv1.sum <- cv1 %>% select(year, a.var) %>%
                              rename(Simulated = a.var)

  merged <- merge(long.sum, cv1.sum, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Raw:Simulated,
                          names_to = 'Data',
                          values_to = 'Consumption')
  ggplot(pivoted, aes(x = year, y = Consumption, color = Data)) +
    geom_line() +
    labs(title = var, x = 'Year')

  ggsave(filename = paste(var, '_crossval.png'),
         path = 'Deliverables/Conferences/NordicDemographic/cv')

}

#compare_cv_continuous(long, cv1, 'alcbase')
```

## Remove files
```{r}
rm(cv1, cv1.alcbase.sum, ldown, long, long.alcbase.sum)
```

# LYs and DFLYs

We are going to calculate Life Years and Disability Free Life Years for the full population outputs, then we can compare the baseline case to the Minimum Unit Price intervention. We need detailed output for this, and there is a script in stata that has done this for a cohort output that I think we can learn from. Some of the key points I know about this so far:

- We need to create a dummy variable for each person to determine whether they are alive at each wave. Summing this variable per person will be the LIFEYEARS
- Following this, we can create a dummy var for each person to determine whether they have a chronic disease and another for disability. Summing these and calculating Life Years minus x will give us disease and disability free Life Years

First get hold of the appended detailed output data.
```{r}
# path <- 'input_data/detailed_output/'
# 
# base <- read_dta(paste0(path, 'ELSA_cohort_append.dta'))
# mup <- read_dta(paste0(path, 'ELSA_alcInt_cohort_append.dta'))
```

```{r}

## Think it through - Steps:
# 1. Add a variable for lifeyears
  # Var needs to == 1 if not died and 0 if died
  # With this var we can calculate the total number of lifeyears for each person
# 2. Calculate lifeyears per person per repetition
  # Sum lifeyears per person
  # Divide sum lifeyears by number of reps

# generate lifeyear for people who have not died
# base$lifeyear <- 1 - base$died
# # generate the number of repetitions for each person
# base$nreps <- max(base$mcrep) + 1
# # now generate the lifeyear intermediate (lifeyears per person)
# base2 <- base %>% group_by(hhidpn, year) %>%
#                     mutate(nreps = n()) %>%
#                     ungroup() %>%
#                     group_by(hhidpn) %>%
#                     mutate(n_LY = sum(lifeyear),
#                            mean_LY = n_LY / nreps) %>%
#                     mutate(n_DFLY = sum(not_disabled),
#                            mean_DFLY = n_DFLY / nreps) %>%
#                     mutate(two_yr_LY = mean_LY * 2,
#                            two_yr_DFLY = mean_DFLY * 2)
#   
# 
# mean <- base2 %>% summarise(LY = mean(mean_LY),
#                             DFLY = mean(mean_DFLY))

```

```{r}
source('../visualisation_utilities.r')
```

# MultipleContinousOut

```{r}
vis_multi_continuous_out_compare <- function(data1, data2, prefix, var, suffix.list, horizon = 50, save = FALSE) {

  # calculate horizon
  start.yr <- 2012
  end.yr <- start.yr + horizon

  # add prefix and suffix
  key.var.list <- lapply(1:length(suffix.list), function(k) paste0(prefix, '_', var, '_', suffix.list[[k]]))

  data1.select <- data1 %>% select(year, all_of(unlist(key.var.list))) %>%
                            rename(Baseline_Moderate = key.var.list[[1]],
                                   Baseline_Increasing = key.var.list[[2]],
                                   Baseline_High = key.var.list[[3]]) %>%
                          filter(year < end.yr) %>%
                          filter(year > start.yr + 2)
  data2.select <- data2 %>% select(year, all_of(unlist(key.var.list))) %>%
                          rename(Intervention_Moderate = key.var.list[[1]],
                                 Intervention_Increasing = key.var.list[[2]],
                                 Intervention_High = key.var.list[[3]]) %>%
                          filter(year < end.yr) %>%
                          filter(year > start.yr + 2)

  merged <- merge(data1.select, data2.select, by = 'year')

  pivoted <- pivot_longer(data = merged,
                          cols = Baseline_Moderate:Intervention_High,
                          names_sep = '_',
                          names_to = c('Scenario', 'Subgroup'),
                          values_to = 'Proportion')

  #ggplot(pivoted, aes(x = year, colour = Scenario)) +
  #  geom_line(aes(y = Proportion))

  p <- ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
        geom_line() +
        scale_linetype_manual(values = c("TRUE" = "solid", "FALSE" = "dashed")) +
        guides(linetype = FALSE)

  if(save) {
    ggsave(filename = 'alcbase_comparison_multisuffix.png',
           path = 'Deliverables/Conferences/NordicDemographic/ordinal_comparisons')
  }

  p
}
```

```{r}
data1 <- base
data2 <- mup
prefix <- 'a'
var <- 'alcbase'
suffix.list <- c('moderate', 'increasingRisk', 'highRisk')
horizon <- 50

vis_multi_continuous_out_compare(data1, data2, prefix, var, suffix.list, horizon = 30)
```

```{r}
data1 <- base
data2 <- mup
prefix <- 'a'
var <- 'alcbase'
suffix.list <- c('moderate', 'increasingRisk', 'highRisk')
horizon <- 50

# calculate horizon
start.yr <- 2012
end.yr <- start.yr + horizon

# add prefix and suffix
key.var.list <- lapply(1:length(suffix.list), function(k) paste0(prefix, '_', var, '_', suffix.list[[k]]))

data1.select <- data1 %>% select(year, all_of(unlist(key.var.list))) %>%
                          rename(Baseline_Moderate = key.var.list[[1]],
                                 Baseline_Increasing = key.var.list[[2]],
                                 Baseline_High = key.var.list[[3]]) %>%
                        filter(year < end.yr) %>%
                        filter(year > start.yr + 2)
data2.select <- data2 %>% select(year, all_of(unlist(key.var.list))) %>%
                        rename(Intervention_Moderate = key.var.list[[1]],
                               Intervention_Increasing = key.var.list[[2]],
                               Intervention_High = key.var.list[[3]]) %>%
                        filter(year < end.yr) %>%
                        filter(year > start.yr + 2)

merged <- merge(data1.select, data2.select, by = 'year')

pivoted <- pivot_longer(data = merged,
                        cols = Baseline_Moderate:Intervention_High,
                        names_sep = '_',
                        names_to = c('Scenario', 'Subgroup'),
                        values_to = 'Proportion')

#ggplot(pivoted, aes(x = year, colour = Scenario)) +
#  geom_line(aes(y = Proportion))

ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
  geom_line() +
  scale_linetype_manual(name = 'Scenario', values = c("TRUE" = "solid", "FALSE" = "dashed"), guide = 'legend', labels = c('Baseline', 'Intervention')) +
  guides(linetype = FALSE)

ggplot(pivoted, aes(x = year, y = Proportion, group = interaction(Scenario, Subgroup), colour = Subgroup, linetype = Scenario %in% c("Baseline"))) +
  geom_line() +
  scale_linetype_manual(name = 'Scenario', values = c("TRUE" = "solid", "FALSE" = "dashed"), guide = 'legend', labels = c('Intervention', 'Baseline'))

#ggplot(pivoted, aes(x = year, y = Proportion, group =))
```

# Reduction

Now producing graphs to show reduction in disease prevalence over time because the prevalence graphs are a bit weird.

```{r}

base1 <- base
mup1 <- mup

base2 <- base1 %>% select(year, p_hearte_all, p_hearte_moderate, p_hearte_increasingRisk, p_hearte_highRisk) %>%
                    rename(Baseline_All = p_hearte_all, 
                           Baseline_Moderate = p_hearte_moderate, 
                           Baseline_IncreasingRisk = p_hearte_increasingRisk,
                           Baseline_HighRisk = p_hearte_highRisk)

mup2 <- mup1 %>% select(year, p_hearte_all, p_hearte_moderate, p_hearte_increasingRisk, p_hearte_highRisk) %>%
                    rename(Intervention_All = p_hearte_all, 
                           Intervention_Moderate = p_hearte_moderate, 
                           Intervention_IncreasingRisk = p_hearte_increasingRisk,
                           Intervention_HighRisk = p_hearte_highRisk)

merged <- merge(base2, mup2, by = 'year')

differenced <- merged %>% mutate(Moderate = Baseline_Moderate - Intervention_Moderate,
                                 IncreasingRisk = Baseline_IncreasingRisk - Intervention_IncreasingRisk,
                                 HighRisk = Baseline_HighRisk - Intervention_HighRisk)

transformed <- differenced %>% group_by(year) %>%
                                summarise(mean_Mod = mean(Moderate),
                                          min_Mod = min(Moderate),
                                          max_Mod = max(Moderate),
                                          mean_Inc = mean(IncreasingRisk),
                                          min_Inc = min(IncreasingRisk),
                                          max_Inc = max(IncreasingRisk),
                                          mean_High = mean(HighRisk),
                                          min_High = min(HighRisk),
                                          max_High = max(HighRisk),)

pivoted <- pivot_longer(data = transformed,
                        cols = mean_Mod:max_High,
                        names_sep = '_',
                        names_to = c('Stat', 'Subgroup'),
                        values_to = 'Prevalence')

ggplot(pivoted, aes(x = year, y = Prevalence, group = interaction(Stat, Subgroup), colour = Subgroup)) +
  geom_line()

```

```{r}

```