### 2022-12-05 FEM_R/Loneliness_Social_Isolation/PAPER_WORKING_SCRIPT.Rmd
Added to the utilities script, and created new R Notebook for the loneliness and social isolation work

---
title: "Social_Isolation_and_Loneliness"
output: html_notebook
---

In this notebook, we prepare all the figures and tables for the paper 'Projecting the impact of loneliness and social isolation on the health outcomes of England's future elderly population'. We want to investigate how loneliness and social isolation will affect health outcomes, both together and separately, when we project an elderly population into the future using the English Future Elderly model.

# Setup

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/E_FEM_clean/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(reactable)
require(haven)
require(ggplot2)
require(jtools)

```

## Data

```{r}
# Directories
base_output_dir <- "output/COMPLETE/"
input_dir <- "input_data/"

source('FEM_R/utilities.R')

# Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))

# Outputs
base_full <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort_full <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
```

# Sample Statistics

In this section, we will calculate a lot of sample statistics on the stock, replenishing, and transition populations, and generate a table that can be easily copied into latex. We will also investigate the distribution of both the loneliness and social isolation measures we have added to the FEM, and particularly look to see how both loneliness and social isolation is distributed among different demographic groups. I am also interested in the individual elements of the index of social isolation, because I want to investigate some of the reasons why social isolation may be distributed differently to loneliness.

## Sample Statistic Calculations

We want sample statistics on a number of characteristics of our input populations:
- Demographics
    - Age
    - BMI
    - Gender
    - Education
- Married/cohabiting %
- Loneliness score
- Social Isolation Index
- Risk Behaviours
    - Smoking
    - Alcohol frequency
    - Physical Activity
- Mental conditions
    - Depression
    - More?
- Chronic Diseases
    - Any
    - Hypertension
    - CVD
    - Diabetes
    - Cancer
    - Alzheimers/Dementia

Samples must be weighted by survey cross-sectional weight - `cwtresp`

```{r, echo=FALSE}
### Going to update the previous function from utilities.R here
# copy the data so we don't overwrite anything
s <- stock

# try it first to see what we get
#weighted.survey.means(s)

weighted.survey.means <- function(init.pop, transition=FALSE) {

  require(survey)

  if(!transition) {
    design <- svydesign(id = ~idauniq,
                        weights = ~weight,
                        data = init.pop)
  } else if(transition) {
    # Error from missing cwtresp var - drop all that are missing to solve
    init.pop <- init.pop[complete.cases(init.pop$cwtresp),]
    design <- svydesign(id = ~idauniq,
                        weights = ~cwtresp,
                        data = init.pop)
  }
  print(svymean(~age, design, na.rm=TRUE))
  print(svyvar(~age, design, na.rm=TRUE))
  print(svymean(~male, design, na.rm=TRUE))
  print(svymean(~bmi, design, na.rm=TRUE))
  print(svyvar(~bmi, design, na.rm=TRUE))
  print(svymean(~smoken, design, na.rm=TRUE))
  print(svymean(~smokev, design, na.rm=TRUE))
  print(svymean(~hsless, design, na.rm=TRUE))
  print(svymean(~college, design, na.rm=TRUE))
  print('---------------------')
  print(svymean(~cancre, design, na.rm=TRUE))
  print(svymean(~diabe, design, na.rm=TRUE))
  print(svymean(~hearte, design, na.rm=TRUE))
  print(svymean(~stroke, design, na.rm=TRUE))
  print(svymean(~lunge, design, na.rm=TRUE))
  print(svymean(~hibpe, design, na.rm=TRUE))
  print(svymean(~alzhe, design, na.rm=TRUE))
  print(svymean(~demene, design, na.rm=TRUE))
  print(svymean(~hchole, design, na.rm=TRUE))
}

s <- s %>% filter(!is.na(s$strat))

weighted.survey.means(s)

```

```{r}
## Lets try to make a table
require(survey)

# First create survey design object for weighted stats
design <- svydesign(id = ~idauniq,
                    weights = ~weight,
                    data = s)

# Now calculate means (and SE)
# Age + Sex + BMI + Smoking Status + Education MEANS
means <- svymean(~age+male+bmi+smoken+smokev+hsless+college, design, na.rm=TRUE)
mean.df <- as.data.frame(means, optional = TRUE)
mean.df <- tibble::rownames_to_column(mean.df, 'Variable')

# Need counts of non missing data per variable to calculate SD from SE
counts <- s %>% summarise_each(funs(sum(!is.na(.)))) %>%
  select(age, male, bmi, smoken, smokev, hsless, college) %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column('Variable') %>%
  rename(n = 'V1')

# Lovely, now merge
mean.df <- merge(counts, mean.df, by = 'Variable')

# Now we can calculate SD
mean.df$SD <- mean.df$SE * sqrt(mean.df$n)

print(mean.df)
```

## Distribution of Lnly
```{r}

```

## Distribution of SocIso

## Latex outputs

# Results

## Functions

## Baseline

### Life Years

### Disability-free Life Years

### Disease-free Life Years

## Interventions