### 2023-02-24 FEM_R/variable_testing/workstat.Rmd
Started a directory for variable testing in R. Added script for workstat and physically_active

---
title: "Workstat"
output: html_notebook
---

```{r setup, include=FALSE}
# Need to set up the working directory
workingDir <- "/home/luke/Documents/WORK/E_FEM/E_FEM/"
knitr::opts_knit$set(root.dir = workingDir)

require(tidyverse)
require(reactable)
require(haven)
require(ggplot2)
require(jtools)
require(ggfortify)
require(survival)
require(survminer)
require(lazyeval)

```

```{r}
source('../HAZARDS/source/hazard_utils.R')
source('FEM_R/utilities.R')
```

```{r}
## Directories
base_output_dir <- "output/COMPLETE/"
input_dir <- "input_data/"
## Inputs
long <- read_dta(paste0(input_dir, 'ELSA_long.dta'))
stock <- read_dta(paste0(input_dir, 'ELSA_stock.dta'))
repl <- read_dta(paste0(input_dir, 'ELSA_repl.dta'))
trans <- read_dta(paste0(input_dir, 'ELSA_transition.dta'))
## Outputs
### Baseline
base <- read_dta(paste0(base_output_dir, 'ELSA_core_base/ELSA_core_base_summary.dta'))
cohort <- read_dta(paste0(base_output_dir, 'ELSA_core_cohort/ELSA_core_cohort_summary.dta'))
### Cross-Validation
cv1 <- read_dta(paste0(base_output_dir, 'ELSA_CV1/ELSA_CV1_summary.dta'))
#cv1.full <- read_dta(paste0(base_output_dir, 'ELSA_CV1/detailed_output/ELSA_CV1_summary.dta'))
cv2 <- read_dta(paste0(base_output_dir, 'ELSA_CV2/ELSA_CV2_summary.dta'))
### Interventions
#no.lonely <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_cohort/ELSA_nolnly_cohort_summary.dta'))
#no.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nosociso_cohort/ELSA_nosociso_cohort_summary.dta'))
#no.lnly.sociso <- read_dta(paste0(base_output_dir, 'ELSA_nolnly_nosociso_cohort/ELSA_nolnly_nosociso_cohort_summary.dta'))
```

Newly recoded var is 3 levels:

- Employed
- Inactive
- Unemployed

Model won't run, saying l2inactive is not a valid variable. First thing to check
is that the variable has been properly created and check balance.

```{r}
# First long file
workstat.long <- long %>% 
  select(idauniq, wave, workstat, l2workstat, employed, 
         inactive, retired, l2employed, l2inactive, l2retired) %>%
  group_by(wave) %>%
  summarise(employed = sum(employed),
            inactive = sum(inactive),
            retired = sum(retired)) %>%
  pivot_longer(cols = employed:retired,
               names_to = 'workstat',
               values_to = 'count')

test.long$wave <- as.integer(test.long$wave)
```

```{r}
ggplot(data = workstat.long, aes(x = wave, y = count, group = workstat, color=workstat)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Count')
```

#### Outputs

```{r}
# First long file
test.CV1 <- cv1 %>% 
  select(year, a_workstat_all, p_employed_all, 
         p_inactive_all, p_retired_all) %>%
  group_by(year) %>%
  summarise(employed = sum(p_employed_all),
            inactive = sum(p_inactive_all),
            retired = sum(p_retired_all)) %>%
  pivot_longer(cols = employed:retired,
               names_to = 'workstat',
               values_to = 'prevalence')

test.long$wave <- as.integer(test.long$wave)
```

```{r}
ggplot(data = test.CV1, aes(x = year, y = count, group = workstat, color=workstat)) +
  geom_line() +
  scale_x_continuous(breaks=c(1:9)) +
  labs(title = 'Workstat in ELSA') +
  xlab('Wave') +
  ylab('Prevalence')
```